### 2017-08-15 Ben_Manuscripts/Masters_Bypass/master.tex
A separate folder for my Master's Bypass paper

\documentclass[journal=ancac3]{achemso}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{geometry}
\usepackage{amsmath} % or simply amstext
\usepackage{siunitx}
\usepackage{booktabs}
\newcommand{\angstrom}{\textup{\AA}}

\title{Predicting Transport in Lyotropic Liquid Crystal Membranes with Molecular Dynamics Simulations}
\author{Benjamin J. Coscia}
\email{benjamin.coscia@colorado.edu}
\author{Douglas L. Gin}
\author{Richard D. Noble}
\author{Michael R. Shirts}
\affiliation{University of Colorado, Boulder, Colorado 80309, United States}

\begin{document}

\graphicspath{{./figures/}}  % put all the figures here
\maketitle

\include{intro}
\include{results}
\include{conclusion}
\include{methods}
%\include{supp} supplementary info should be included as a separate document for ACSnano

\newpage
\bibliography{llc.bib}
\end{document}