### 2017-08-15 Ben_Manuscripts/Masters_Bypass/conclusion.tex
Create conclusion.tex

The procedure used to create and validate our model can be used
to evaluate other liquid crystalline assemblies. Using the design 
framework and analysis methods applied herin, we have the ability 
to reliably predict structures of new nanoporous membranes.