### 2017-08-15 Ben_Manuscripts/Masters_Bypass/conclusion.tex
additions

\section{Conclusion}

In this work, we have illustrated a more detailed picture of the structure 
of a self-assembled thermotropic liquid crystal membrane using an atomistic 
molecular model. Using the model we have learned that these liquid crystal
monomers prefer to stack in a sandwiched configuration with five monomers
in each layer. By a thorough analysis of the membrane structure, we have
discovered the existence of two metastable basins that both persist at room
temperature. Structurally, Basin A matches closest to experiment, while 
Basin B has not been seen experimentally. We hypothesize that Basin B will
be dominant at higher temperatures. This hypothesis will be tested in future 
work. In addition to structural accuracy, our model's physical properties
are consistent with experimental measurements. Ionic conductivity, calculated
in two ways, is in reasonable agreement with experiment. The procedure used
to create and validate our model can be used to evaluate other liquid 
crystalline assemblies. Using the design framework and analysis methods 
applied herin, we have the ability to understand structures of new and
unsynthesized nanoporous membranes.