### 2017-08-15 Ben_Manuscripts/Masters_Bypass/master.tex
Formatting update

\documentclass{article}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{geometry}
\usepackage{amsmath} % or simply amstext
\usepackage{siunitx}
\usepackage{booktabs}
\newcommand{\angstrom}{\textup{\AA}}

\title{Predicting Transport in Lyotropic Liquid Crystal Membranes with Molecular Dynamics Simulations}
\author{Benjamin J. Coscia}
\author{Douglas L. Gin}
\author{Richard D. Noble}
\author{Michael R. Shirts}

\begin{document}

\graphicspath{{./figures/}}  % put all the figures here
\maketitle

\input{intro}
\input{methods}
\input{results}
\input{conclusion}

%\include{supp} supplementary info should be included as a separate document for ACSnano

\newpage
\bibliography{llc.bib}
\end{document}