### 2018-06-16 Ben_Manuscripts/structure_paper/supporting_information.tex
spellchecking.

\documentclass[journal=jpcbfk,manusciprt=article]{achemso}

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{amsmath} 
%\usepackage[margin=1in]{geometry}\usepackage{amsmath} % or simply amstext
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\newcommand{\angstrom}{\textup{\AA}}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}
\usepackage{listings}
\usepackage{xr}

\SectionNumbersOn

\externaldocument[M-]{draft}
%MRS8: can use document class here - looks weird with title/authors as written.
\title{Supporting Information : Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes}
\author{Benjamin J. Coscia \and Douglas L. Gin \and Richard D. Noble \and Joe Yelk \and Matthew Glaser \and Xunda Feng \and Michael R. Shirts}

\begin{document}

%  \bibliographystyle{ieeetr}
  \graphicspath{{./figures/}}  % put all the figures here
  \maketitle

  \section{Attempted Self-assembly}\label{section:self_assembly}

  We attempted self assembly of Na-GA3C11 monomers by using an isotropic configuration
  generated using Packmol \cite{martinez_packmol:_2009} (Figure~\ref{fig:isotropic_box}).
  The input file given to Packmol is shown below.

  \lstset{language=bash}
  \begin{lstlisting}
  tolerance 2.0
  output packed.pdb
  filetype pdb
  structure monomer.pdb
  		number 400
  		inside box 0. 0. 0. 84.3 84.3 84.3
  end structure
  \end{lstlisting}

  We used the same number of monomers (400) that we used in all other simulations. 
  Since Packmol does not have the capability to create monoclinic boxes, we used a cubic
  box with a volume of 660 $nm^3$, 25\% larger than those used to create ordered unit
  cells we studied. We tested systems with semi-isotropic and anisotropic pressure 
  coupling since the shape of the unit cell will likely need to change in order to 
  accommodate hexagonally packed pores.

  We use the nematic order parameter as defined in section~\ref{method:nematic_order} to 
  observe any progress towards system ordering. In both cases, we see that the nematic
  order parameter stays close to zero for the duration of the simulation time, indicating
  an isotropic arrangement of head groups (Figure~\ref{fig:nematic_isotropic}). The values
  are slightly negative since there are only 400 monomers. 
  Equation~\ref{eqn:nematic_order_parameter} is bounded between -1 and 1, however, with a
  proper choice of nematic director, measured values will range between 0 and 1. With a 
  large number of samples, an isotropic system would have an order parameter of 0.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{isotropic_box.png}
  \caption{}\label{fig:isotropic_box}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{nematic_isotropic.png}
  \caption{}\label{fig:nematic_isotropic}
  \end{subfigure}
  \caption{(a) We created a box of isotropically packed monomers and allowed it to simulate
  for 100 ns using isotropic and anisotropic pressure coupling. (b) The nematic order 
  parameter hovers close to zero for the duration of the simulation meaning the system
  maintains its isotropic alignment.}\label{fig:self_assembly}
  \end{figure}

  \section{Further details regarding monomer parameterization}\label{section:parameterization}

  \vspace{1em}
  We parameterized monomers according to the following procedure:
  \begin{enumerate}
	\item \textit{Create monomer structure file with connectivity} : We
	drew atomistic structures using MarvinSketch
	17.13~\cite{chemaxon_marvinsketch_2017} with all hydrogen atoms drawn out
	explicitly. We optimized the 3D geometry of the structure using the 'Clean in
	3D' function of MarvinSketch.  We saved the structure as a .mol file, then
	converted it to .pdb format using Open Babel 2.4.1
	\cite{oboyle_open_2011,noauthor_open_nodate}. 
	\item \textit{Assign GAFF atomtypes using \texttt{antechamber}} : Using
	the .pdb structure file as input, we ran
	\texttt{antechamber}~\cite{wang_automatic_2006} using the AM1-BCC charge model.
	The net charge on the monomer is input as -1 since the sodium ion is kept as 
	a separate residue. We use \texttt{LEaP}~\cite{case_ambertools16_2016} and the
	output of \texttt{antechamber} to create Amber topology files. A detailed
	tutorial can be accessed elsewhere \cite{walker_antechamber_nodate}.
	\item \textit{Create GROMACS topologies from Amber output} : The output
	of \texttt{LEaP} is a .inpcrd and a .prmtop file which are Amber topology
	files. Using acpype.py \cite{sousa_da_silva_acpype_2012}, we converted the
	\texttt{LEaP} output into GROMACS .gro and .top files. 
	\item \textit{Perform a simulated annealing procedure on the monomer} :
	We created a cubic box around the monomer using the GROMACS command \texttt{gmx
	editconf}. The monomer was centered in the box with edges of the
	box spaced at least 3 nm from the monomer on all sides. We ran an energy minimization
	on the system with the steepest descent algorithm. Next we performed an NVT
	simulated annealing procedure. We linearly decreased the temperature of the
	system from 1000 K to 50 K over the course of 10 ns. We randomly chose a monomer
	configuration from the last 10 \% of the trajectory. 
	\item \textit{Reassign charges with \texttt{molcharge}}: With the monomer
	configuration taken from the annealed trajectory, we reassigned charges using
	\texttt{molcharge} with the am1bccsym method in order to ensure charges 
	are symmetric. This condition is not guaranteed with \texttt{antechamber}. 
	The charges in the GROMACS topology file (.top) were replaced with the 
	new charges calculated by \texttt{molcharge}. 
	\item \textit{Anneal again to get final structure} : We repeated the
	same simulated annealing procedure using the monomer topology with
	\texttt{molcharge} charges. A random monomer configuration was pulled from the
	last 10 \% of the trajectory and was used to build all assemblies reported
	(Figure~\ref{fig:monomer}).
  \end{enumerate}

  \begin{figure}[!htb]
	\centering
        \includegraphics[width=0.9\textwidth]{monomer.png}
	\caption{Atomistic representation of the monomer Na-GA3C11. White atoms
		represent hydrogen, cyan atoms represent carbon, red atoms represent oxygen and
		the blue atom is sodium.}\label{fig:monomer}
  \end{figure}

  \begin{figure}[!htb]
	\centering
        \includegraphics[width=0.9\textwidth]{monomer_color_coded.png}
	\caption{The groups used for radial distribution calculations. Red atoms are in the
		tails group. Green atoms are in the head group region. The blue atom is sodium. 
		}\label{fig:monomer_color_coded}
  \end{figure}

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.5\textwidth]{monomer_dihedrals.png}
  \caption{We measured the autocorrelation functions of the dihedrals highlighted in
  black. We only included dihedrals associated with the outer tails in our calculation
  since the position of the center tail leads to fundamentally different interactions and
  dynamics.}\label{fig:dihedrals}
  \end{figure}

  \begin{figure}[!htb]
        \centering
        \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=\textwidth]{sandwiched.png}
                \caption{}\label{fig:sandwiched}
        \end{subfigure}
        \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=\textwidth]{PD.png}
                \caption{}\label{fig:pd}
        \end{subfigure}
        \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=\textwidth]{Tshaped.png}
                \caption{}\label{fig:tshaped}
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.475\textwidth}
                \centering
                \includegraphics[width=\textwidth]{sandwichedlayers.png}
                \caption{}\label{fig:sandwichedlayers}
        \end{subfigure}
        \begin{subfigure}[b]{0.475\textwidth}
                \centering
                \includegraphics[width=\textwidth]{offsetlayers.png}
                \caption{}\label{fig:offsetlayers}
        \end{subfigure}
        \caption{(a) Sandwiched benzene dimers stack 3.8 \angstrom~apart. (b) Parallel-Displaced benzene dimers stack
        3.4 \angstrom~vertically and 1.6 \angstrom~horizontally apart. (c) T-shaped benzene dimers stack 5.0 \angstrom~apart.
        (d) Two monomer layers stacked in the sandwiched configuration (e) Two monomer layers stacked in the parallel-displaced
        configuration }\label{fig:stacking}
  \end{figure}

  \begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{pore_radius_illustration.png}
	\caption{When creating an initial configuration, the pore radius is defined based
	on the distance of the carbonyl carbon from the pore's central axis}
	\label{fig:pore_radius_illustration}
  \end{figure}

  %Our model is insensitive to initial configuration if monomer placement is chosen within reason.

  \section{Number of monomers per column}\label{section:monomers_per_column}

  We chose to build all systems using 20 monomers per column. We made this choice
  by considering a number of factors that directly affect the quality of our results. 
  Clearly, choosing a small number of layers will cause the system to suffer from 
  finite size effects. Even in our 20 monomer-per-column system, axial correlations
  between stacked monomers persist throughout the entire correlation function (See
  Figure ~\ref{M-fig:correlation}). The correlation function of the system which we
  built with 40 monomers per column continues to oscillate for its full length
  (Figure~\ref{fig:tall_correlation_full}). If we continue adding monomers, the 
  correlation function will eventually fully decay but it is not worth the 
  computational expense. Regardless, the correlation lengths calculated from 20
  monomer per column simulations are in sufficient agreement with experiment.

  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{tall_correlation_full.png}
  \caption{Oscillations persist throughout the length of the correlation 
  function for the sandwiched system in the ordered basin built with 40 monomers
  per column. The full system is ca. 17 nm tall so the full correlation function
  covers half this length since periodicity forces the remaining length to be its
  mirror image.}\label{fig:tall_correlation_full}
  \end{figure}

  We may be able to simulate systems using less than 20 monomers per column, but chose
  to use 20 since it gives fairly good resolution when simulating X-ray diffraction 
  patterns. The size of the Fourier space bin in each dimension is determined by 
  $dfrac{2\pi}{L}$ where L is the length of a given box vector. The z-box vector 
  for 20 monomer-per-column systems is $\approx$ 85 \AA which means we will see 
  a reciprocal space resolution of 0.074 \AA$^{-1}$ in the z direction. Further decreasing
  the resolution would increase the noise in the simulated patterns and make it more
  difficult to measure the intensity of reflections of interest.

  \section{Initial Configuration Dependence}\label{section:initial_config_dependence}
%  \begingroup
%	\fontsize{14pt}{12pt}\selectfont
%	\textbf{Initial Configuration Dependence}
%  \endgroup

  \vspace{1em}
  We addressed any major dependence on initial configuration in the main text.
  There we showed that systems are stable when made with 4, 5, 6, 7, and 8
  monomers per layer. We also showed that systems are stable when monomer head
  groups are oriented in the parallel displaced and sandwiched configurations
  (See Figure~\ref{fig:stacking}). Our model is most consistent with experiment
  when built with 5 monomers per layer, with layers stacked 3.7 \AA~apart in 
  the parallel displaced configuration.

  There are three other parameters whose choice may influence the equilibrium
  structure: initial pore spacing, initial pore radius and initial distance
  between layers. Here we show the results of a sensitivity analysis performed
  on the three parameters. To reduce the size of the sensitivity analysis, we
  only tested systems built with 5 monomers per layer with layers stacked in 
  the parallel displaced configuration. We equilibrated all systems according
  to the dry equilibration procedure.

  %\begin{enumerate}

  %	  \item \underline{Initial pore spacing}
  \subsection{Initial pore spacing}\label{section:initial_pore_spacing}

	  We tested five different initial pore spacings, defined as the
	  distance between the central axis of each pore with all others
	  (Figure~\ref{fig:p2p_diagram}). To reduce the number of variables, we held the
	  pore radius constant, at 6 \AA~and the distance between layers at 3.7 \AA~since
	  those were the values used in our optimal system in the main text. We
	  prioritized ensuring that resulting configurations maintain the expected
	  hexagonal symmetry. If initial pore spacing is too small, we observe repulsion
	  between columns which disrupts equilibration. If the pore spacing is too large
	  the pores squeeze together, but in a distorted hexagonal array because the xy
	  initial translation of the pores is somewhat erratic.

%	  If the initial pore spacing is chosen to be
%	  too small, pore columns abruptly repel each other as soon as they are able to
%	  overcome the restraining potential present as part of the dry equilibration
%	  procedure. We avoided using systems that exhibit this behavior since the
%	  behavior increases the chances of a system to become kinetically trapped in an
%	  undesired metastable free energy basin.

	  \begin{enumerate}

		\item 39 \AA~: We tested a pore spacing of 39 \AA~in order to
		have a test system with an initial spacing below the experimental value. As soon
		as the restraining potential switches to 56 kJ mol$^{-1}$ nm$^{-2}$, the
		columns are able to repel resulting in a large jump in pore spacing
		(Figure~\ref{fig:p2p_39}). %After 400 ns, the pore spacing is 4.05 $\pm$ 0.17 nm.
%		There is a large uncertainty in this value because the 1-4 pore spacing
%		(Figure~\ref{fig:p2p_diagram}) deviates significantly from all other pore
%		spacings. The 1-4 pore spacing is 4.31 nm while the average of all other pore
%		spacings is 3.96 nm. The system deviates from hexagonal geometry and is trending
%                towards cubic packing.

		\item 41 \AA~: We chose to test a pore spacing of 41 \AA~since
		it closely matches the experimental pore spacing. Again, we observe abrupt
		repulsion of columns once the restraining potential is switched to 56 KJ
		mol$^{-1}$ nm$^{-2}$ (Figure~\ref{fig:p2p_41}). %After 400 ns, the
%		average pore spacing settles below the experimental value at 4.07 $\pm$ 0.21
%		nm. Again, a large uncertainty in this value occurs because the 1-4 pore spacing
%		(Figure~\ref{fig:p2p_diagram}) deviates significantly from all other pore
%		spacings. The 1-4 pore spacing is 4.44 nm while the average of all other pore
%		spacings is 3.96 nm.

		\item 45 \AA~: A pore spacing of 45 \AA~is about 10 \% larger
		than the experimental value. This systems exhibits the smallest response to our 
 		equilibration procedure (Figure~\ref{fig:p2p_45}) and we chose to use this
        value for all our simulations in the main text.
        %The equilibrated pore spacing is 4.23 $\pm$ 0.03 nm.
		%The system maintains hexagonal symmetry and exhibits the lowest uncertainty in
		%pore spacing. We chose to use this value for all our simulations in the main
		%text.

		\item 50 \AA~: We tested a pore spacing of 50 \AA, about 20 \%
		larger than the experimental value. Once the restraining potential reaches 56
		KJ mol$^{-1}$ nm$^{-2}$, the pore spacing begins to decrease rapidly, following
		a linear trend (Figure~\ref{fig:p2p_50}).
%		After 400 ns, the average pore spacing is 4.37 $\pm$ 0.13 nm.
%		There is a large
 %               uncertainty with this value because, again, the array of pores deviates from 
 %               hexagonal character. Using Figure ~\ref{fig:p2p_diagram} as a reference, pore 
  %              2 drifts outward from the array and forces the 1-4 distance to shrink in 
   %             response.

		\item 55 \AA~: We tested a pore spacing of 55 \AA~which is at
		a distance where monomers in each pore no longer intersect those of adjacent pores. Once
		the restraining potential reaches 56 KJ mol$^{-1}$ nm$^{-2}$, the pore spacing
		changes erratically until it begins to settle when the force constants are
		below 3 KJ mol$^{-1}$ nm$^{-2}$ (Figure~\ref{fig:p2p_55}). We recommend
		avoiding a system such as this where vacuum gaps between pore columns are
		introduced unnecessarily. %The system equilibrates to a pore spacing of 4.27
%		$\pm$ 0.18 nm. The 1-2 pore spacing is the outlier in this case with a pore
 %               spacing of 3.99 \AA~ which distorts the hexagonal array.

	  \end{enumerate}

	  \begin{figure}[!htb]
		\centering
		\includegraphics[width=0.4\textwidth]{p2p_diagram.png}
		\caption{In a perfect hexagonal, there are 5 distances that should be exactly
		equal. As pictured in this diagram the distance from 1-2, 1-3, 1-4, 2-4, and 3-4
		are the same.}~\label{fig:p2p_diagram}
	  \end{figure}

	  %BJC: Could put vertical lines to indicate where position restraint values are reduced
  	  \begin{figure}[!htb]
		\centering
		\begin{subfigure}{0.3\textwidth}
			\includegraphics[width=\textwidth]{p2p_39.png}\quad
			\vspace{-1.25em}
			\caption{39 \AA~initial pore spacing}~\label{fig:p2p_39}
		\end{subfigure}
		\begin{subfigure}{0.3\textwidth}
			\includegraphics[width=\textwidth]{p2p_41.png}\quad
			\vspace{-1.25em}
			\caption{41 \AA~initial pore spacing}~\label{fig:p2p_41}
		\end{subfigure}
		\begin{subfigure}{0.3\textwidth}
			\includegraphics[width=\textwidth]{p2p_45.png}
			\vspace{-1.25em}
			\caption{45 \AA~initial pore spacing}~\label{fig:p2p_45}
		\end{subfigure}
		\begin{subfigure}{0.3\textwidth}
			\includegraphics[width=\textwidth]{p2p_50.png}\quad	
			\vspace{-1.25em}
			\caption{50 \AA~initial pore spacing}~\label{fig:p2p_50}
		\end{subfigure}
		\begin{subfigure}{0.3\textwidth}
			\includegraphics[width=\textwidth]{p2p_55.png}
			\vspace{-1.25em}
			\caption{55 \AA~initial pore spacing}~\label{fig:p2p_55}
		\end{subfigure}

		\caption{The pore spacing during the restrained portion of the
			dry equilibration procedure is shown. Every 50 ps (0.05 ns) position restraints
			are reduced according to the sequence: 1000000, 3162, 56, 8, 3, 2, 1, 0 KJ
			mol$^{-1}$ nm$^{-2}$. When the initial pore spacing is chosen below the
			experimental value (a) or at the experimental value (b), there is an abrupt
			change in pore spacing when the position restraints are reduced to 56 KJ
			mol$^{-1}$ nm$^{-2}$. When pores are started 45 \AA~apart (c) the pore spacing
			remains relatively stable. When pores are spaced 50 \AA~apart (d), the pore
			spacing decreases nearly linearly once the restraints are reduced to 56 KJ
			mol$^{-1}$ nm$^{-2}$. When pores are spaced 55 \AA~apart (e), so that monomers
			do not intersect with adjacent pores, and position restraints are reduced to 56
			KJ mol$^{-1}$ nm$^{-2}$, the pore spacing changes erratically before stabilizing
			when force constants are reduced below 3 KJ mol$^{-1}$ nm$^{-2}$.} 
		\label{fig:p2p}
	  \end{figure}

%	  \item \underline{Initial pore radius}
	  \subsection{Initial pore radius}\label{section:initial_pore_radius}

	  We tested 3 different pore radii, defined as shown in
	  Figure~\ref{fig:pore_radius_illustration}. For each system, we held the initial
	  pore spacing constant at 45 \AA~and the initial distance between layers
	  constant at 3.7 \AA. Equilibrated values of pore radii are presented in
	  Table~\ref{table:radii}. The pore radius in each simulation frame is calculated
	  as the average distance of all carbonyl carbons (See
	  Figure~\ref{fig:pore_radius_illustration}) from their associated pore center.
	  Statistics for the pore radii reported were generated from the time series
	  representing the average pore radius at each frame. We detected equilibration
	  using \texttt{pymbar.timeseries.detectEquilibration}. We calculated the average
	  and standard deviation of pore radii using only data points collected after
	  equilibration was detected.

	  \begin{enumerate}

		\item 2.5 \AA~: The smallest pore radius that we can achieve
		before energy minimization becomes problematic is 2.5 \AA. After being run
		through the full dry equilibration procedure, the average pore radius is 0.40
		$\pm$ 0.01 \AA.

		\item 5 \AA~: We tested a pore radius of 5 \AA~because it is
		slightly larger than the equilibrated pore radius of the system simulated with
		an initial pore radius of 2.5 \AA. After being run through the full dry
		equilibration procedure, the average pore radius is 0.42 $\pm$ 0.01 \AA, which
		agrees with the 2.5 \AA~configuration within uncertainty.

		\item 8 \AA~: The largest pore radius that can achieved before
		energy minimization becomes problematic is 8 \AA. One should use caution with
		such a structure because of the relatively large vacuum space that is created
		in the pore region of the initial configuration. After being run through the
		full dry equilibration procedure, we see a combination of cylindrical and
		slit-like pores (Figure~\ref{fig:slits}). Measuring the pore radius of this
		system does not have a concrete meaning since slits do not have a single
		radius, but its calculated value is reported in Table ~\ref{table:radii}. We
		are wary of such a non-symmetric structure and choose not to use a pore radius
		of 8 \AA~for our starting configurations.

	  \end{enumerate}

	  \begin{table}[h]
	  \centering
	  \begin{tabular}{cc}
	  \toprule
	  Initial Pore Radius & Equilibrated Pore Radius \\
	  \midrule
	  % BJC: to be updated once simulations are run out long enough
	  2.5 \AA & $0.40 \pm 0.01$ \AA \\
	  5 \AA   & $0.42 \pm 0.01$ \AA \\ % final result
	  8 \AA   & $0.69 \pm 0.01$ \AA \\
	  \bottomrule
	  \end{tabular}
	  \caption{The average pore radii of systems built with an initial pore
		  radius of 2.5 \AA~and 5 \AA~equilibrate to values that agree within error. If
		  the pore radius is too large, slit pores may form.}~\label{table:radii}
	  \end{table}

	  \begin{figure}[!htb]
	  \centering
          \includegraphics[width=\textwidth]{slit_pores.png}
	  \caption{A system that was built with an initial pore radius of 8
		  \AA~equilibrates to a structure that exhibits both cylindrical and slit-like
		  pores. As pictured here, sodium ions are colored blue, carbon atoms in the
		  aromatic ring of the head group are colored orange and all else is colored
		  cyan.}~\label{fig:slits}
	  \end{figure}

%	  \item \underline{Initial distance between layers}
	  \subsection{Initial distance between layers}\label{section:initial_dbwl}

	  We tested 3 different initial layer spacings, defined as the distance
	  between the planes of aromatic rings in each layer. Systems built with layers
	  stacked 3.7 \AA~and 5 \AA~apart are discussed extensively in the main text. We
	  also tested a system built with layers stacked 10 \AA~ apart.

	  Figure~\ref{fig:dbwl_10} shows the structure of an assembly built
	  with an initial layer spacing of 10 \AA~immediately after the restrained
	  portion of the equilibration procedure. Since we used position restraints, the
	  simulations were run in the NVT ensemble. When layer spacing is large, such as
	  this situation, there is a significant amount of vacuum space which the monomer
	  attempts to fill. Even if turning pressure control on allows the system to
	  recover the geometry of the hexagonal phase, we would likely need much longer
	  equilibration times, and it will almost certainly get trapped in a metastable
	  configuration that bears no resemblance to the experimental profile.

	  \begin{figure}[!htb]
		\centering
		\includegraphics[width=0.5\textwidth]{dbwl_10.png}
		\caption{When layers are initially stacked 10 \AA~apart and the system
                is equilibrated using the dry equilibration procedure, large vacuum gaps
		form as the monomers attempt to fill space.}\label{fig:dbwl_10} 
	  \end{figure}

%  \end{enumerate}

%  \noindent
%  \begingroup
%        \fontsize{14pt}{14pt}\selectfont
%	\textbf{Calculation of pore-to-pore spacing statistics}
%  \endgroup

  \begin{figure}[!htb]
	\centering
	\includegraphics[width=0.25\textwidth]{p2p_diagram.png}
	\caption{There are five pore-to-pore distances that are equal in a perfect
		hexagon. If each number in the diagram represents the center of a
		pore in a hexagonal unit cell, then the distance from 1 to 2, 2 to 4, 
		4 to 3, 3 to 1 and 1 to 4 should be equal. Only 4 of the distances
		are independent. For example, the distance from 1 to 4 is defined by
		the location of all other pore centers.}
	\label{fig:p2p_diagram}
  \end{figure}

  \section{Equilibration Details}\label{section:equilibration}

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We run a 50 ps simulation in
  the NVT ensemble which allows the monomer tails to settle without disrupting
  the ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^{-1}$ nm$^{-2}$, we reduce the restraints in a sequence with values of
  8, 3, 2, 1, and 0 KJ mol$^{-1}$ nm$^{-2}$ respectively. We allow the resulting
  unrestrained structure to equilibrate for 5 ns in the NPT ensemble
  with pressure controlled by the Berendsen barostat. Next, we run long NPT
  equilibration simulations for at least 400 ns using the Parrinello-Rahman
  barostat with a time constant of 10 ps.

  % MRS6: some parameters missing here (time run between insertions, etc). 
  % I assume they are in supporting information?
  % BJC5: all water is present in initial configuration. This did not follow the more
  % complex procedures that I've come up with since then. 
  In order to create a ``wet'' system, we solvated an initial configuration with
  water using \texttt{gmx solvate}. We remove all water molecules placed outside
  the pore region. Then we randomly remove water molecules inside the pore region
  until the pores reach the desired concentration of water. The remainder of the
  equilibration follows the same procedure as the dry system.

  \section{Calculation of pore-to-pore spacing statistics}\label{section:p2p_stats}

 %BJC: is it really necessary to explain this if I release the code? I don't explain
 %any of the other stats I calculate.
  \vspace{1em} 
  We are interested in 5 pore-to-pore distances which should all be
  equal in a perfect hexagonal array (Figure~\ref{fig:p2p_diagram}). Only 4 of
  the 5 distances are independent. We can calculate a trajectory of spacing
  versus time for each of the 5 distances. We calculated the average pore-to-pore
  spacing and its uncertainty according to the following procedure:
   \begin{enumerate}
	\item We calculated the time when each of the pore-to-pore distances
	were equilibrated using \\ \texttt{pymbar.timeseries.detectEquilibration()}
	\cite{chodera_simple_2016,shirts_statistically_2008}. We began calculations
	after the largest of the five values.
	\item We calculated how long it takes for the data in each of the 5
	trajectories to become uncorrelated using \texttt{pymbar.timeseries.integratedAutocorrelationTime()}
	\cite{chodera_simple_2016,shirts_statistically_2008}. 
	\item We broke the full equilibrated trajectory into blocks of length $\tau$, where
	$\tau$ is the maximum of the five autocorrelation times calculated. Each block contains
	five sub-trajectories of pore-to-pore spacings.   
	\item We generate statistics using the bootstrapping technique. For
	each bootstrap trial, we reconstruct an equilibrium trajectory by randomly
	sampling from the trajectory blocks. 
	\item The average pore-to-pore distance is the mean of all pore
	spacings among all bootstrap trials.
	\item To calculate the uncertainty in pore-to-pore distance, we calculate the average 
	pore spacing for each pore over all bootstrap trials. Using the 5 average 
	pore-to-pore distances, we calculate the spread with:
	\begin{equation}
	s = \sqrt{\frac{1}{4} \sum_{i=1}^5 (x_i - \overline{x})^2} 
	\label{equation:p2p_stats}
	\end{equation}
	where $\overline{x}$ is the average pore-to-pore distance. We divided the sum in 
	equation~\ref{equation:p2p_stats} by 4 since there are only 4 independent pore
	spacings.

  \end{enumerate}

  \section{The nematic order parameter}\label{method:nematic_order}

  We calculated the nematic order parameter for our system in order to
  understand the degree of ordering among monomer head groups. Typically, the
  nematic order parameter is calculated for nematic liquid crystal systems which
  are characterized by unidirectional ordering of liquid crystal monomers. The
  preferred direction of monomers is defined by the unit director vector,
  $\mathbf{n}$. Assuming a single preferred direction of alignment, the nematic
  order parameter, $S$, is defined as:
  \begin{equation}
	 S = \frac{1}{2} \langle(3\cos^2\theta -1)\rangle
	\label{eqn:nematic_order_parameter}
  \end{equation}
  where $\theta$ is the angle between the molecular long axis and $\mathbf{n}$.
  In a perfectly ordered nematic liquid crystal system, the molecular axis of each
  monomer should be aligned with $\mathbf{n}$ and give an order parameter of $S=1$. 
  We are interested in quantifying the degree of monomer head group alignment 
  between systems. We use Eq.~\ref{eqn:nematic_order_parameter} to accomplish this
  by defining $\mathbf{n}$ as the z-axis (or pore axis), and then measuring the angles,
  $\theta$, between $\mathbf{n}$ and the vectors perpendicular to the plane of
  the aromatic head groups ($\mathbf{v}$ in Figure~\ref{fig:director}).

  %BJC: change n to v in picture.
  \begin{figure}[!htb]
	\centering
	\includegraphics[width=0.75\linewidth]{nematic_director.png}
	\caption{We calculate the nematic order parameter by measuring the angle
	between $\mathbf{v}$ and the z-axis.}~\label{fig:director}
  \end{figure}

  The nematic order parameter is highest in sandwiched systems. 
  Figure~\ref{fig:nematic_distribution} shows the distribution of angles between
  the nematic director and the vector perpendicular to the monomer head groups. 
  Systems in the ordered basin have higher nematic order parameters than their
  disordered basin counterparts. Both sandwiched system have higher nematic order
  parameters than both parallel displaced systems. This may occur because monomers
  in the sandwiched configuration have their motion restricted by vertically adjacent
  monomers, while those in the parallel displaced configuration have more room to 
  rotate.

  \begin{figure}[!htb]

  \begin{subfigure}{\linewidth}
        \centering
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{layered_nematic_order.png}
                \caption{Sandwiched, Ordered}~\label{fig:sandwich_nematic}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{offset_nematic_order.png}
                \caption{Parallel Displaced, Ordered}~\label{fig:offset_nematic}
        \end{subfigure}
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_sandwich_nematic_order.png}
                \caption{Sandwiched, Disordered}~\label{fig:disorder_sandwich_nematic}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_offset_nematic_order.png}
                \caption{Parallel Displaced, Disordered}~\label{fig:disorder_offset_nematic}
        \end{subfigure}

  \end{subfigure}
  \caption{The distribution of angles between nematic director vector (See
	  Figure~\ref{fig:director}) and the z-axis averaged over the equilibrated portion
	  of each trajectory}~\label{fig:nematic_distribution}
  \end{figure}

  \section{Measurement of intensities of experimental and simulated XRD patterns}\label{section:xrd_intensities}

  We developed standardized methods of measuring the intensity of reflections of interest
  in the experimental and simulated X-ray diffraction patterns presented in this work. Here,
  we graphically illustrate the methods used for measuring these intensities.

  All diffraction patterns are normalized according to R-alkanes and the intensities
  of all reflections are reported relative to the intensity of R-alkanes. The region
  used to calculate the average intensity of R-alkanes is shown in 
  Figure~\ref{fig:ralkanes}. The top region is excluded because, in the simulated
  patterns, R-$\pi$ intersects with R-alkanes.

  R-$\pi$ and R-double are reported as the maximum values of the peaks of the 
  $q_z$ cross-section of the experimental pattern at $q_r=0$ (Figure~\ref{fig:rpi_rdouble}).
  In the case of the experimental pattern, the peak heights are not symmetrical,
  so we report the average of the two heights.

  We measured R-spots by calculating the average intensity within a region 
  selected by visual inspection. We report the average intensity within the regions 
  highlighted in red in Figure~\ref{fig:rspots}. If there are no easily discernible
  spots, we report the average intensity at the intersection of R-alkanes with
  the $q_z$ value of R-double since that is where we expect it to appear based on
  experiment. Since R-double does not appear in our simulated patterns, we 
  estimate where it should appear as half of the $q_z$ value of R-$\pi$.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{WAXS_raw.png}
  \caption{Unaltered experimental WAXS pattern}\label{fig:unaltered}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{ralkanes.png}
  \caption{R-alkanes}\label{fig:ralkanes}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{rpi_rdouble.png}
  \caption{$q_z|_{r=0}$}\label{fig:rpi_rdouble}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{rspots.png}
  \caption{R-spots}\label{fig:rspots}
  \end{subfigure}
  \caption{}\label{fig:xrd_intensities}
  \end{figure}

  \section{Noise in simulated diffraction patterns}\label{section:xrd_noise}

  The magnitude of noise in the simulated diffraction patterns decrease as 
  the number of independent configurations increases. We simulated the structure
  factor of 100 randomly placed particles and varied the number of independent 
  configurations (N). It is expected that the structure factor should resemble
  a two-dimensional Gaussian. We cannot sample enough independent configurations
  of our system to completely eliminate noise.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{xrd_1frame.png}
  \caption{N=1}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{xrd_10frame.png}
  \caption{N=10}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{xrd_100frame.png}
  \caption{N=100}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{xrd_10000frame.png}
  \caption{N=10000}
  \end{subfigure}
  \caption{The magnitude of noise, especially along $q_r$=0, decreases
  as the number of independent configurations (N) increases.}\label{fig:xrd_noise}
  \end{figure}

  \section{Experimental correlation length of R-$\pi$}\label{section:correlation_length}

  \vspace{1em}
  Experimentalists did not measure the correlation length, so we estimated it
  from the raw data. We took a slice of the 2D WAXS data along the dashed line in
  Figure ~\ref{fig:waxs_dashed}. We removed background noise by subtracting the
  intensity far from the pattern (high $q$) uniformly. We used a least squares
  algorithm to fit a Lorentzian curve to the data. Only values recorded above
  $q_z$ = 1.6 were considered for the fit in order to mitigate interference from
  R-alkanes (Figure~\ref{fig:correlation_length_exp}). The full width at half
  maximum (FWHM) is related to the correlation length, L, by the relationship: $
  L = \frac{1}{FWHM} $. The error in the value is calculated as the square root
  of the diagonal entry of the covariance matrix of optimized fit parameters.
  We estimate the value of the correlation length to be 10 $\pm$ 1 using this 
  method. Since the experimental peak width is also influenced by strain and 
  instrumental broadening in addition to finite size broadening, the value 
  calculated here is really only an upper bound on the correlation length. 
  % need citations and more on Lorentzian parameters
  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}	
  	\includegraphics[width=\textwidth]{waxs_dashed.png}
	\caption{}~\label{fig:waxs_dashed}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
	\includegraphics[width=\textwidth]{Correlation_length_exp.png}
	\caption{}~\label{fig:correlation_length_exp}
  \end{subfigure}
  \caption{(a) We estimated the system's experimental correlation length using a 
  1D slice of the 2D WAXS patterns along $q_r$=0. (b) We fit a Lorentzian curve 
  to the 1D slice. One of the fit parameters is the full width at half maximum (FWHM).
  The correlation length is 1/FWHM.}~\label{fig:correlation}
  \end{figure}
  \vspace{1em} \\

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.47\textwidth}
        \includegraphics[width=1\linewidth]{offset_density.png}
        \caption{Parallel Displaced, Ordered Pore Basin}
        \label{fig:offset_density}
  \end{subfigure}
  \begin{subfigure}{0.47\textwidth}
        \includegraphics[width=1\linewidth]{layered_density.png}
        \caption{Sandwiched, Ordered Pore Basin}
        \label{fig:layered_density}
  \end{subfigure}
  \begin{subfigure}{0.47\textwidth}
        \includegraphics[width=1\linewidth]{disordered_offset_density.png}
        \caption{Parallel Displaced, Disordered Pore Basin}
        \label{fig:disordered_offset_density}
  \end{subfigure}
  \begin{subfigure}{0.47\textwidth}
        \includegraphics[width=1\linewidth]{disordered_density.png}
        \caption{Sandwiched, Disordered Pore Basin}
        \label{fig:disorder_layered_density}
  \end{subfigure}
  \caption{In all cases, the component radial distribution functions exhibit a
	  composition gradient transitioning from the hydrophilic to the hydrophobic
	  regions. Systems with layers initially spaced 3.7 \AA~apart in the  parallel
	  displaced (a), and sandwiched (b) configurations both show the highest
	  concentrations of ions and head groups away from the pore center. Systems with
	  layers initially spaced 5 \AA~apart (c) and (d), both show the highest
	  concentrations of ions and head groups near the pore center, implying a more
	  uniform, disordered pore.}~\label{fig:densities}
  \end{figure}

% BJC: Unnecessary, supporting info already long
%  \begin{figure}
%	\centering
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{zdf_layered_4.png}
%                \caption{4 mon/layer, Sandwiched}\label{fig:zdf_layered_4}
%        \end{subfigure}
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{zdf_offset_4.png}
%                \caption{4 mon/layer, Parallel Displaced}\label{fig:zdf_layered_4}
%        \end{subfigure}
%	\begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{zdf_layered_6.png}
%                \caption{6 mon/layer, Sandwiched}\label{fig:zdf_layered_6}
%        \end{subfigure}
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{zdf_offset_6.png}
%                \caption{6 mon/layer, Parallel Displaced}\label{fig:zdf_layered_6}
%        \end{subfigure}
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{zdf_layered_7.png}
%                \caption{7 mon/layer, Sandwiched}\label{fig:zdf_layered_7}
%        \end{subfigure}
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{zdf_offset_7.png}
%                \caption{7 mon/layer, Parallel Displaced}\label{fig:zdf_layered_7}
%        \end{subfigure}
%	\begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{zdf_layered_8.png}
%                \caption{8 mon/layer, Sandwiched}\label{fig:zdf_layered_8}
%        \end{subfigure}
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{zdf_offset_8.png}
%                \caption{8 mon/layer, Parallel Displaced}\label{fig:zdf_layered_8}
%        \end{subfigure}
%	\caption{$g(z)$ for all other configurations built with layers stacked 3.7 \AA~apart}\label{fig:zdf}
%  \end{figure}

%BJC: unnecessary. Supporting info already long..and more is coming
%  \begin{figure}
%	\centering
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{rzplot_layered_4.png}
%                \caption{4 mon/layer, Sandwiched}\label{fig:rzplot_layered_4}
%        \end{subfigure}
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{rzplot_offset_4.png}
%                \caption{4 mon/layer, Parallel Displaced}\label{fig:rzplot_offset_4}
%        \end{subfigure}
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{rzplot_layered_6.png}
%                \caption{6 mon/layer, Sandwiched}\label{fig:rzplot_layered_6}
%        \end{subfigure}
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{rzplot_offset_6.png}
%                \caption{6 mon/layer, Parallel Displaced}\label{fig:rzplot_offset_6}
%        \end{subfigure}
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{rzplot_layered_7.png}
%                \caption{7 mon/layer, Sandwiched}\label{fig:rzplot_layered_7}
%        \end{subfigure}
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{rzplot_offset_7.png}
%                \caption{7 mon/layer, Parallel Displaced}\label{fig:rzplot_offset_7}
%        \end{subfigure}
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{rzplot_layered_8.png}
%                \caption{8 mon/layer, Sandwiched}\label{fig:rzplot_layered_8}
%        \end{subfigure}
%        \begin{subfigure}{0.40\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{rzplot_offset_8.png}
%                \caption{8 mon/layer, Parallel Displaced}\label{fig:rzplot_offset_8}
%        \end{subfigure}
%	\caption{Simulated XRD patterns for all other configurations built with
%		 layers stacked 3.7 \AA~apart}\label{fig:XRDsim}
%  \end{figure}

%  \begin{wrapfigure}{R}{0.4\textwidth}
%      \centering
%      \includegraphics[width=0.4\textwidth]{tilt.png}
%      \caption{The average angle between alkane chains and the xy plane is nearly zero degrees}\label{fig:tilt}     
%  \end{wrapfigure}

%  \begin{figure}
%	\centering
%	\begin{subfigure}{0.45\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{ps5layered.png}
%		\caption{}\label(fig:ps5layered}
%	\end{subfigure}
%	\begin{subfigure}
%		\centering
%		\includegraphics[width=\textwidth]{ps5offset.png}
%		\caption{}\label{fig:ps5offset}
%	\end{subfigure}
%  \end{figure}

  \begin{figure}[!htb]
        \centering
%        \begin{subfigure}{0.45\textwidth}
%                \centering
%                \hspace{-1cm}
%                \vspace{1cm}
%                \includegraphics[width=\textwidth,scale=2]{tails_topview.png}
%                \caption{}\label{fig:tails_topview}
%        \end{subfigure}
%        \begin{subfigure}{0.45\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{tails_rzplot.png}
%                \caption{}\label{fig:tails_rzplot}
%        \end{subfigure}
%        \begin{subfigure}{0.45\textwidth}
                \includegraphics[width=0.5\textwidth]{centroids_box.png}
		\caption{Monomer tails pack together hexagonally. The centroid
			of each tail is visualized as a blue sphere. The centroids are calculated based
			on the red atoms in Figure~\ref{fig:monomer_color_coded}. The red sphere
			highlights an example of an alkane tail centroid with its nearest neighbors
			(green spheres) surrounding it in a hexagonal pattern.}\label{fig:centroids}
%        \end{subfigure}
%        \begin{subfigure}{0.45\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{angles_traj_layered.png}
%                \caption{}\label{fig:angle_distribution}
%        \end{subfigure}
%        \caption{(a) The trajectory can be stripped of all atoms except carbon
%        atoms in monomer tails. (b) Simulated diffraction of the tail-only trajectory
%        still gives rise to R-spots. (c) Finding the center of mass and visualizing
%        their coordinates reveals the hexagonal-like packing of the tails. (d) The
%        distribution created by measuring the angle between each centroid (e.g. red
%        in (c)) and its nearest neighbors (e.g. green in (c)) with respect to the xy
%        plane has distinct spikes near 30\degree, which is consistent with the location
%        of R-spots}\label{fig:tail_packing}
  \end{figure}

  \begin{figure}[!htb]
	\centering
	\includegraphics[width=0.75\linewidth]{radial_distribution_annulus.png}
	\caption{Looking down onto the plane of membrane, this diagram
		illustrates how we calculated radial distribution
		functions. We binned the radial distance of all atoms in 
                chosen groups from pore centers. The pore centers are 
                defined as the average coordinates of sodium ions in each
                pore. The bins are defined by the annulus bounded by concentric
		circles centered at the pore centers, as shown. To normalize, we
		divide the count of atoms within the bin annulus by the volume of
                the annulus where the volume is the area of the annulus times the
                height of the membrane in the z-direction.}~\label{fig:rdf_diagram}
  \end{figure}

  \begin{table}[h]
  \centering
  \begin{tabular}{cc}
  \toprule
  System & Diffusion Constant ($m^2/s$) \\ 
  \midrule
  Sandwiched, Ordered & $1.4\mathrm{e}{-14}$ \\
  Parallel Displaced, Ordered & $2.0\mathrm{e}{-14}$ \\
  Sandwiched, Disordered & $2.0\mathrm{e}{-14}$ \\
  Parallel Displaced, Disordered & $3.5\mathrm{e}{-14}$ \\
  \bottomrule
  \end{tabular}
  \caption{The diffusion constants for all systems are on the order of $10^{-14} m^2/s$ 
  which is three orders of magnitude lower than literature values.}~\label{table:msd}
  \end{table}

  \begin{figure}[!htb]
  \centering
        \begin{subfigure}{0.45\textwidth}
                \includegraphics[width=\textwidth]{water_filled_pore.png}
                \caption{}\label{fig:water_filled_pores}
        \end{subfigure}
        \begin{subfigure}{0.45\textwidth}
                \includegraphics[width=\textwidth]{water_removed.png}
                \caption{}\label{fig:water_removed}
        \end{subfigure}
  \caption{(a) Pores built in the parallel displaced configuration with 5
	  monomers per layer are filled with 5 wt\% water. (b) The same system is
	  visualized with water molecules and sodium ions removed. Head groups vacate the
	  pore region leaving an aqueous solution of water and sodium ions.}\label{fig:water_pores}
  \end{figure}

  \section{Demonstrative Fourier transforms}\label{section:fourier}

  Here we use Fourier transforms to demonstrate some of the arguments made in 
  the main text. Interpretation of real and simulated X-ray diffraction patterns
  is non-trivial and requires careful thought. They are a function of the positions
  of all atoms in space, and suffer from the phase problem. Namely, there are multiple
  solutions that could explain a single diffraction pattern. Using simplified models, 
  we can hypothesize about the effect of various configurations on the simulated patterns.

  \subsection{Magnitude of R-$\pi$}\label{section:rpi_ft}

  In all cases, the magnitude of R-$\pi$ in our system is larger than that in experiment.
  Some of the discrepancy can be explained by defects and imperfect pore alignment. 
  However, there are significant effects resulting from the ordered states that we simulate.

  Persistent oscillations in the z-correlation function, $g(z)$, increase the intensity of R-$\pi$. 
  In Figure~\ref{fig:z_correlation}, we fit a decaying sinusoidal function 
  (equation~\ref{M-eqn:decaying_sinusoid} of the main text) to $g(z)$ calculated from the
  sandwiched configuration in the ordered basin. While the sinusoidal decay fit decays to one
  rather quickly, $g(z)$ continues to oscillate with nearly the same amplitude as its third peak. 
  In Figure~\ref{fig:theoretical_correlation_functions}, we overlay the same decaying sinusoidal
  fit with a function that continues to oscillate with the same amplitude as the third peak of
  the fit to $g(z)$. In Figure~\ref{fig:power_spectrum_comparison}, we plot the power spectrum
  generated from the discrete Fourier transform of the plots in Figure~\ref{fig:theoretical_correlation_functions}. 
  The maximum intensity of the power spectrum increases by nearly 60 \% when the correlation 
  function does not decay to one. This explains, in part, the large simulated value of the
  intensity of R-$\pi$.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{z_correlation_layered_sinefit.png}
  \caption{}\label{fig:z_correlation}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{theoretical_correlation_functions.png}
  \caption{}\label{fig:theoretical_correlation_functions}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{power_spectrum_comparison.png}
  \caption{}\label{fig:power_spectrum_comparison}
  \end{subfigure}
  \caption{(a) We fit a decaying sinusoidal function to the z-correlation function, $g(z)$, 
  calculated from the ordered basin sandwiched system. The oscillations in $g(z)$ continue
  to oscillate at nearly the same amplitude as its third peak. (b) We overlaid a plot of the
  fit to $g(z)$ with a similar function that continues to oscillate with the same amplitude
  as its third peak. (c) The power spectrum of each plot in (b) exhibits a 60 \% increase in
  intensity of the peaks when the function does not decay to 1.}\label{fig:rpi_ft}
  \end{figure}

  \subsection{Simplified assemblies}\label{section:simplified_assemblies}

  We explored the sandwiched and parallel displaced configurations as potential ways for
  aromatic head groups to stack. We used copies of the same monomer to build each system.
  We were unable to create any assemblies which exhibited R-double using this technique. 
  We simulated diffraction patterns of simplified configurations to understand why. We 
  placed points in array that resemble the sandwiched and parallel displaced configurations.
  Points are space 3.7 $\AA$ apart so it is easy to relate them to LLC membrane system.
  5 columns of points are placed 5 $\AA$ from a defined "pore center".

  When points are stacked in either configuration, R-double cannot appear. In the case
  of the sandwiched configuration, the fundamental frequency is set at 1.7 $\AA^{-1}$,
  the same location as R-$\pi$. Subharmonics appear at integer multiples of this frequency.
  The parallel displaced configuration exhibits reflections at half the $q_z$ value of
  R-$\pi$, which is where we expect to see R-double. However, just like the LLC membrane
  system, the reflections appear away from $q_r=0$ and are caused by the helical stacking
  arrangement.

  From these observations, it is clear that, in order to make R-double appear in our simulated
  diffraction patterns, we must create systems with asymmetries along the z-axis which lead to
  a fundamental frequency of $0.85 \AA^{-1}$. This requires a modulation in electron density 
  every 7.4 $\AA$. This line of thought gave rise to the systems studied in 
  section~\ref{M-section:rdouble} of the main text.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{simple_sandwich_realspace.png}
  \caption{}\label{fig:simple_sandwich_realspace}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{simple_sandwich_rzplot.png}
  \caption{}\label{fig:simple_sandwich_rzplot}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{simple_offset_realspace.png}
  \caption{}\label{fig:simple_offset_realspace}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{simple_offset_rzplot.png}
  \caption{}\label{fig:simple_offset_rzplot}
  \end{subfigure}
  \caption{}\label{fig:simple_FTs}
  \end{figure}

  \section{Solvation}\label{section:solvation}

  % formatting taken from : https://tex.stackexchange.com/questions/239715/add-titles-for-rows-and-columns-in-a-subfloat
  \newlength{\tempdima}
  \newcommand{\rowname}[1]% #1 = text
  {\rotatebox{90}{\makebox[\tempdima][c]{\textbf{#1}}}}

  \renewcommand{\thesubfigure}{\alph{subfigure}}
  \newcommand{\mycaption}[1]% #1 = caption
  {\refstepcounter{subfigure}\textbf{(\thesubfigure) }{\ignorespaces #1}}

  \begin{figure}[!htb]
  \begin{subfigure}{0.915\textwidth}
  	\settoheight{\tempdima}{\includegraphics[width=.32\linewidth]{solvated_offset_rzplot_1.png}}%
  	\centering\begin{tabular}{@{}c@{ }c@{ }c@{ }c@{}}
  	&\textbf{1 wt\%} & \textbf{\hspace{2em}2.5 wt\%} & \textbf{5 wt\%} \\
  	\rowname{Parallel Displaced}&
  	\includegraphics[width=.28\linewidth,trim={1cm 0 1.3cm 0},clip]{solvated_offset_rzplot_1.png}&
  	\includegraphics[width=.28\linewidth,trim={1cm 0 1.3cm 0},clip]{solvated_offset_rzplot_25.png}&
  	\includegraphics[width=.28\linewidth,trim={1cm 0 1.3cm 0},clip]{solvated_offset_rzplot_5.png}\\[-1ex]
  	%&\mycaption{0.2} & \mycaption{0.2} & \mycaption{0.3}\\
  	\rowname{Sandwiched}&
  	\includegraphics[width=.28\linewidth,trim={1cm 0 1.3cm 0},clip]{solvated_layered_rzplot_1.png}&
  	\includegraphics[width=.28\linewidth,trim={1cm 0 1.3cm 0},clip]{solvated_layered_rzplot_25.png}&
  	\includegraphics[width=.28\linewidth,trim={1cm 0 1.3cm 0},clip]{solvated_layered_rzplot_5.png}\\[-1ex]
  	%\mycaption{0.5} & \mycaption{0.6} & \mycaption{0.7} \\
  	\end{tabular}
  \end{subfigure}
  \begin{subfigure}{0.075\textwidth}
  	\includegraphics[width=\textwidth]{colorbar_jet.png}
  \end{subfigure}
	\caption{Adding 1 wt \% water to the ordered basin dry systems causes an increase in the intensity
	R-spots. R-double becomes visible in the parallel displaced configuration. Additional amounts
	of water increase discrepancies with experiment} %MRS7:by increasing disorder?
  \label{fig:solvation}

  \end{figure}

  \section{Crosslinking algorithm details}\label{section:xlink}

  Crosslinking of the the monomer Na-GA3C11 occurs by a UV initiated free
  radical polymerization (FRP) (Figure~\ref{fig:xlink_mech}). Head-to-tail addition
  takes place between terminal vinyl groups on each of the monomer tails. We
  only considered head-to-tail addition since it is the dominant propagation mode
  in the real system.

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{Crosslink_mechanism.eps}
  \caption{Terminal vinyl groups present on each monomer tail react with free
	  radical initiators to create monomers with terminal vinyl radicals.  Vinyl
	  radicals react with the vinyl groups of other monomers in to propagate
	  crosslinking.}\label{fig:xlink_mech}
  \end{figure}

  We based our cross-linking algorithm on the known reaction mechanism.
  FRPs require an initiator which bonds to the system, meaning new atoms are
  introduced into the system. For simplicity, we simulated the initiator as
  hydrogen and made it present in the simulation by including them as dummy atoms
  in all possible locations where an addition could occur. We carry out the
  cross-linking procedure iteratively. During each iteration, the algorithm
  selects eligible bonding carbon atoms based on a distance cut-off. The algorithm
  then updates the topology with new bonds and changes dummy hydrogen atoms to appropriate
  hydrogen atom types. We only considered head-to-tail addition
  due to its dominance in the real system \cite{young_introduction_2011}. We did
  not consider direction of attack because the resultant mixture is racemic.

  Our implementation requires long simulation times to achieve high cross-link 
  densities. A typical cross-linking procedure can take up to 24 hours. In
  order to collect equilibrated data, further NPT simulation is necessary. We
  typically run a cross-linked system for an additional 100 ns to allow the system
  to readjust. For those reasons we did not cross-link all systems tested.

  \section{Ionic Conductivity}\label{section:ionic_conductivity}

  In the main text, we calculated ionic conductivity using the Nernst Einstein 
  relationship. We also measured ionic conductivity using a second method, 
  termed the Collective Diffusion model, for robustness. The collective 
  diffusion model measures the movement of the collective variable, Q, which
  is defined as the amount of charge transfer through the system and can be 
  thought to represent the center of charge of the system. The conductance, 
  $\gamma$, of the system can be calculated as:
  \begin{equation}
	 \gamma = \dfrac{D_Q}{k_b T} 
	\label{eqn:collective_diffusion}
  \end{equation}
  $D_Q$ is the diffusion coefficient of the collective variable Q and is 
  calculated using the Einstein relation. We convert the resulting value
  to ionic conductivity by multiplying by channel length and dividing by
  the membrane cross sectional area. One can access a detailed derivation of 
  the model elsewhere\cite{liu_collective_2013}.

  \begin{figure}[!htb]
        \centering
        \includegraphics[width=0.5\linewidth]{Ionic_conductivity.png}
        \caption{The Collective Diffusion model and the Nernst-Einstein relation yield
        agreeing values of ionic conductivity for both types of ordered basin systems 
        simulated. Both methods estimate the value of the ionic conductivity to be an 
        order of magnitude higher than the experimental value. There is more noise in
        the collective diffusion model because there is inherently less data that 
        can be used for its calculation. For that reason, we use only the Nernst 
        Einstein relationship in the main text.}
        \label{fig:conductivity}
  \end{figure}

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.5\textwidth]{z_correlation_overlay.png}
  \caption{The correlation functions generated from ordered basin sandwiched configurations
  with different z-dimensions. The correlation function is nearly the same when the z-dimension
  of the system is doubled.}\label{fig:z_correlation_overlay}
  \end{figure}

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.5\textwidth]{carboxylate_dihedral_rb.png}
  \caption{A potential energy diagram for the dihedral angle about the carbon-carbon bond of
  a carboxylate group attached to an aromatic ring, using parameters from the generalized amber 
  forcefield. It is energetically favorable for the carboxylate group to stay in plane with 
  the aromatic ring. There is a large energy penalty as the dihedral becomes antiparallel to 
  the plane of the aromatic ring.}\label{fig:carboxylate_dihedral_rb}
  \end{figure}

%  \begin{figure}[!htb]
%  \centering
%  \begin{subfigure}{width=0.45\textwidth}
%  \includegraphics[width=\textwidth]{rdouble_offset_solvated_rzplot.png}
%  \caption{}\label{fig:rdouble_offset_solvated_rzplot}
%  \end{subfigure}
%  \begin{subfigure}{width=0.45\textwidth}
%  \includegraphics[width=\textwidth]{rzplot_pore1_removed.png}
%  \caption{}\label{fig:rzplot_pore1_removed}
%  \end{subfigure}
%  \caption{(a) The simulated diffraction pattern of a system simulated in an ordered basin 
%  parallel displaced configuration with 1 wt\% water shows R-double at $mathbf{q}=(0, 0.7) \AA^{-1}$.
%  (b) When we remove }\label{fig:pore1_removed}
%  \end{figure}

\clearpage
\bibliography{llc}
\end{document}

% LocalWords:  Na GA Packmol pdb filetype MarvinSketch noauthor nodate GROMACS
% LocalWords:  atomtypes antechamber inpcrd prmtop acpype py silva gro gmx NVT
% LocalWords:  editconf molcharge bccsym cyan png dihedrals pore's ca pt xy kJ
% LocalWords:  kinetically metastable outlier BJC Berendsen barostat Parrinello
% LocalWords:  Rahman solvated solvate Eq XRD alkanes WAXS exp FWHM mon alkane
% LocalWords:  transitioning topview rzplot traj binned wt LLC Subharmonics UV
% LocalWords:  realspace Crosslinking polymerization FRP crosslinking FRPs rb
% LocalWords:  racemic Nernst carboxylate antiparallel llc