### 2018-11-13 Ben_Manuscripts/Masters_Bypass/master.tex

\documentclass{article}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath} % or simply amstext
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\newcommand{\angstrom}{\textup{\AA}}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}

\title{Predicting Transport in Lyotropic Liquid Crystal Membranes with Molecular Dynamics Simulations}
\author{Benjamin J. Coscia \and Xunda Feng \and Joe Yelk \and Matthew A. Glaser \and Douglas L. Gin \and Richard D. Noble \and Michael R. Shirts}

\begin{document}

\bibliographystyle{ieeetr}
\graphicspath{{./figures/}}  % put all the figures here
\maketitle

\input{intro}
\input{methods}
\input{results}
\input{conclusion}

%\include{supp} supplementary info should be included as a separate document for ACSnano

\newpage
\bibliography{llc}
\end{document}