### 2018-11-18 Ben_Manuscripts/structure_paper/response_to_reviewers.tex
Added many responses to Reviewer 1. Outlined else needs to be done

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{xr}

\externaldocument[S-]{supporting_information}
\externaldocument[M-]{Final_Draft}

\title{Response to reviewers: Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes}
\author{Benjamin J. Coscia \and Joseph Yelk \and Matthew A. Glaser \and Douglas L. Gin \and Xunda Feng \and Michael R. Shirts}

\begin{document}

\maketitle

We thank the reviewers for carefully reading over our manuscript and providing helpful comments. We have taken the suggestions into consideration and made appropriate revisions to the manuscript document. All changes to the text have been documented below.

\section{Response to Reviewer 1}

\begin{enumerate}

    \item \begin{quote} \textit{This manuscript describes atomistic molecular
	    dynamics simulations and diffraction analysis of hexagonally-arranged pores,
	    each of which is made of many columns of the monomer Na-GA3C11. This physical
	    system is of interest because such nanopores have potential to transform
	    filtration applications. This work is of potential interest to the JPC
	    readership because it provides atomistic insight into the origins of
	    diffraction reflections observed in experiments, which could be used to resolve
	    discrepancies about what the structure of these hexagonal arrays are, or used
	    to engineer better filters. Overall, the manuscript provides comprehensive
	    descriptions of metrics used to evaluate the simulated structures against
	    experiments, but the number of things described and length of these
	    descriptions obscures whether significant new physical insight into this system
	    is made.} 
    \end{quote}

    Author reply: We thank the reviewer for their constructive feedback and aim to address all of their concerns.

    \item \begin{quote}
    \textit{As written, it is not clear if the authors are presenting this study as a validation of 
	GAFF and the present equilibration techniques, or whether this study is providing significant new
	physical insight into these lyotropic liquid crystals. The paper reads a bit more like the former,
	with a comprehensive discussion of the degree to which five scattering features expected in 
	experiments are predicted in simulations. However, at the same time, the authors make arguments
	including (1) pores are made from five columns of monomers, (2) that water is present in the "dry"
	structures, and (3) that a scattering feature comes from tail packing, not tilt: These observations
	*could* be new physical insight into lyotropic liquid crystal structure, but as presented seem more
	like hypotheses than conclusions:}

	%BJC: want to convery that we want to reproduce all of the experimental features as best as possible
	% using an out-of-the-box force field. That way we know we are approximating the chemical environment
	% as close to the experimental system as possible. The goal is to eventually test large numbers of 
	% monomers without the needing to dive into the excruciating detail shown here.

    \end{quote}

    \item \begin{quote}
    \textit{1. For the simulations performed here, the authors do make a strong case that 5 columns per
    core gets the best pore-to-pore spacing (Figure 5). However, insufficient time is accessed to see 
    whether the 4-, 5-, and 6-column initial conditions move towards or away from the expected pore 
    spacing over many-microseconds trajectories. That said, for the quenches performed here, 5-columns
    per core seems more plausible than 4.}
    \end{quote}

    Author reply: We agree with the reviewer that 5 columns-per-core is more plausible than 4 and that much
    longer simulations may provide further supporting evidence. However, there is very little motion on
    the hundreds of nanoseconds timescale that we simulated which suggests that we will not gain any 
    significant new insight by extending our simulations into the microsecond regime. It is not feasible 
    for us to carry out these simulations as it would consume our computational resources with little reward.
    %BJC: The following might apply:
    Some confusion may have been generated by agreement of the disordered basin 6 column-per-pore system
    with experimental pore spacing. As mentioned in the text, the vertical stacking distance between monomers
    in that system is far too large. The columns are `stretched' in the $z$-direction allowing the pores take
    up less space on the $xy$ plane and thus pack closer together.

    \item \begin{quote}    
    \textit{Is this significant physical insight? It is insight, but the
    authors do not make a compelling case for how 5 columns would change how we think about these crystals,
    or how it would impact membrane engineering (or more importantly, whether these 5-column structures
    assemble in the solution phase, or are only locally stable).}
    \end{quote}

    Author reply: We thank the reviewer for pushing us to provide more significant physical insight based
    on the number of monomers per column. The number of columns-per-pore influences the pore radius which
    is important for determining the limits of size-exclusion-based separations. It also tells use the 
    density of interacting sites within the pore which may play a role in small molecule transport. We have
    added the following text to the main document at the end of Section~\ref{M-section:mon_per_pore} on 
    page _:

    ``The number of columns per pore dictates the density of interaction sites within each pore and the 
    pore radius. A higher number of interaction sites will surely play a role in transport of molecules
    that have an affinity for the monomer head groups. 
    %We measured the pore radius as the average distance from the pore center to each head group COM. 
    The pore radius increases with the number of columns per pore (See Figure~\ref{S-fig:pore_radius}. 
    This radius is likely related to the size of the molecules which can be excluded by this type of 
    membrane. We may be able to control LLC membrane pore size by selecting monomers that are stable
    in configurations with more columns-per-pore."

    \item \begin{quote}
    \textit{2. The authors make a pretty compelling case that water causes the R-double scattering feature
    by showing its absence when water is added, but that the R-double appears and persists for 200ns after
    the water makes its way into the crystal. However, the authors also show that nonuniformly spaced
    monomers can give rise to the R-Double feature, a seeming counterexample to the necessity of water.
    Is the R-double feature with water because the monomers are now irregularly spaced? What happens when
    the waters are removed and the scattering re-performed in Fig 17c(bottom)?}
    \end{quote}

    Author reply: We thank the reviewer for exposing a potentially unclear
    explanation of the origin of R-double. R-double appears in a dry system with
    non-uniformly spaced monomers \textit{only} if $z$-direction position
    restraints are applied to monomer head groups in order to maintain unevenly
    spaced stacking. As soon as the restraints are released, R-double fades and the
    monomer spacing becomes uniform. We modified the text on page _ in order to more
    clearly convey this message:

    Old text: ``We can also produce R-double if the LLC monomers are not
    uniformly spaced in the $z$-direction, but instead form pairs that stack less
    than 3.7~\AA~apart, with COMs that are spaced 7.4 \AA~ from neighboring pairs
    of monomers (Figure~\ref{fig:staggered_rzplot_norestraints}).  Our force field
    causes our system to tend towards uniformly spaced layers. Simulations of
    unevenly spaced systems are only stable if position restraints are applied to
    heavy atoms of the phenyl rings. Additionally, there is little evidence from QM
    studies of stacked $\pi-\pi$ systems that such uneven stacking could be
    energetically stable.~\cite{tauer_beyond_2005}"

    New text: ``We can also produce R-double if the LLC monomers are not
    uniformly spaced in the $z$-direction, but instead are placed in pairs that
    stack less than 3.7~\AA~apart, with COMs that are spaced 7.4 \AA~ from
    neighboring pairs of monomers (Figure~\ref{fig:staggered_rzplot_norestraints}).
    Simulations of unevenly spaced systems are only stable while position
    restraints are applied to heavy atoms of the phenyl rings. There is little
    evidence from quantum mechanical studies of stacked $\pi-\pi$ systems that such
    uneven stacking could be energetically stable.~\cite{tauer_beyond_2005} As soon
    as we remove position restraints, our system to tends towards uniformly spaced
    monomers."

    R-double is only long-term stable without position restraints if a small
    amount of water is present. We believe that water facilitates the appearance
    of R-double by stabilizing non-uniform spacing through hydrogen bonding, as the
    reviewer suggests. We clarified the text on page _ in order to make this 
    conclusion clear:

    Old text: ``When two vertically stacked monomer head groups hydrogen bond
    with a shared water molecule, the monomers are drawn closer together (as
    illustrated in Figure~\ref{fig:hbond_visualization}), which creates an
    asymmetry that allows R-double to appear. If a monomer head group shares a
    hydrogen-bonded water molecule with a head group above itself, it will be less
    likely to share a water molecule with a head group below it due to geometric
    constraints. The monomer head group below can just as easily share a water
    molecule with a head group below itself. In this scenario, the COMs of each
    pair are 2 times the $\pi$-stacking distance apart which would lead to R-double
    (much like the configuration in
     Figure~\ref{fig:staggered_rzplot_norestraints}). There are a modest number of
    occurrences of this scenario, which we quantify in further detail in the SI,
    Section \ref{S-section:rdouble}."

    New text: ``The addition of a small amount of water to the pores stabilizes
    non-uniform stacking of head groups much like that shown in
    Figure~\ref{fig:staggered_rzplot_norestraints}. When two vertically stacked
    monomer head groups hydrogen bond with a shared water molecule, the monomers
    are drawn closer together (as illustrated in
    Figure~\ref{fig:hbond_visualization}), which creates the asymmetry that allows
    R-double to appear. If a monomer head group shares a hydrogen-bonded water
    molecule with a head group above itself, it will be less likely to share a
    water molecule with a head group below it due to geometric constraints. The
    monomer head group below can just as easily share a water molecule with a head
    group below itself. There are a modest number of occurrences of this scenario,
    which we quantify in further detail in the SI, Section
    \ref{S-section:rdouble}."

    The reviewer asked us to re-perform the scattering analysis on the same
    system with water molecules removed. We would like to emphasize that we
    attempted to make a similar argument by removing just the head group atoms from
    the trajectory and re-performing the scattering analysis
    (Figure~\ref{M-fig:rdouble_nophenyls}). In the absence of head groups, R-double
    is not present, implying that their structure (which is influenced by the
    presence of water molecules) is responsible for the R-double reflection. For
    completeness, we performed the scattering analysis on the same system with
    water molecules removed, as suggested by the reviewer, and R-double is still
    present. This is the expected result since the structure of water itself does
    not give rise to R-double. We've added the resulting simulated XRD pattern to
    the Supporting Information, Figure~\ref{S-TBD}.

    \item \begin{quote}
    \textit{3. The authors make the case that the R-spots feature comes from the structure of the hexagonal
    tails, not their tilt. The authors should perform the scattering analysis to check if the R-spots are
    present on the tilted-tails they constrain: This quick experiment would nail down their interpretation
    here, or show that the tilted tails are also plausible (but not observed in this model).}
    \end{quote}

    Author reply: We thank the reviewer for their suggestion. We do not deny
    that the tilt of the monomer tails is a plausible explanation for the
    appearance of reflections that resemble R-spots in many liquid crystal systems.
    While the reviewer's suggested experiment is a good idea, its results in this
    case are ambiguous for a couple reasons. First, the tails may still pack
    hexagonally in the tilted configuration which would independently give rise to
    R-spots. Second, it is necessary to impose high force constant (>1000 kJ mol
    nm$^{-2}$) position restraints in order to maintain a tilt angle close to that
    implied by the 2D WAXS data. The resulting high degree of order in the system 
    gives way to reflections which are easily misinterpretted. We have added the
    following text highlighted in red to the text shown on Page _ in order to
    acknowledge that tail tilt is a valid explanation of R-spot-like reflections 
    and to offer a reason a molecular reason why our monomers in our system do not
    tilt:

    Modified text: ``Previous literature has attributed the R-spots reflection
    in this particular WAXS dataset as the result of tilted alkane
    chains~\cite{feng_scalable_2014}. {\color{red} This explanation is not
    unfounded as there are examples in literature of systems where tilted liquid
    crystals give rise to reflections that resemble R-spots (Cite, Cite).} We
    measured the tilt angle of the alkane chains of the 280 K system by measuring
    the angle made by the vector extending from top to bottom of each tail with
    respect to the membrane plane.  We found that it equilibrates to an average
    tilt angle of -2 $\pm$ 13\degree~(Figure~\ref{fig:tilt}), far from the
    37\degree~tilt angle previously used to explain R-spots. Even when we placed
    position restraints on the monomers in a tilted initial configuration, monomers
    quickly reduced their tilt relative to the $xy$ plane to an angle statisically
    indistinguishable from zero (see Section \ref{S-section:tail_tilt} of the SI).
    {\color{red} While it is plausible that our force field may in part be
    responsible for the tendency of monomers to stay nearly flat, the monomer tails
    lack the functional groups which are known to facilitate tail tilting (cite).}"

    \item \begin{quote}
    \textit{These arguments would be stronger if there weren't so many
	    qualifications about simulation timescales and GAFF, in particular its ability
	    to predict the shortest-scale feature (pi-stacking) for this system. How does
	    everything else change if the force-field were tuned to get that feature
	    correct?}
    \end{quote}

    Author reply: We thank the reviewer for their constructive criticism. Many
    of our findings may be relatively independent of the force fields used, but we
    feel that we should include these qualifications in order to not overstate our
    case. It is common for researchers to run simulations for an insufficient
    length of time and to make bold claims. We acknowledge the shortcomings of our
    techniques in order to inspire other researchers to build on our initial work.

    We agree with the reviewer that we could tune our force field in order to
    exactly reproduce the $\pi-\pi$ stacking distance seen experimentally. Surely
    there will be some structural differences. There are many ways to tune a force
    field in order to achieve this. We prefer to work with an "out-of-the-box"
    force field since they are widely used and can be applied to vast number of
    systems. The ultimate goal of our work is to create new monomers that can
    achieve specfic separations.  If we can gain experimentally-consistent insights
    using our model simulated with GAFF, then we can have increased confidence in
    future predictions generated from systems built with unsynthesized monomers.
    With that said, we think that it would be interesting to repeat our analysis
    using different and/or optimized force fields in a future study in order to
    comment on the signficance of any differences.

    %BJC: Emphasize main goal in intro (or pull a quote that says it)
    %BJC: add justification of gaff to methods?
    %BJC: Add force field optimization to future work

    \item \begin{quote}
    \textit{The MD-Structure-Factor repository is not available at the provided
	    github link, nor does it appear to be online anywhere. So, the correctness of
	    that code cannot be checked.}
    \end{quote}

    Author reply: We thank the reviewer for pointing this out and apologize for
    our mistake. The referenced GitHub repository was set to private due to
    miscommunication with our collaborators but it has now been made public. Please
    feel free to read through the code. We have performed checks to ensure that the
    code works properly, especially when simulating monoclinic unit cells. For
    example, we generated test trajectories containing thousands of frames made up
    of configurations of randomly placed point scatterers in a monoclinic unit cell.
    We expect that the simulated diffraction pattern will have a gaussian shape.
    This exact result is shown in the Figure S-thingy of the supporting information.
    With improper treatment of the monoclinic geometry, the gaussian shape is
    skewed into an elliptical shape.

    \item \begin{quote}
    \textit{One might want to check that code based on Figure 7f, where
	    presumably spherical scatterers generate reflections with anisotropy (diamonds
	    instead of squares or circles). This kind of an artefact is plausible if there
	    are errors handling the non-cubic box, or with scattering angles that aren't
	    parallel to a box axis. llcsim/analysis/correlation.py looks like it has the
	    machinery for doing the scattering analysis.}
    \end{quote}

    Author reply: We thank the reviewer for their astute observation and
    the resulting improvements to our manuscript. We generated Figure 7f by
    calculating the structure factor of a 2-dimensional array of hexagonally-packed
    scatterers. It is meant to be a simplified representation of the scattering
    exhibited by hexagonally packed alkane tails so that we can qualitatively
    understand where we would expect R-spots to appear if tails indeed pack in that
    way. Since it is possible to generate a rectangular unit cell which contains
    hexagonally packed point scatterers, there are no artifacts due to improper
    treatment of periodicity.  Although we believe our calculations are correct, we
    originally used only eight total scatterers in the 2D unit cell and then
    plotted the structure factor with a contour plot
    (\texttt{matplotlib.pyplot.contourf}) which does not interpolate the data. We
    re-created Figure 7f by simulating the structure factor of 100 hexagonally
    packed scatterers, in order to get higher fourier space resolution, and then
    plotting the result with \texttt{matplotlib.pyplot.imshow} and Gaussian
    interpolation. We have added the code used to generate this figure to the
    \texttt{llcsim} repository which can be accessed at
    \texttt{https://github.com/bencoscia/llcsim/blob/master/analysis/hexagonal\_packing.py}.

    % Add code and clarify main text

    \item \begin{quote}	
    \textit{Regarding the ion conductivities, how do the numbers presented here
	    compare against ion conductivities predicted with GAFF and other polarizable
	    and nonpolarizable force fields? If GAFF is systematically too high in other
	    cases, it could help explain the order-of-magnitude discrepancy observed here.}
    \end{quote}

	\item \begin{quote}
    \textit{If the authors either (a) strengthen the cases for their proposed structural insights being  highly 
    likely (not just plausible), or (b) make a stronger case for why these particular insights have important 
    ramifications for our understanding of LLC's (their significance), then it would be easier to publish this 
    immediately. This reviewer's opinion is that this work would find a broader interested audience with a more
    concise focus on these physical insights, which could be accomplished by moving more content 
    (e.g., dihedral correlations) to the SI.}
    \end{quote}

\end{enumerate}

\section{Response to Reviewer 2}

\begin{enumerate}

	\item \begin{quote} \textit{The authors report a computer simulation study of the detailed structure
	of several systems, which are similar to the cross-linked inverted hexagonal phase of self-assembled
	lyotropic liquid crystals (LLCs) used to develop porous polymer membranes for separations. Specifically,
	the authors studied the Col-h thermotropic (solvent-free) phase formed by the assembly of the Na-GA3C11
	monomer, which is similar to the LLC phases used in the development of polymeric membranes. The Col-h 
	phase has been characterized experimentally using wide-angle X-ray scattering (WAXS), but nonetheless 
	important gaps in its structure still remains. Using all-atom classical molecular dynamics simulations,
	the authors performed a very careful and very detailed study of the structure of the Col-h phase formed
	by Na-GA3C11. The authors compared their simulated X-ray diffraction (XRD) patterns with existing 2D-WAXS
	experimental data, carefully discussed and explained the observed differences between simulations and 
	experiments, and provided molecular-level details of the structure of this system that cannot be obtained
	from experiments. This study is very detailed; all relevant information is provided for interested readers
	to replicate results from this study (including links to Python scripts used to set up the simulations, 
	analyze MD trajectories, and obtain simulated XRD patterns). The simulation results has been analyzed and
	discussed comprehensively (perhaps too comprehensively, as this paper is 63 pages long, has 21 figures and
	its Supporting Information file has 19 additional pages; however, I don’t think the authors should make any
	efforts to shorten the paper, as I feel every presented component is required to fully understand the 
	contributions of this study) Overall this is an extremely solid manuscript}\end{quote}

	Author reply: We thank the reviewer for the kind assessment of our work. We also acknowledge that this is a long article

	\item \begin{quote}

	\textit{I have no suggestions to further improve this paper, other than fixing a few small typos:}

    \textit{(1) Page 10, first paragraph: remove the ‘page 58’ associated with reference 25}

    \textit{(2) Page 54: reference to the paper of Feng et al is missing}

	\end{quote}

    Author reply: We thank the reviewer for pointing out these minor errors in our main text. These issues have been corrected in
    the revised main text file. % need to add page numbers once finished

    \item \begin{quote} \textit{(3) Page 54: mention examples of force fields that explicitly include pi-pi interactions
    } \end{quote}

    Author reply: We have adjusted our language in order to specifically mention and cite polarizable force fields as follows:

    ``Polarizable force fields explicitly including $\pi-\pi$ interactions may be able to draw stacked
    monomers closer together"

\end{enumerate}

\end{document}