### 2019-05-08 Ben_Manuscripts/transport/cover_letter.tex
updated tex.

\documentclass[fontsize=11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{amsmath} 
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{hyperref}
\usepackage[export]{adjustbox}
\usepackage{anyfontsize}

\begin{document}

	\graphicspath{{./figures/}}

	%BJC: check address, phone number and fax (?) number
	\begin{figure}
	\centering
	\begin{minipage}{0.37\textwidth}
	\includegraphics[width=2.14in,left]{CUBoulder.pdf}
	\end{minipage}
	\begin{minipage}{0.62\textwidth}
	\scriptsize
	\noindent Department of Chemical and Biological Engineering \hfill t 303-735-7860~~~~~~~~~~~~~~~~~~ \\
	\noindent 596 UCB \hfill f 303-492-8425~~~~~~~~~~~~~~~~~~ \\
	\noindent Boulder, Colorado 80309 \hfill michael.shirts@colorado.edu \\
	\end{minipage}
	\end{figure}

	\noindent May 3\textsuperscript{rd}, 2019\\

	\noindent Dear Editors,\\

	\newcommand{\ManuscriptTitle}{Chemically Selective Transport in a Cross-linked 
	H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane}

	We are submitting ``\ManuscriptTitle'' by Benjamin J. Coscia and Michael R. Shirts, for
	consideration for publication as an Article in the Journal of Physical Chemistry
	B. We expect this work to be of significant interest to researchers who study 
	nanostructured polymer membranes for aqueous separations.

	In this article, we use molecular dynamics (MD) simulations in
        order to observe transport of small solutes in an inverted
        hexagonal (H\textsubscript{II}) phase self-assembled lytropic
        liquid crystal (LLC) polymer membrane. This is one of the
        first, and far the most extensive, uses of molecular dynamics
        to understand the molecular details of transport in such
        chemical complex nanostructured channels. Most importantly, we
        show how such channels can give rise to selective transport
        that is not simply size dependent, but also varies according
        to chemical functionality of the solutes.

        Recently, we characterized an atomistic molecular model of an
        LLC membrane that is maximally consistent with experimental
        observations (Coscia et al. \textit{J. Phys. Chem. B}, 123,
        289--309 (2019)). In this work, we observe and quantify
        transport of water, sodium and 20 small polar solutes within
        the pores of our atomistic H\textsubscript{II} phase LLC
        membrane model. In general, all solutes exhibit subdiffusive
        transport behavior characterized by intermittent hops between
        long periods of entrapment. However, due the pores'
        inhomogeneous architecture, we observed 3 different mechanisms
        of entrapment. First, solutes can diffuse out of the pores and
        become entangled between tails. Second, many of the solutes
        are capable of donating hydrogen bonds to monomer head groups
        for extended periods of time. Finally, solutes can associate
        with monomer head group counter-ions within nanopores. Each
        solute is influenced by each mechanism to varying degrees
        dependent on its size and chemical functionality. We begin our
        discussion with a broad description of these mechanisms in the
        context of all solutes before restricting the discussion to
        more detailed analyses of subsets of chemically-similar
        solutes.

	%BJC: included phone numbers because it asked to do so in author's guide.
	\noindent Some suggestions for reviewers are:
	\begin{enumerate}

		\item Menachem Elimelech is a leader in the development of membrane technologies
		for	water desalination and water reuse (Yale University, 203-432-2789,
		\href{mailto:menachem.elimelech@yale.edu}{menachem.elimelech@yale.edu}).

		\item Francisco Hung has used molecular simulations to study nematic liquid crystals
		and various	nanostructured materials (Northeastern University, 617-373-8619 	
		\href{mailto:f.hung@northeastern.edu}{f.hung@northeastern.edu})

		\item Eric Jankowski has experience modeling soft materials using molecular dynamics
		including work done simulating structure factors from simulations, much like
		the techniques we use in our work (Boise State University, 208-426-5681,
		\href{mailto:ericjankowski@boisestate.edu}{ericjankowski@boisestate.edu})

	\end{enumerate}

	\noindent Please send correspondence regarding this paper to Michael R. Shirts (contact
	details in letterhead).\\

	\noindent Sincerely,

	\noindent Michael R. Shirts \\
	\noindent Benjamin J. Coscia \\

\end{document}

% LocalWords:  BJC