### 2019-05-13 Ben_Manuscripts/comps/Draft.tex
Added draft

\documentclass{article}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath} % or simply amstext
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}
\usepackage[x11names,table]{xcolor}
\usepackage{setspace}
\usepackage[font=small]{caption}
\setstretch{1.5} % line spacing
\captionsetup[table]{font={stretch=1,small}} % line spacing in tables
\captionsetup[figure]{font={stretch=1,small}} % line spacing in figures
\usepackage{url}
% For timeline
\newcommand{\foo}{\color{LightSteelBlue3}\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}}
\DeclareCaptionFont{blue}{\color{LightSteelBlue3}}
\usepackage{array,booktabs}
\usepackage[superscript,biblabel]{cite}  % make citations superscripts to save space
%\setlength{\belowcaptionskip}{-1.2\baselineskip}

\title{PhD Research Proposal: Molecular Level Design of Nanoporous Lyotropic Liquid
Crystal Membranes for Aqueous Separations \\ \vspace{0.5cm}
\large Advisors: Michael Shirts and Richard Noble}

%MRS: overall, really nice start.  You can cut back a little on the detail and emphasize a bit more ``This is what the data means and what the consequnces are.  Need to flesh out objective 3 a more (and objective 4 a bit more). 
\author{Benjamin J. Coscia}

\begin{document}

  \graphicspath{{./figures/}}

  \maketitle
  \thispagestyle{empty}
  \clearpage
  \setcounter{page}{1} % don't count title page as a page

  \section{State of the Art}\label{section:state-of-the-art}

  \subsection*{Commercial Membranes for Small Molecule Separations}

  % perhaps a word or two about micro and ultra filtration

  %MRS: probably can do a little less introduction of the overall membrane needs, as we want to make sure 
  %MRS: the focus is on the research (thesis, you can include everything).  Can probably hold off that until you get the rest done, then trim as needed for space.
  %
  %MRS: remember the purpose of the introduction to any document is to lay out the rules of engagement; you want 
  %MRS: to raise the questions that the rest of the document will answer.
  % 
  %MRS: try to make sure the application motivation and the research described match up when possible.
  %MRS: if particular applications wouldn't be directly enabled by the research, at least draw to the connection
  %MRS: i.e. talk about the general tools of computation to understand nanostructured membranes, even if the current
  %MRS: membranes under development aren't suitable for the particular separations described.
  %BJC: Something like: The computational techniques developed here are suitable not just for LLC Membranes, but for any nanostructured 
  % membrane material.

  More highly selective nanoporous membranes would be useful for 
  performing complex aqueous separations with seawater and various
  types of wastewater. For example, sodium chloride and boron in seawater 
  \cite{fritzmann_state---art_2007} and organic micropollutants found in
  municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006}
  represent a small subset of the diverse contaminants of water sources. 
  By efficiently separating contaminants from feed solutions with
  highly selective membranes, it is possible to reduce the number of 
  required membrane passes and post-treatment steps needed for a given 
  filtration process \cite{werber_materials_2016}, thus lowering energy 
  requirements and cost. Additionally, one can extract valuable 
  resources from the feed streams. For example, flowback water produced 
  during hydraulic fracturing of shale formations contains dissolved 
  species like acetate whose extraction has economic value \cite{dischinger_application_2017}.

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing commercial membrane
  filtration processes that can be used to separate solutes on the order of
  1 nm in size and smaller, including ions. Both apply hydraulic pressure 
  to the feed solution in order to overcome osmotic pressure and force water
  and unfiltered components through the membrane. RO membranes are typically
  thin film composite with a porous mechanical support layer and a thin but
  dense polymer matrix active layer where separations occur.\cite{jeong_interfacial_2007}
  RO separates solutes based on the solute's ability to dissolve into and 
  diffuse through the tortuous pathways available in the membrane's dense 
  active layer. RO offers high selectivity at the cost of relatively high 
  energy requirements since one must apply hydraulic pressure on the order of
  50--100 bar to the feed solution in order to achieve an economical flux~\cite{van_der_bruggen_review_2003}.
  In contrast to RO membranes, NF membranes have explicit pores on the order
  of 1 nm in size. Typically, separations are achieved based on size exclusion
  and Donnan exclusion if the membrane's surface has a net charge.~\cite{donnan_theory_1995}
  NF membranes require significantly less applied pressure in order to achieve solute flux
  comparable to RO. Unfortunately, conventional NF membrane synthesis processes, such as 
  phase-inversion\cite{smolders_microstructures_1992}, are stochastic in
  nature which yields pores that are polydisperse in size.\cite{werber_materials_2016}
  Pore size polydispersity is detrimental to membrane selectivity.

  The downfall of RO and NF membranes can be summarized by the well-known
  permeability-selectivity tradeoff. Namely, it is difficult to increase the
  permeability of a desired molecular or atomic species, while maintaining
  the same retention of an undesired species.\cite{werber_materials_2016}

  \subsection*{Nanostructured Membranes}

  Nanostructured membranes attempt to overcome the permeability-selectivity 
  tradeoff through intelligent design at the molecular level. Graphene sheets, 
  carbon nanotubes (CNTs) and zeolites are three highly studied nanostructured
  technologies.

  Ultrathin-film graphene and graphene oxide membranes are a very active
  area of research because they offer potential for extremely high 
  permeability membranes. An ideal graphene membrane is a 2D material 
  composed of a single layer of graphene.\cite{humplik_nanostructured_2011}
  Synthesis of a single layer of graphene at a large scale without introducing
  defects is a challenge that has not yet been overcome. Multilayered graphene
  membranes are potentially easier to synthesize and more scalable which has 
  resulted in an increasing interest over recent years.\cite{cohen-tanugi_multilayer_2016,wei_multilayered_2018}

  Carbon nanotubes (CNTs) have shown promise as aqueous separations membranes
  due to unprecedentedly fast water transport.\cite{humplik_nanostructured_2011,hummer_water_2001} 
  Practically, dispersing CNTs into a polymer matrix is extremely difficult
  because they tend to agglomerate due to Van der Waals forces. Functionalization
  of the CNT walls has been heavily investigated  as a way to overcome this issue, however 
  alignment of the CNTs into an array feasible as a nanoporous membrane 
  persists as a challenge to CNT membranes.\cite{sahoo_polymer_2010}

  Zeolite-coated ceramic membranes offer the potential for permeabilities
  comparable to ultrafiltration, molecular sieves with pores up to 100 nm in 
  size~\cite{werber_materials_2016}, with selectivities as good as NF and RO. 
  Zeolites have highly uniform nm-sized crystalline structures with cage-like
  cavities that allow movement and trapping of small solutes. The crystalline
  frameworks are typically formed by networks of silicon and aluminum each 
  attached to 4 oxygen atoms in a tetrahedral arrangement.~\cite{auerbach_handbook_2003}
  One can replace the silicon and aluminum atoms via ion exchange in order to
  control the size of the cavities and hence its molecular-seiving properties.~\cite{li_novel_2007}
  A number of studies have tested the permeability and sodium salt rejection
  of various zeolite membranes, however none have fully overcome the 
  permeability-selectivity tradeoff.~\cite{pendergast_review_2011}
%    \item Most are prone to defects in the crystalline structure.

  \subsection*{Lyotropic Liquid Crystal Membranes}

  Preliminary evidence has shown that cross-linked lyotropic liquid crystal
  (LLC) membranes can be produced at moderate scale and may be capable of 
  performing highly selective separations. LLCs are amphiphilic molecules 
  that have the ability to self-assemble into porous nanostructures~\cite{smith_ordered_1997}
  that can be cross-linked to create mechanically strong membrane films with 
  periodic uniform-sized pores on the order of 1 nm in diameter~\cite{zhou_supported_2005}. 
  Since LLC polymer membranes lack an appreciable pore size distribution, they 
  inherently exhibit high selectivity due to their molecular weight cut-off 
  (MWCO)~\cite{zhou_supported_2005}. Additionally, LLC monomers can be salts,
  and therefore lead to Donnan exclusion of ions at the membrane-feed solution 
  interface.\cite{donnan_theory_1995}
%    The membrane gains a net surface charge when counterions from
%  the head groups that line the pore walls escape into the feed solution in an
%  effort to balance the gradients of concentration and electric potential
%  \cite{donnan_theory_1995}.

  % BJC: should show figure with HII and QI monomer structures
  The feasibility of nanostructured LLC polymer membranes for selective separations
  has been demonstrated using LLC monomers that form the type 1 bicontinuous cubic
  (Q\textsubscript{I})~\cite{hatakeyama_water_2011,hatakeyama_nanoporous_2010,carter_glycerol-based_2012}
  and the inverted hexagonal (H\textsubscript{II}) 
  phases (see Figure~\ref{fig:structures}).~\cite{zhou_supported_2005} When separating organic solutes from NaCl, Q\textsubscript{I}-phase
  membrane filtration experiments have shown selectivity 2--3 times higher than
  commercial RO and 6--12 times higher than commercial NF membranes.~\cite{dischinger_application_2017}
  When separating a series of various sized dyes, an H\textsubscript{II}-phase 
  LLC membrane showed complete rejection of dyes bigger than 1.2 nm in size~\cite{zhou_supported_2005}.

%  \begin{wrapfigure}{L}{0.5\textwidth}
%  \centering
%  \includegraphics[width=\linewidth]{NaGA3C11_colorcoded.eps}
%  \includegraphics[width=0.8\linewidth]{q1_monomer.png}
%  \end{wrapfigure}

%  \begin{wrapfigure}{L}{0.75\textwidth}
  \begin{figure}
  \centering
  \begin{subfigure}{0.55\linewidth}
  \hspace{0.5cm}
  \includegraphics[width=0.8\linewidth]{NaGA3C11_colorcoded.eps}
  \vspace{1.4cm}
  \caption{}\label{fig:HII_mon}
  \end{subfigure}
  \begin{subfigure}{0.4\linewidth}
  \hspace{0.4cm}
  \includegraphics[width=0.7\linewidth]{q1_monomer.png}
  \caption{}\label{fig:q1_mon}
  \end{subfigure}
  \caption{H\textsubscript{II} and Q\textsubscript{I} phase LLC membranes have 
  been synthesized from monomers in (a) and (b) respectively. One can 
  alter membrane properties by modifying the length and number of tails (red), the
  functional head group (green) and the counterion identity (blue) of LLC monomers.}\label{fig:structures}
  \end{figure}

  Q\textsubscript{I}-phase membranes consist of a tortuous network of three
  dimensionally interconnected pores that prevent optimal through-plane
  transport. In contrast, the densely packed, non-tortuous and uniform sized
  pores of H\textsubscript{II}-phase membranes represent the ideal geometry
  for achieving high solute flux\cite{matyka_tortuosity-porosity_2008}. However,
  the hexagonally packed LC domains of the H\textsubscript{II}-phase are generally
  unaligned, which hurts membrane permeability. This domain scale misalignment
  had inhibited further development of this technology, and research efforts were
  focused on the Q\textsubscript{I} phase, whose geometry does not require 
  alignment~\cite{zhou_new_2007}.

%  The H\textsubscript{II}-phase pore geometry (Figure~\ref{fig:assembly}) has a
%  high theoretical capacity for transport than the Q\textsubscript{I} phase.
%  \begin{itemize}
%	  \item The H\textsubscript{II} phase forms at room temperature in the 
%	  presence of ca.~10 wt\% water and consists of hexagonally packed, 
%	  hydrophilic pore columns\cite{smith_ordered_1997}. 
%%	  \item The hexagonal, straight-pore geometry provides a dense array
%%	  of hydrophilic channels and short pathways for solutes to travel.
%	  % BJC: should rethink the following since we think that thermotropic doesn't exist
%	  \item In the absence of water, neat monomer will form the same hexagonal
%	  columnar structure which, in the literature, has been referred to as the
%	  Col\textsubscript{h} thermotropic phase\cite{feng_scalable_2014}.
%  \end{itemize}

  Recently, researchers learned how to macroscopically align the hexagonal 
  domains which has revived research into H\textsubscript{II}-phase LLC
  polymer membranes. In 2014, Feng et al.~showed that one can align Col\textsubscript{h}
  domains, a temperature-dependent hexagonal phase created by neat LLC monomers,
  using a magnetic field with subsequent cross-linking to lock the structure 
  in place\cite{feng_scalable_2014}. In 2016, Feng et al.~showed that one could 
  also obtain the same result by confining the neat monomer between PDMS or glass 
  substrates since hexagonal mesophases preferentially anchor perpendicular to 
  both surfaces\cite{feng_thin_2016}.

  Unfortunately, reproducing the work of Feng et al. with the hydrated 
  H\textsubscript{II} phase has been an experimental challenge. Therefore, the
  primary focus of experimental research efforts has been with the Q\textsubscript{I} 
  phase.

  \section{Project Objectives}\label{section:objectives}

  Our current understanding of the molecular details of LLC membranes'
  nanostructure is not sufficient to be able to precisely design them for
  specific separations. Dischinger et al.~attempted to use an empirical model
  that correlates the physicochemical properties of the counterion used in
  a Q\textsubscript{I}-phase LLC membrane to solute rejection. Although there
  was some agreement with their empirical model, it does not offer a 
  sufficiently detailed explanation of the critical molecular interactions 
  leading to the observed behavior.~\cite{dischinger_effect_2017} 
  H\textsubscript{II}-phase LLC polymer membrane studies have been limited
  primarily to the Na-GA3C11 monomer (Figure~\ref{fig:HII_mon}) with
  some characterization done after minor monomer structural modifications. 
  These studies emphasized large scale structural features such as pore spacing
  as well as size-based rejection.~\cite{zhou_supported_2005,resel_structural_2000}
  Given the similar topology of H\textsubscript{II} and Q\textsubscript{I}
  membrane pores, we expect the H\textsubscript{II} phase will also be 
  subject to complex solute-membrane interactions that are not easily explained
  by an empirical model.

  A molecular-level understanding of structure and transport in LLC polymer 
  membranes, enabled by molecular dynamics (MD) simulations, can provide 
  guidelines to reduce the large chemical space available to design
  monomers for creation of separation-specific membranes. Using a sufficiently
  accurate molecular model, we can observe transport of solutes within LLC 
  membrane nanopores with atomistic resolution and infer mechanisms. 
  Based on this information we will have a much greater capability to 
  intelligently design new membranes by screening new liquid crystal monomer
  designs with MD simulations. The principles learned can be implemented 
  and tested experimentally.

  \noindent There are four primary objectives of this PhD research.
  \begin{enumerate}

    \item Develop techniques to build and understand the nanoscopic structure
    of LLC membranes.

    %MRS: probably should talk about this 1) with proper tense (some of these things have been done) and 2) talk about what YOU are doing, not we. I helped design the research, but you carried essentially all of it out! If phrased in future tense, committee will be confused, thinking those are things still to be done.

    A useful molecular-level model should incorporate a detailed picture 
    of the nanoscopic pore structure, which is crucial to understanding
    the role of monomer structure in solute transport and membrane design.
    I have created a maximally consistent structural model by comparing 
    simulated X-ray diffraction (XRD) patterns, generated from MD trajectories,
    to an experimental 2D wide angle X-ray scattering (WAXS) spectrum of a 
    Col\textsubscript{h} phase LLC membrane. I have used this model as an 
    example in order to understand the ease with which we can apply our approach
    to alternate monomers and other LC phases.

%    shown that we achieve a
%    similar pore composition regardless of initial configuration which is 
%    influenced by a number of variables. This implies that our model can
%    
%        We will assess the extent to which
%    we can apply our understanding to the H\textsubscript{II} phase, as well
%    as systems built with alternate monomers.

    \item Determine dominant solute-membrane interactions that give rise to
    transport mechanisms.

    I observed the transport of a relatively large set of small polar solutes
    placed within the H\textsubscript{II} phase membrane nanopores. I analyzed
    the time series of each solute's position in addition to directly measuring
    the physical interactions, such as hydrogen bonding and ion coordination, 
    between solutes and LLC monomers.

    \item Create a stochastic model which can project long timescale 
    transport behavior.

    I will combine my qualitative knowledge of the solute transport mechanisms
    with simulation data in order to inform a stochastic model. This model
    should closely reproduce the properties of the time series that we observe 
    in our simulations. Due to the low computational cost of a stochastic model
    relative to MD simulations, I will be able to forecast long timescale transport
    behavior and make well-converged predictions of macroscopic transport properties.

    \item Adapt the same analysis to the Q\textsubscript{I} phase.

    Over the course of this project, experimental research surrounding
    LLC membranes has shifted nearly all focus towards the Q\textsubscript{I}
    phase due to its more facile synthesis.	Although most of my work has 
    been applied to the H\textsubscript{II} phase, much of the same analyses
    can be applied to the Q\textsubscript{I} phase. The biggest challenge will
    be adapting my techniques to its more complex three dimensional geometry.

%    \item Enable easy continuation of our work with a dedicated and 
%    well-documented python package.
%    
%	Although molecular simulations have become popular
%	for studying systems at the atomic level, LLCs used in this context
%	have not been heavily investigated. Consequently, much of the 
%	analysis developed for this project is not widely applied. Therefore, 
%	it is important for us to make available the scripts that reproduce
%	the exact results presented in our published papers along with detailed
%	documentation of the scripts. This will ensure near-seamless continuation
%	of this project and accelerate the development of LLC membranes.

  \end{enumerate}

%MRS: overall, good description of wha has been done.  You can cut back a little if space is an issue.  Give the paper citations with the points. 
  \section{Progress to Date}\label{section:progress}

  \textbf{\large Objective 1:} {\large Build and understand the nanoscopic structure
  of the H\textsubscript{II} phase} (\textcolor{green!40!olive}{\textbf{Complete}})

  My first task was to develop a procedure for building and equilibrating 
  an atomistic LLC membrane system (see Figure~\ref{fig:assembly}). I chose to build
  a monoclinic unit cell consisting of four pores. Each pore is composed of monomer
  columns, where each column consists of 20 monomers stacked on top of each other
  so that the phenyl groups are coplanar with each other and the $xy$ plane. The
  columns are oriented so that the hydrophilic monomer head groups face towards
  the pore center. I did not add any water to the initial configuration because I
  compared the structure to experimental data for a system claimed to be
  synthesized with only neat monomer. I equilibrated unit cells using a series
  of simulations with head groups held in place by position restraints, gradually
  reducing the force constant of the position restraints until the system
  was completely unrestrained. I let the system equilibrate for 400 ns of simulation
  time.

  \begin{figure}[!htb]
      \centering
      \begin{subfigure}{.325\textwidth}
              \centering
              \vspace{0.5cm}
              \includegraphics[width=\textwidth]{NaGA3C11_wedge.eps}
              \vspace{0.05cm}
              \caption{}~\label{fig:monomer}
      \end{subfigure}
%      \begin{subfigure}{.3\textwidth}
%              \centering
%              \includegraphics[width=\textwidth]{monomer_twocolor.pdf}
%              \caption{}~\label{fig:atomistic_monomer}
%      \end{subfigure}
%      \begin{subfigure}{0.3\linewidth}
%              \centering
%              \includegraphics[width=0.6\textwidth]{wedge_thick.pdf}
%              \caption{}~\label{fig:wedge}
%      \end{subfigure}
              \begin{subfigure}{0.325\linewidth}
              \centering
              \includegraphics[width=0.6\textwidth]{columns.pdf}
              \caption{}~\label{fig:wedge_layer}
      \end{subfigure}
      \begin{subfigure}{0.325\linewidth}
              \centering
              \includegraphics[width=0.6\textwidth]{hexagonal_packing.pdf}
              \caption{}~\label{fig:hex_packing_simple}
      \end{subfigure}
      \caption{(a) The LLC monomer Na-GA3C11 exhibits wedge-like character. (b) 
      Monomers stack on top of each other to create columns with short range order,
      then assemble into pores with hydrophilic head groups (blue) facing towards 
      the pore center. (c) The pores assemble into hexagonally packed columnar
      mesophases.}~\label{fig:assembly}
      \vspace{-0.5cm}
  \end{figure}

  I evaluated a number of different initial column architectures. I stacked 
  monomers in parallel displaced and sandwiched configurations, two possible
  $\pi$-$\pi$ stacking modes.\cite{sinnokrot_estimates_2002} I also varied 
  the initial distance between stacked monomers, $d$. I chose $d$ values of 
  3.7~\AA, based on experimental WAXS measurements, as well as 5 \AA~as a 
  test of its sensitivity.

  My model's geometry is most consistent with experiment for systems built
  with 5 columns per pore and monomers initially stacked 3.7 \AA~apart
  (see Figure~\ref{fig:p2p}). I equilibrated systems with 4, 5, 6, 7 and 
  8 columns per pore. The pore spacing of 5 column-per-pore systems agree
  well with experiment. 6 column-per-pore systems built with $d$ = 5 \AA~appear
  to yield an experimentally consistent pore spacing however, the equilibrated
  distance between stacked monomers stay close to 5 \AA~which is inconsistent
  with experiment. In general, the equilibrated distance between monomers stays
  close to its initial value.

  \begin{wrapfigure}{L!}{0.5\textwidth}
    \centering
    \vspace{-0.5cm}
    \includegraphics[width=\linewidth]{p2p.pdf}
    \caption{Systems with 5 columns per pore have equilibrated pore spacings
            closest to the experimental value of 4.12 nm. The equilibrated 
            pore spacing of the model increases as the number of columns in
            each pore increases.}~\label{fig:p2p}
    \vspace{-1.75cm}
  \end{wrapfigure}

%  \noindent On timescales that we can reasonably simulate, our equilibrated
%  structures show some initial configuration dependence. 
%  \begin{itemize}
%    \item All of the systems used to generate Figure~\ref{fig:p2p} are at 
%    least metastable.
%    \item We observe no large scale rearrangements that would be necessary
%    for all structures to converge to the same equilibrium structure. 
%  \end{itemize}

  I further validated the structure of our molecular model by verifying 
  its consistency with five major reflections present in the experimental
  WAXS pattern. Using MD trajectories, we simulated X-ray diffraction (XRD)
  patterns that we could compare to the WAXS data by taking the appropriate
  cross-section of the time-averaged 3D structure factor. The five major 
  reflections and the structural features leading to them are summarized
  and described in the caption of Figure~\ref{fig:WAXS_comparison}. None of
  the models simulated to this point could reproduce R-double.

  It is necessary to add a small amount of water to the model in order to 
  fully reproduce all features of the WAXS pattern (see Figure~\ref{fig:solvated_XRD}).
  I obtain the most experimentally consistent structure when I build systems
  in the parallel displaced configuration with 1 wt\% water added to the pores. 
  This is the only configuration that gives rise to the R-double feature. R-double
  appears because vertically adjacent monomer head groups hydrogen bond with shared 
  water molecules, causing them to be drawn closer together. When multiple 
  pairing interactions occur in series along the same pore axis, the center
  of masses of the pairs are spaced apart at twice the $\pi$-stacking distance. 
  The necessity of water in my model suggests that the membrane synthesized by
  Feng et al. was slightly hydrated due to water molecules leached from 
  surroundings by the hydroscopic monomers. Similar suspicions have been voiced
  by experimentalists in unpublished communications.

  \begin{wrapfigure}{R!}{0.65\textwidth}
  	\centering
  	\vspace{-0.5cm}
    \begin{subfigure}{0.49\linewidth}
    	\includegraphics[width=\linewidth]{WAXS_annotated_words.pdf}
        \caption{Experiment}\label{fig:WAXS}
    \end{subfigure}
	\begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{solvated_offset_rzplot_1.pdf}
        \caption{Simulation}\label{fig:solvated_XRD}   
	\end{subfigure}
    \caption{(a) 2D-WAXS gives details about repeating features on the order of
    angstroms. Explanations for each of the 5 major reflections present are
    as follows: (R-$\pi$) Aromatic head groups $\pi-\pi$ stack 3.7 \AA~apart. 
    (R-double) Monomer head groups associate into pairs by hydrogen bonding with
    a shared water molecule. (R-alkanes) Alkane chain tails pack 4.5 \AA~apart. 
    (R-spots) Monomer tails pack hexagonally. (R-pores) The pores are spaced 
    4.12 nm apart and pack hexagonally. (b) We obtain a maximally consistent
    match between simulated XRD patterns and experimental WAXS data when we
    build systems in the parallel displaced configuration with 1 wt \% water
    included in the pores.}\label{fig:WAXS_comparison}
    \vspace{-0.75cm}
 \end{wrapfigure}

  %BJC: I think this is important to say somewhere
  %BJC: but I don't have an ensemble of 1 wt % systems to fully prove that with.
  %BJC: I could show cross-sections of R-pi, but that would require more context. 
  On the timescales that we can reasonably simulate, our model exhibits slow 
  dynamics. Consequently, there are very few uncorrelated frames in our 
  trajectories which leads to noise and sharpened reflections in the simulated
  XRD patterns. We can overcome this issue by combining the structure factors
  generated from an ensemble of trajectories produced starting from an ensemble
  of uncorrelated initial configurations.

  The composition of the pores does not change regardless of which
  initial configuration we study. In Figure~\ref{fig:overlaid_densities}, we 
  plot the radial density of various monomer components as a function of distance
  from the pore centers. All are qualitatively similar meaning that a solute 
  placed in any of these systems should experience a similar chemical environment.
  Although we will move forward with our most promising configuration, this is 
  an important finding since it implies that we do not need to apply the same level
  of rigor when screening new monomers.

  My work towards this objective is published in the Journal of Physical Chemistry B:
  Coscia et al. \textit{J. Phys. Chem. B}, 123, 289-309 (2019).

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{\textwidth}
  %BJC: probably change this legend to `d = x Angstroms` instead of 
  % disordered/ordered, since I don't have space to talk about ordered / disorderd configurations (requires 
  % nematic order parameter to give a complete justification of the names)
  \includegraphics[width=\textwidth]{regional_density_legend.pdf}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{head_group_density.pdf}
        \caption{Head Groups}
        \label{fig:head_groups_regional_density}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{tails_density.pdf}
        \caption{Tails}
        \label{fig:tails_regional_density}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{sodium_density.pdf}
        \caption{Sodium Ions}
        \label{fig:sodium_regional_density}
  \end{subfigure}
  \caption{In all systems studied, the component radial distribution functions are similar.
      They exhibit a composition gradient transitioning from the hydrophilic to the hydrophobic
          regions. The biggest differences are at r=0 where noise is higher due to
          decreased sampling. The center of the pore is not hollow, but contains sodium ions and
          head groups, even when the system is solvated. This architecture may impede transport in
          the real system in a chemically-dependent manner.
          The solvated system has a lower density of head groups near the
          pore center which is likely due to the swelling that is necessary in order to fit water
          molecules in the pore region.}~\label{fig:overlaid_densities}
  \vspace{-0.5cm}
  \end{figure}

  \noindent \textbf{\large Objective 2:} \textit{\large Determine transport mechanisms} (\textcolor{green!40!olive}{\textbf{Complete}})

  We added additional water to our most experimentally consistent structural
  model in order to create a higher water content H\textsubscript{II} phase
  model. Researchers have synthesized the H\textsubscript{II} phase using 
  the NaGA3C11 monomer with water contents ranging from 7 - 20 wt\%, with 
  10 wt\% being the most common.~\cite{smith_ordered_1997,zhou_supported_2005} 
  However, Resel et al. stated that the system is likely fully hydrated at
  7 wt \% water with additional water trapped in defects between mesophases.~\cite{resel_structural_2000}
  Consequently, we chose to build two systems with 5 and 10 wt\% water.

  \begin{wrapfigure}{l}{0.65\textwidth}
  \centering
  \vspace{-0.4cm}
%  \begin{subfigure}{0.325\textwidth}
%  \includegraphics[width=\textwidth]{component_density_dry.pdf}
%  %BJC: simulation is extended to smooth out these curves
%  \caption{Dry System}\label{fig:component_density_dry}
%  \end{subfigure}
%  \begin{subfigure}{0.45\textwidth}
%  \vspace{-0.5cm}
%  \includegraphics[width=\textwidth]{monomer_color_coded.pdf}
%  \label{fig:monomer_color_coded}
%  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{component_density_5wt.pdf}
  \caption{5 wt\% water}\label{fig:component_density_5wt}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{component_density_10wt.pdf}
  \caption{10 wt\% water}\label{fig:component_density_10wt}
  \end{subfigure}
  \caption{The radial densities of various monomer components paint a
  picture of the pore topology where the pore centers are primarily composed
  of water and sodium ions. The monomer groups labeled in each plot correspond to
  the color-coded monomer pictured in Figure~\ref{fig:HII_mon}. All RDFs
  represent the number of atoms located at a given distance from the pore center
  normalized by the volume of the annular bin to which they belong. The black
  dashed lines are positioned so that they intersect with the maximum head group
  density. (a) In the 5 wt \% system, monomer head groups are situated about 0.45 nm
  from the pore center. Sodium and water are densest near the pore center 
  (b) Monomers in the 10 wt \% system retreat an additional 0.2 nm to make room
  for more water. Water is densest at the pore center, but sodium ion density
  peaks closer to the head groups since many stay bound to carboxylate groups.}\label{fig:component_densities}
  \vspace{-0.75cm}
  \end{wrapfigure}

  Upon simulating each system, we observed that water partitions into the
  distal tail region. Based on the radial distance from the pore center where 
  the minimum water density occurs, we defined the distal tail region to 
  be greater than 1.5 nm from any pore center. There
  is approximately a 3:2 ratio of water in the pores to water in the 
  distal tails. Due to the wedge shape of the monomers, the distal tail
  region has a relatively low density leaving space for water molecules 
  to fill.

  The pores of the H\textsubscript{II} phase are primarily a mixture of 
  water and sodium ions. The maximum density of head groups occurs 0.45
  and 0.65 nm from the pore center in the 5 and 10 wt\% water systems 
  respectively (see Figure~\ref{fig:component_densities}). There are 
  almost no head groups occupying the pore center.

  Water and sodium transport is significantly faster in the less crowded 
  10 wt\% water pores. The mean squared displacement (MSD) of water is about
  51 times higher and the MSD of sodium is about 49 times higher compared to
  the 5 wt\% system. In general we observe similar solute transport mechanisms
  in the 5 and 10 wt\% water systems but on different time scales.
    %BJC: might as well leave 5 wt% data in there.
%    \item Therefore, we focus the remainder of this discussion on the 10 wt\%
%    water system since we were able to collect a larger quantity of mechanistic
%    data from the solute trajectories.

  % BJC: do you think I can get away with an appendix with the solute table + abbreviations that
  % doesn't count towards the page limit?
  We observed transport of 20 different small polar solutes in the pores of
  our model. We created a separate initial configuration for each solute 
  studied and placed 6 solutes, equally spaced in $z$, in each nanopore.
  After a 5 ns equilibration, we allowed the solutes to simulate for
  1 microsecond.

  \begin{figure}
  \includegraphics[width=\textwidth]{solute_table.pdf}
  \caption{The solutes studied in this work along with the abbreviations used
  in subsequent figures and their chemical structures. Solutes are color-coded
  according to similarities in their structure. Blue corresponds to simple 
  alcohols, red to diols, triols and sugars, green to ketone-like solutes, 
  orange to sulfur-containing solutes and yellow to solutes that can only accept
  hydrogen bonds but not donate.}\label{fig:solute_table}
  \vspace{-0.5cm}
  \end{figure}

%  \begin{wrapfigure}{R}{0.6\textwidth}
%  \centering
%  \begin{subfigure}{0.45\linewidth}
%  \includegraphics[width=\linewidth]{all_10wt_tamsds.pdf}
%  \caption{10 wt \% water}\label{fig:all_msds_10wt}
%  \end{subfigure}
%  \begin{subfigure}{0.45\linewidth}
%  \includegraphics[width=\linewidth]{msd_radius_10wt.pdf}
%  \caption{10 wt \% water}\label{fig:msd_radius_10wt}
%  \end{subfigure}
%  \caption{The MSDs of solutes in the 5 wt \% water system (a) are significantly
%  smaller than those of the solutes in the 10 wt \% water system (b). The
%  MSDs are not a monotonic function of molecular size (c and d). A significant
%  number of solute MSDs fall below the theoretical lines predicted by the
%  Stokes-Einstein equation and Gierer and Wirtz' corrected Stokes-Einstein equation.
%  }\label{fig:msds}
%  \vspace{-1.5cm}
%  \end{wrapfigure}

  The solute MSDs are not a monotonic function of solute size. In 
  Figures~\ref{fig:all_msds_5wt} and~\ref{fig:all_msds_10wt}, we plot
  the average MSD of each solute. In Figure~\ref{fig:msd_radius_5wt} 
  and~\ref{fig:msd_radius_10wt}, we plot solute MSDs against their
  molecular radius. We also show theoretical curves which illustrate the 
  expected MSD of each solute if they were to travel unhindered. We 
  ensured that the Stokes-Einstein equation with the correction factor
  of Gierer and Wirtz\cite{gierer_molekulare_1953} passed through methanol's MSD and radius
  in order to give an approximate frame of reference. The correction 
  factor attempts to include the effects of microfriction that begin to 
  play a role when solute size becomes on the order of solvent size. 
  Since methanol is small, we assume that it travels unhindered relative 
  to all other solutes. The theoretical line serves as an approximate 
  boundary between subdiffusive, Brownian and superdiffusive behavior.
  In the majority of cases, solute MSDs fall well below the theoretical
  lines and therefore are subject to more hindrance than methanol due to
  factors other than size.

  % 5 and 10 wt %
  \begin{wrapfigure}{r}{0.65\textwidth}
  \centering
  \vspace{-0.4cm}
  \begin{subfigure}{0.45\linewidth}
  \includegraphics[width=\linewidth]{all_5wt_tamsds.pdf}
  \caption{5 wt \% water}\label{fig:all_msds_5wt}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \includegraphics[width=\linewidth]{all_10wt_tamsds.pdf}
  \caption{10 wt \% water}\label{fig:all_msds_10wt}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \includegraphics[width=\linewidth]{msd_radius_5wt.pdf}
  \caption{5 wt \% water}\label{fig:msd_radius_5wt}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \includegraphics[width=\linewidth]{msd_radius_10wt.pdf}
  \caption{10 wt \% water}\label{fig:msd_radius_10wt}
  \end{subfigure}
  \caption{The MSDs of solutes in the 5 wt \% water system (a) are significantly
  smaller than those of the solutes in the 10 wt \% water system (b). The
  MSDs are not a monotonic function of molecular size (c and d). A significant
  number of solute MSDs fall below the theoretical lines predicted by the
  Stokes-Einstein equation and Gierer and Wirtz' corrected Stokes-Einstein equation.
  }\label{fig:msds}
  \vspace{-0.4cm}
  \end{wrapfigure}

  Intermittent hops between long periods of entrapment lead to subdiffusive
  solute transport behavior. As implied by Figures~\ref{fig:msd_radius_10wt} and~\ref{fig:msd_radius_5wt}, 
  most solutes move significantly slower than expected. In 
  Figure~\ref{fig:example_ztraces}, we've plotted the $z$-direction trajectory
  of three ethanol molecules. Typically, long periods of entrapment occur
  when ethanol molecules are far from the pore center, while there is a much 
  greater degree of mobility close to the pore center. The MSD curve averaged
  over all ethanol trajectories is shown in Figure~\ref{fig:example_msd}.
  The curve is sub-linear and thus subdiffusive.

  \begin{wrapfigure}{l}{0.6\textwidth}
  \centering
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{colorful_example_ztraces.pdf}
  \caption{}\label{fig:example_ztraces}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{example_msd.pdf}
  \caption{}\label{fig:example_msd}
  \end{subfigure}
  \caption{All solutes show subdiffusive transport behavior inside the membrane's
  nanopores, similar to that exhibited by ethanol. (a) The $z$-coordinate trace of
  3 representative ethanol COMs shows clear periods of entrapment separated by hops.
  In general, the longest dwell times occur when solutes are situated far from the
  pore center and the hops occur when solutes are close to the pore center. (b) The
  time-averaged MSD of ethanol is sub-linear which suggests transport is governed
  by an anomalous subdiffusion process.}\label{fig:qualitative_mechanisms}

  \end{wrapfigure}

  Due to the crowded environment among the monomer tails, solutes generally move
  faster in the less dense pore region. Figure~\ref{fig:hop_lengths} shows that 
  this is the case for all solutes. Hops made in the pore region are on average
  59 \% larger than those made outside the pore region.

  However, time spent in the pore region does not necessarily result
  in a high MSD. For example, ribose spends the largest fraction of time in 
  the pore region, but has the fifth lowest hop frequency and the third 
  lowest average MSD (see Figures~\ref{fig:frac_time} and~\ref{fig:hopfreq}).
  Therefore, proximity to the pore center cannot be the sole variable that
  determines transport properties.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\linewidth]{hop_length.pdf}
  \caption{}\label{fig:hop_lengths}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{frac_time_spent.pdf}
  \caption{}\label{fig:frac_time}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{hopfreq_total.pdf}
  \caption{}\label{fig:hopfreq}
  \end{subfigure}
  \caption{(a) Hops made in the pore region of the 10 wt\% water system
  are, on average, 59 \% larger than those made outside the pore region.
  The trend in hop lengths is similar to the trend in MSDs shown in
  Figure~\ref{fig:all_msds_10wt} implying that solutes which make consistently
  larger hops have higher MSDs. The fraction of time spent by a solute in the
  pore region (b) does not necessarily lead to more frequent hopping (c). For
  example, ribose spends the largest fraction of time in the pore region, yet
  performs the the fifth lowest number of hops.}\label{fig:hops}
  \end{figure}

  Among the set of solutes studied, we observe three different mechanisms 
  of entrapment that are responsible for subdiffusive behavior:
  \begin{enumerate}
    \item As already demonstrated by Figures~\ref{fig:example_ztraces} and 
    Figure~\ref{fig:hop_lengths}, solutes that drift away from the pore center
    can become entangled in the monomer tails. 
    \item Many of the solutes we studied are capable of donating hydrogen bonds to
    monomer head groups and thus are prone to temporary immobilization through 
    this interaction.
    \item Because all solutes are polar, they have regions of concentrated electron
    density, modeled as partial charges, which can associate with sodium ions
    bound to monomer head groups.
  \end{enumerate}

  \begin{wrapfigure}{L}{0.7\textwidth}
  \centering
  \vspace{-0.35cm}
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{nhbonds_all.pdf}  % unique_hbonds.py
  \caption{}\label{fig:nhbonds}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{hbond_lifetime.pdf}  % hbond_dwells.py
  \caption{}\label{fig:hbond_dwells}
  \end{subfigure}
  \caption{(a) Solutes capable of donating hydrogen bonds to monomer head groups
  do so to varying degrees. The reported percentages represent unique solute-monomer
  hydrogen bonds. Individual solutes that hydrogen bond with multiple head groups
  simultaneously are only counted once. (b) The lifetime of individual hydrogen
  bonds appears correlated to the percentage of solutes involved in hydrogen bond
  interactions. Hydrogen bond lifetimes tend to be longer for solutes that
  hydrogen bond frequently. Note that solutes incapable of donating hydrogen
  bonds are omitted from this figure.}\label{fig:hbonds}
  \vspace{-0.5cm}
  \end{wrapfigure}

  The frequency with which solutes donate hydrogen bonds to monomer head 
  groups is related to the number of hydrogen bond donating atoms as well
  as their identity. The percentage of solutes actively participating in
  at least one hydrogen bond with a head group each frame descends as the
  number of hydroxyl groups decreases (Figure~\ref{fig:nhbonds}). Solutes with many hydroxyl groups,
  such as ribose, tetrose and glycerol, can donate multiple hydrogen bonds
  to monomer head groups simultaneously. When one hydrogen bond is broken,
  other hydrogen bonds work to hold the solute in place, which allows 
  broken hydrogen bonds to reform. Solutes containing sulfur and 
  nitrogen atoms in place of oxygen atoms hydrogen bond less frequently
  since they are less electronegative elements.~\cite{biswal_hydrogen_2015} % citation
  The lifetime of hydrogen bonds follows nearly the same trend 
  (Figure~\ref{fig:hbond_dwells}). Hydrogen bonds of solutes that hydrogen
  bond more frequently last longer.

  \begin{wrapfigure}{r}{0.7\textwidth}
  \centering
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{all_solutes_NA_coordination.pdf}
  \caption{}\label{fig:all_solutes_NA_coordination}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{all_NA_dwells.pdf}
  \caption{}\label{fig:all_NA_dwells}
  \end{subfigure}
  \caption{(a) Solutes, especially those with carbonyl groups, spend a
  significant fraction of time coordinated to sodium ions. (b) The length
  of time a solute-sodium pairs spends associated tends to be higher for
  pairs that associate more frequently.}\label{fig:na_coordination}
  \vspace{-0.5cm}
  \end{wrapfigure}

  Solutes with carbonyl groups tend to associate with sodium ions most frequently.
  Nearly all of the most coordinated solutes contain a carbonyl group, except
  for DMSO which has an analogous sulfinyl group (see Figure~\ref{fig:all_solutes_NA_coordination}). There is a significant drop
  in sodium ion association for solutes that do not contain carbonyl groups or
  multiple hydroxyl groups to compensate.
  The corresponding dwell times follow a similar trend, however the dwell
  times of highly coordinated solutes with multiple hydroxyl groups are generally
  lower since association between hydroxyl groups and sodium is apparently a 
  weaker interaction (Figure~\ref{fig:all_NA_dwells}). Solutes with nitrogen 
  atoms adjacent to the carbonyl groups tend to associate with sodium ions 
  significantly more.

  A manuscript based on the work presented towards this objective is under 
  review for publication in the Journal of Physical Chemistry B.

  \begin{figure}[h!]
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\linewidth]{hop_distribution_GCL.pdf}
  \caption{}\label{fig:hop_dist}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\linewidth]{dwell_distribution_GCL.pdf}
  \caption{}\label{fig:dwell_dist}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\linewidth]{hop_acf_GCL.pdf}
  \caption{}\label{fig:hop_acf}
  \end{subfigure}
  \caption{The movement of solutes may be well-described by an sFBM process 
  which can be simulated using parameters that describe the solute's hop lengths, 
  dwell times and degree of correlation between hops. Here I use the parameters 
  derived from ethylene glycol as an example. (a) The distribution of hop
  lengths made by ethylene glycol appears Gaussian, which can be described by the
  maximum likelihood estimate (MLE) of the $\sigma$ parameter, with assumed mean, 
  $\mu$, of 0. (b) The distribution of dwell times between hops can be fit to a 
  power law whose decay is described by the MLE of the parameter $\alpha$. 
  (c) Hops are anticorrelated to their previous hop as indicated by the negative
  value at $k=1$ of the autocorrelation function, $\gamma$. The Hurst 
  parameter, $H$, is a measure of long term memory of the hops and is directly
  related to the autocorrelation function of fractional Brownian process.}\label{fig:distributions}
  \end{figure}

  \noindent \textbf{\large Objective 3:} \textit{\large Create a stochastic model} (\textcolor{blue}{\textbf{In Progress}})

  While one can clearly learn a great deal of qualitative information about 
  mechanistic behavior on MD-accessible timescales, my work's impact would
  greatly increase if I could create a model that could predict macroscopic
  quantities such as permeability and selectivity within a reasonable degree
  of certainty. A stochastic model derived from simulation data could achieve 
  this however, there is not a well-defined procedure for this type of 
  extrapolation. Therefore, I am exploring two different approaches towards
  this end.

  \begin{wrapfigure}{r}{0.4\textwidth}
  \centering
  \vspace{-0.6cm}
  \includegraphics[width=\linewidth]{all_tamsds.pdf}
  \caption{The trends in the MSD calculated directly from MD simulations and from
  realizations of the sFBM process mostly agree, but the sFBM predictions are generally
  far too low.}\label{fig:sFBM_MSDs}
  \vspace{-0.5cm}
  \end{wrapfigure}

  \textit{Approach 1: subordinated fraction Brownian motion}: My first route 
  towards a stochastic model is to simulate realizations of a subordinated 
  fractional Brownian motion (sFBM) process based on parameters extracted 
  from solute time series.~\cite{meroz_subdiffusion_2010} As demonstrated 
  by the example in Figures~\ref{fig:hop_dist} and~\ref{fig:dwell_dist},
  hop length distributions in the $z$-direction appear Gaussian while the 
  distribution of dwell times between hops appears to be distributed 
  according to a power law. Alone, these distributions characterize a 
  continuous time random walk (CTRW), which is a subdiffusive process where
  particles hop a random distance in a random direction and stay there for
  a random amount of time.~\cite{meroz_toolbox_2015}. Unlike a CTRW, the 
  hops in our system appear anti-correlated to their previous hop as shown
  by the autocorrelation function in Figure~\ref{fig:hop_acf}. A CTRW with
  anticorrelated hops is well-described by an sFBM process.

  \begin{wrapfigure}{l}{0.6\textwidth}
  \centering
  \vspace{-0.5cm}
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{breakpoints.pdf}
  \caption{}\label{fig:breakpoints}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{first_order_difference.pdf}
  \caption{}\label{fig:first_order_difference}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{hop_density.pdf}
  \caption{}\label{fig:hop_density}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{naive_density_forecast.pdf}
  \caption{}\label{fig:naive_density_forecast}
  \end{subfigure}
  \caption{In contrast to our first approach which relies on accurate 
  detection of hops, density forecasting uses all of the time series 
  information at our disposal. (a) Our first approach identifies hops 
  based on breakpoints, indicated by color transitions. All information 
  is lost between breakpoints. (b) Instead we can calculate hop lengths
  between each time step and (c) compile them into a single probability 
  density. (d) As a first approximation, one can simulate time series
  by randomly drawing from the distribution in (c).}\label{fig:density_forecast}
  \vspace{-1.5cm}
  \end{wrapfigure}

  Solute MSDs calculated based on realizations of an sFBM process mostly
  preserve the trends in MSD calculated directly from MD simulations (see
  Figure~\ref{fig:sFBM_MSDs}), but the predictions are generally too low. 
  For each solute, I used the parameters described in Figure~\ref{fig:distributions}
  to generate $1e4$ realizations of an sFBM process, each with the same
  length as the MD simulations (1000 ns) so that I could make a direct 
  comparison between their MSDs. While the sFBM MSD prediction for methanol
  is too high, the rest are on average only 46\% of the MD calculated MSD
  values.

  This approach might be improved by incorporating radial dependence into
  the three sFBM parameters described in Figure~\ref{fig:distributions}. 
  In Figure~\ref{fig:hop_lengths}, we showed that hops by solutes situated 
  inside the pores are 59\% larger than hops by solutes in the tails, which 
  leads to two distinct values of $\sigma$. Similarly, we can determine 
  radially dependent values of $\alpha$ and $H$.

  \begin{wrapfigure}{r}{0.4\textwidth}
  \centering
  \vspace{-0.5cm}
  \includegraphics[width=\linewidth]{z_histogram.pdf}
  \caption{The histogram of $z$ values coming from $z_t$ calculated with 
  Equation~\ref{eqn:pit} appears uniform meaning $p(h)$ is well-sampled 
  by the out-sample data.}\label{fig:z_histogram}
  \vspace{-1cm}
  \end{wrapfigure}

  A simple two state model, in pore versus out of pore, may be sufficient to 
  fully incorporate radially dependent effects. Based on the fraction of time 
  spent in the pore region (Figure~\ref{fig:frac_time}), we can determine
  the appropriate balance of each type of parameter to use when simulating sFBM
  realizations. We would also need some kind of switching frequency, or dwell
  time in each region since it does not make physical sense for a solute to 
  enter and exit the pore region every other time step.

  If the two state model proves insufficient, I can add increasing levels of
  complexity. For example, I might make the radial dependence a continuous 
  function, or I could assign parameters that are dependent on the type of
  trapping mechanism affecting the solute.

  \textit{Approach 2: density forecasting}: Density forecasting can provide
  predictions based on properties of a time series without any knowledge
  of the physical interactions leading to those properties.

  \begin{wrapfigure}{l}{0.6\textwidth}
  \centering
  \vspace{-0.75cm}
  \includegraphics[width=\linewidth]{correlograms.pdf}
  \caption{Correlograms of $z_t$ reveal dependence for all 4 powers of 
  $(z - \overline{z})$.}\label{fig:correlograms}
  \vspace{-0.5cm}
  \end{wrapfigure}

  In contrast to the first approach, density forecasting allows one to use
  all of available trajectory information. In order to construct distributions
  of hop lengths and dwell times with my first approach, I needed to 
  locate precise time points at which hops occurred. We achieved this using 
  an off-line breakpoint detection algorithm implemented in the python package 
  \texttt{ruptures}.~\cite{truong_review_2018} The algorithm does a decent
  job of identifying hops, but still loses information as demonstrated in
  Figure~\ref{fig:breakpoints}. With density forecasting, one can instead 
  compile the hop length and magnitude between each time step 
  (Figure~\ref{fig:first_order_difference}) into a probability density
  like that shown in Figure~\ref{fig:hop_density}. As a first attempt, 
  one could randomly draw from this distribution in order to construct
  a time series like that shown in Figure~\ref{fig:naive_density_forecast}.

  One can test how well a time series of hop lengths, $h_t$, is described by the 
  hop density, $p(h)$, in Figure~\ref{fig:hop_density} using the probability
  integral transform:~\cite{diebold_evaluating_1998}
  \begin{equation}
  z_t = \int_{-\infty}^{h_t} p(u) du
  \label{eqn:pit}
  \end{equation}
  Generally, one uses the first half of a time series, or in-sample data, to
  create $p(h)$ and applies Equation~\ref{eqn:pit} to the second half of
  the time series, or out-sample data. If the distribution of $z_t$ values
  is uniform, then the cumulative distribution function of $p(h)$ has been
  sampled uniformly by the out-sample data and there is evidence that the 
  time series is well-described by $p(h)$. Using the out-sample data from 
  Figure~\ref{fig:density_forecast}, we observe a uniform distribution of
  $z_t$ values (Figure~\ref{fig:z_histogram}).

  However, it is also necessary to look for correlations in $z_t$. Diebold
  et al. found that examination of the correlograms of $(z - \overline{z})$,
  $(z - \overline{z})^2$, $(z - \overline{z})^3$ and $(z - \overline{z})^4$,
  which represent dependence of the mean, variance, skewness and kurtosis 
  respectively, are adequate to identify potentially sophisticated and 
  nonlinear forms of past dependence.~\cite{diebold_evaluating_1998} Continuing
  with our example above, the correlograms of $z_t$ shown in Figure~\ref{fig:correlograms} 
  exhibit serial dependence for all powers of $(z - \overline{z})$. Our 
  time series exhibits dependence on its past values, a feature that was 
  not included in the forecast shown in Figure~\ref{fig:naive_density_forecast}.
  %This likely explains the qualitative difference in the time series of 
  %Figure~\ref{fig:breakpoints} versus~\ref{fig:naive_density_forecast}.

  Going forward, I will need to incorporate previous dependence in the 
  form of a conditional density forecast with the form $p(h|\Omega_t)$ where
  $\Omega_t$ incorporates past hop lengths and radial distances from
  the pore center. These types of autoregressive models have been widely
  studied in the context of economics, but are still well-suited for 
  application to solute transport.~\cite{hamilton_time_1994}

  \noindent \textbf{\large Objective 4:} \textit{\large Apply analyses to 
  Q\textsubscript{I} phase} (\textcolor{blue}{\textbf{In Progress}})

  Paralleling our work on the H\textsubscript{II} phase, my first task towards
  this objective was to build a suitable unit cell representation of the 
  Q\textsubscript{I} phase. Unfortunately, the true space group of the Q\textsubscript{I} phase that we
  are studying is unknown. Experimental diffraction has narrowed down the 
  possible bicontinuous cubic configurations to the Ia3d and Pn3m space groups. 
  Therefore, we will build both types of unit cells and search for clues that
  can be used to differentiate them experimentally.

  \begin{wrapfigure}{l}{0.6\textwidth}
  \centering
  \vspace{-0.5cm}
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{gyroid.png}
  \caption{}\label{fig:Ia3d}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
  \includegraphics[width=\linewidth]{schwarz.png}
  \caption{}\label{fig:pn3m}
  \end{subfigure}
  \caption{I developed a procedure capable of building both (a) Ia3d and (b) Pn3m
  unit cells from arbitrary monomers with controllable pore size. The blue
  glycerol molecules highlight the aqeuous region where transport occurs.}\label{fig:q1_unitcells}
  \vspace{-0.5cm}
  \end{wrapfigure}

  I have developed a procedure that can build both Ia3d and Pn3m unit cells.
  To our knowledge, nobody has ever built an atomistic bicontinuous cubic 
  phase system without the aid of self-assembly. Self-assembly can be a long
  process, especially for fully atomistic systems. Instead, we used known
  analytical equations that describe the surface of these systems in order 
  to place monomers into a unit cell. Monomers are placed perpendicular to
  the surface in random locations that don't overlap. One can control the 
  pore size by translating monomers perpendicular to the surface at the 
  point where they are attached.

  First, I will study the structure of the Q\textsubscript{I} phase using
  methodology similar to that used for the H\textsubscript{II} phase 
  I will attempt to identify one of the space groups as the most likely
  experimental configuration based on simulated XRD signatures, although
  this might not work due to the known similarity of Ia3d and Pn3m XRD 
  patterns. %cite

  Then I will study transport of solutes within these atomistic unit cells, 
  adapting the same analyses used in Objectives 2--3. The primary 
  challenge will be adapting my techniques to the more complex pore geometry
  of the Q\textsubscript{I} phase, however I expect that main conclusions
  will not deviate far from those of the H\textsubscript{II} phase since
  the chemical environments within the nanopores are qualitatively similar.. We may be 
  able to identify which space group is correct if there are significantly
  different transport properties between systems.

%BJC: not sure how much more detail to include. Can go into a lot. Have to expand box and slowly compress. Add glycerol iteratively until there are no gaps.

%  \noindent \textbf{\large Objective 5} \textit{\large Create a well-documented python package} (\textcolor{blue}{\textbf{In Progress}})
%  
%  Python scripts used to simulate X-ray diffraction patterns are available in the following
%  GitHub repository: \url{https://github.com/joeyelk/MD-Structure-Factor}. Documentation
%  is forthcoming. 
%  
%  Python scripts used to conduct all other post-simulation trajectory analysis are 
%  also available on GitHub: \url{https://github.com/shirtsgroup/LLC_Membranes}. 
%  Documentation is a work in progress. It is available for viewing in its 
%  current state at \url{https://llc-membranes.readthedocs.io/en/latest/}.

  \section{Timeline for Completion of Objectives}\label{section:timeline}

  A schematic of the estimated timeline that I plan to follow for the 
  completion of tasks pertinent to finishing all objectives is given 
  in Table~\ref{table:timeline}. Ideally, each task will become a 
  published manuscript. I will run simulations required to 
  study the structure of the bicontinuous cubic phase in
  parallel with the continued development of my stochastic model of 
  transport in the H\textsubscript{II} phase. Simulations and analysis
  required for Q\textsubscript{I} phase solute transport studies analogous
  to those of Objective 2 will be carried out throughout Fall 2019. 
  Application of a stochastic model to the Q\textsubscript{I} phase will
  be finished by May 2020, just before I defend my dissertation.

  \begin{center}  
  \begin{table}[!htb]
	\renewcommand\arraystretch{1.4}\arrayrulecolor{LightSteelBlue3}
	\captionsetup{singlelinecheck=false, font=blue, labelfont=sc, labelsep=quad}
	\caption{Estimated Timeline for Completion of Objectives}\vskip -1.5ex
	\begin{tabular}{@{\,}r <{\hskip 2pt} !{\foo} >{\raggedright\arraybackslash}p{8cm}}
	\toprule
	\addlinespace[1.5ex]
	August 2019 & Complete Stochastic Model for H\textsubscript{II} phase \\
	September 2019 & Finalize Q\textsubscript{I} phase structure \\
    January 2020 & Finish transport study of Q\textsubscript{I} phase \\
    April 2020 & Finish application of stochastic model to Q\textsubscript{I} phase \\
    May 2020 & PhD Defense \\
	\end{tabular}
	\label{table:timeline}
  \end{table}
  \end{center}
  \vspace{-2cm}

  \section{Resource Requirements}\label{section:resources}

  The remainder of our work will require the use of high performance computing (HPC)
  resources. We will continue using Bridges, an XSEDE resource as well as Summit, a
  supercomputer located at CU Boulder.

  \section{Safety Considerations}\label{section:safety}

  Although my work is confined to computer work in an office space, there are 
  associated health risks that I must mitigate. Proper ergonomics ensure 
  maximum comfort and help me to work efficiently. Sitting up straight and
  having proper back support are necessary to relieve pressure on the discs
  of my vertebrae. My monitors are positioned an arm's length directly in 
  front of me with the tops of the monitors at eye level in order to reduce
  head, neck and eye strain. Finally, I take frequent breaks and go for 
  walks around the building.

  \newpage
  \bibliographystyle{ieeetr}
  \bibliography{comps}

\end{document}