### 2020-02-12 Ben_Manuscripts/thesis/thesis.tex

\documentclass[defaultstyle,11pt]{thesis}

\usepackage{amssymb}		% to get all AMS symbols
\usepackage{graphicx}		% to insert figures
%\usepackage{hyperref}		% PDF hyperreferences??
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{textcomp} % gets rid of \perthousand and \micro warnings that come from gensymb
\usepackage{gensymb}
\usepackage{subcaption}
\usepackage{tcolorbox}
\usepackage[export]{adjustbox}
\usepackage{listings}

\DeclareUnicodeCharacter{03C9}{$\omega$}
\DeclareUnicodeCharacter{03C0}{$\pi$}

\definecolor{amber}{rgb}{0.99, 0.76, 0.0} % golden poppy

\title{Understanding Structure and Selectivity in Nanostructured Lyotropic Liquid Crystal Membranes using Molecular Simulations}

\author{Benjamin Joseph}{Coscia}

\otherdegrees{B.A., University of Connecticut, 2014 \\
	          M.S., University of Colorado Boulder, 2017}

\Degree{Doctor of Philosophy}		                %  #1 {long descr.}
	   {Ph.D., Chemical Engineering}		%  #2 {short descr.}

\dept{Department of}			                %  #1 {designation}
	 {Chemical and Biological Engineering}		%  #2 {name}

\advisor{Prof.}				        %  #1 {title}
	    {Michael Shirts}			%  #2 {name}

\reader{Prof. Richard Noble}		%  2nd person to sign thesis
\readerThree{Prof. Douglas Gin}		%  3rd person to sign thesis
\readerFour{Prof. Matthew Glaser}
\readerFive{Prof. Robert Davis}

\abstract{

  % \OnePageChapter	% use if abstract is less than one page

  % BJC: spliced together abstracts from papers with some additions

  The ability to perform highly selective aqueous separations at low cost
  is beneficial to a variety of chemical industries

  Periodic, nanostructured porous polymer membranes made from the cross-linked
  inverted hexagonal phase of self-assembled lyotropic liquid crystals (LLCs) are
  a promising class of materials for selective separations.

  The uniform size and complex chemical topology of the pores formed by
  self-assembled amphiphilic molecules such as liquid crystals make them
  promising candidates for selective separations.

  In this work, we investigate an experimentally characterized LLC polymer 
  membrane using atomistic molecular modeling. We

  First, we gained a deep understanding of nanoscopic pore structure in order 
  to make sure we got the chemical environment in the pores correct.

  In particular, we compare simulated X-ray
  diffraction (XRD) patterns with experimental XRD data to quantify and
  understand the differences between simulation and experiment.

%  We find that the
%  nanopores are likely composed of 5 columns of stacked LLC monomers which
%  surround each hydrophilic core. Evidence suggests that these columns likely
%  move independently of each other over longer time scales than accessible via
%  atomistic simulation. We also find that WAXS structural features previously
%  attributed to monomer tail tilt is likely instead due to ordered tail packing.
%  Although this system has been reported as dry, we show that small amounts of
%  water are necessary to reproduce all features from the experimental XRD pattern
%  due to asymmetries introduced by hydrogen bonds between the monomer head groups
%  and water molecules.

  Finally, we explore the composition and structure of the
  nanopores and reveal that there exists a composition gradient rather than an
  abrupt partition between the hydrophilic and hydrophobic regions. A caveat is
  that the time scales of the dynamics are extremely long for this system,
  resulting in simulated structures that appear too ordered, thus requiring
  careful examination of the metastable states observed in order to draw any
  conclusions.

  %The clearer picture of the nanoscopic structure of these membranes
  %provided in this study will enable a better understanding of the mechanisms of
  %small molecule transport within these nanopores.

  We observe transport of water, sodium ions and 20 small polar
  solutes within the pores of a lyotropic liquid crystal (LLC) membrane using
  atomistic molecular simulations. We find that the transport of a species is
  dependent not only on molecular size, but on chemical functionality as well.
%  The membrane's inhomogeneous composition gives rise to radially dependent
%  transport mechanisms with respect to the pore centers. 
  We observe that all
  solutes perform intermittent hops between lengthy periods of entrapment. Three
  different trapping mechanisms are responsible for this behavior. First, solutes
  that drift out of the pore can become entangled among the dense monomer tails.
  Second, solutes can donate hydrogen bonds to the monomer head groups. Third,
  solutes can coordinate with sodium counter ions. The degree to which a solute
  is affected by each mechanism is dependent on the chemical functionality of the
  solute.

  %Using the insights developed in this study, we can begin to think about
  %how to redesign existing LLC membranes in order to perform solute-specific
  %separations.

  Mathematically modeling complex transport phenomena at the molecular level
  can be a powerful tool for identifying transport mechanisms and predicting
  macroscopic properties. We use two different stochastic time series models,
  parameterized from long molecular dynamics (MD) simulation trajectories of
  a cross-linked H\textsubscript{II} phase lyotropic liquid crystal (LLC)
  membrane, in order to predict solute mean squared displacements (MSDs)
  and solute flux, and thus solute selectivity, in macroscopic length pores.

  First, using anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time
  random walk. From the MD simulations, we parameterize the distribution of
  dwell times, hop lengths between dwells and correlation between hops. We
  explore two variations of the anomalous diffusion modeling approach. The first variation
  applies a single set of parameters to the solute displacements and the second
  applies two sets of parameters based on the solute's radial distance from the
  closest pore center.

  Next, we generalize Markov state models, treating the
  configurational states of the system as a Markov process where each
  state has distinct transport properties. For each state and transition
  between states, we parameterize the distribution and temporal correlation
  structure of positional fluctuations as a means of characterization and to
  allow us to predict solute MSDs. We show that both models reasonably reproduce
  the MSDs calculated from MD simulations. However, qualitative differences
  between MD and Markov state dependent model-generated trajectories may limit
  its usefulness.

  Finally, we demonstrate how one can use these models to
  estimate flux of a solute across a macroscopic-length pore and, based on
  those quantities, the membrane's selectivity towards each solute. This work
  helps to connect microscopic chemically-dependent solute motions that do not
  follow simple diffusive behavior with macroscopic membrane performance.

  }

\dedication[Dedication]{	% NEVER use \OnePageChapter here.
	To all of the fluffy kitties.
	}

%  Here's where you acknowledge folks who helped. But keep it short, i.e., no more than one page,
%  as required by the Grad School Specifications.
\acknowledgements{	
  \OnePageChapter	% *MUST* BE ONLY ONE PAGE!

  I would like to acknowledge and thank a number of people whose support and
  guidance has not gone unnoticed. First, I thank my advisor, Michael Shirts, 
  who has been instrumental to my growth as a competent and confident researcher,
  pushing me to do high quality science. I thank the entire Shirts group for 
  their feedback and encouragement as well as for making our lab an enjoyable to work.
  I thank Richard Noble and Douglas Gin for their guidance on past and present 
  experimental research on LLC membranes. I also thank Michael McGrath, Gregory 
  Dwulet and Sarah Dischinger, members of the Gin and Noble groups, for providing
  invaluable experimental perspective. I thank Matthew Glaser for his 
  contributions to my understanding of liquid crystal systems. Our first paper 
  would not have been complete without his insight. I thank Joseph Yelk for his 
  effort creating the X-ray diffraction code on which our first paper relied.
  I thank Christopher Calderon for his guidance on applying Bayesian non-parametric
  techniques to solute time series.

  % BJC: should I mention the start-up? If so, I'm not sure what to say for that.
  I would like to acknowledge my financial support: The ACS Petroleum Research Fund
  grant \#59814-ND7 and the Graduate Assistance in Areas of National Need (GAANN) 
  fellowship which is funded by the U.S. Department of Education.

  % BJC: This acknowledgement is how they want to be cited in papers. Could I
  % shorten it for this?
  Finally, I would like to acknowledge my high performance computing resources:
  Molecular simulations were performed using the Extreme Science and
  Engineering Discovery Environment (XSEDE), which is supported by National
  Science Foundation grant number ACI-1548562. Specifically, it used the Bridges
  system, which is supported by NSF award number ACI-1445606, at the Pittsburgh
  Supercomputing Center (PSC). This work also utilized the RMACC Summit supercomputer,
  which is supported by the National Science Foundation (awards ACI-1532235 and
  ACI-1532236), the University of Colorado Boulder, and Colorado State
  University. The Summit supercomputer is a joint effort of the University of
  Colorado Boulder and Colorado State University.

  }

% BJC: The list of figures and tables is obnoxious because the captions are so long.
% BJC: Can I leave them out? (would just uncomment \emptyLoF and \emptyLoT)

% \IRBprotocol{E927F29.001X}	% optional!

% \ToCisShort	% use this only for 1-page Table of Contents

% \LoFisShort	% use this only for 1-page Table of Figures
% \emptyLoF	% use this if there is no List of Figures

% \LoTisShort	% use this only for 1-page Table of Tables
% \emptyLoT	% use this if there is no List of Tables

\begin{document}

\graphicspath{{../transport/figures/}{../structure_paper/figures/}{../transport/supporting_figures/}}

\input macros.tex
\input intro.tex

\input structure.tex
%\input structure_si.tex

\input transport.tex
%\input transport_si.tex

%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{ieeetr}	% or "siam", or "alpha", etc.
%\nocite{*}		% list all refs in database, cited or not
\bibliography{refs}		% Bib database in "refs.bib"

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
\appendix
\input structure_si.tex
\input transport_si.tex

\end{document}