### 2020-03-16 Ben_Manuscripts/thesis/intro.tex
Intro outline ready for a lookX

\chapter{Introduction}\label{chapter:intro}

  \section{Membranes for Aqueous Separations}

  %BJC: limiting to pressure driven aqueous membrane separations (so I don't have to get into electrodialysis, gas separations, dialysis etc.)
  Pressure driven membrane processes have become an increasingly useful tool for
  performing aqueous separations.
  \begin{itemize}
    \item Untreated water can be a very complex solution with particles ranging
    in size from microns down to nanometers.~\cite{goosen_fouling_2005}
    \item Sediment, bacteria, algae, proteins, small organic molecules, and ions
    are all common components of aqueous streams.~\cite{werber_materials_2016}
    \item The diversity in size and chemical topology of contaminants makes the
    design of membrane processes meant to separate these components complex.
    \item To deal with complex aqueous streams, and to prevent excessive membrane
    fouling, multiple types of membranes are often used in series, removing 
    larger particles first.~\cite{suk_membranebased_2006}
  \end{itemize}

  % TFC membranes review -- "Composite reverse osmosis and nanofiltration membranes"
  \noindent Modern membrane design is completely dependent on the target particle 
  separation. 
  \begin{itemize}
    \item Separation membranes are most broadly classified based on the size of 
    the particle they reject.
    \item Microfiltration (MF) membranes contain pores with diameters ranging from 
    100-10000 nm. They can separate large particles like bacteria and protozoa.~\cite{ma_functionalized_2014}
    \item Ultrafiltration (UF) membranes have pore diameters of about 5-500 nm and are useful for the 
    separation of sugars, proteins, viruses and colloidal materials.~\cite{yanful_ultrafiltration_2009}
    \item Nanofiltration (NF) membranes have pores on the order of 1 nm in diameter and can 
    be used for the separation of small organic molecules and charged species.~\cite{van_der_bruggen_review_2003}
    \item Reverse osmosis (RO) membranes are dense amorphous polymers with tortuous 
    pathways formed by collapsed polymer networks. Their dense polymer architecture 
    rejects all solutes to a high degree except for water making them useful for 
    separating hydrated ions from water.~\cite{warsinger_review_2018}
    \item We have summarized the uses of the different classes of membrane technologies in 
    Figure~\ref{fig:size_regimes}.
  \end{itemize}

  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/membrane_separation_size_regimes.pdf}
  \caption{Caption of the future}\label{fig:size_regimes}
  \end{figure}

  Membranes are generally designed to optimize two key performance metrics: 
  permeability and selectivity.~\cite{geise_water_2011}
  %BJC2: Not sure the right place to put citation since its all described in the above citation. I could add more, but seems pointless.
  \begin{itemize}
    \item Permeability is used to quantify the rate at which solvent and solutes
    pass through the membrane.
    \item Selectivity quantifies the relative rates of passage of two substances.
    \item A related quantity to selectivity, rejection, quantifies the fraction
    of undesired solute that does not pass through the membrane.
  \end{itemize}

  Although there has been considerable focus on creating membranes with high
  permeabilities, there is a case to be made that shifting the focus to membrane
  selectivity may offer a more effective route towards lowering the cost of 
  membrane separations.~\cite{werber_materials_2016}
  \begin{itemize}
    \item Increasing permeabilities beyond those already achieved may have 
    only a small effect on energy requirements and capital costs. 
    \item Cohen-Tanugi \textit{et al.} calculated that a theoretical tripling of membrane permeability
    would only reduce energy consumption by up to 15\% in a seawater RO plant.~\cite{cohen-tanugi_quantifying_2014}
    \item Perhaps the biggest detriment to energy consumption is the practical
	need for multiple membrane passes in order to achieve a desired purity.~\cite{singh_production_2009}
    \item Additionally, many membranes are not suited for high purity separations
    of small neutral species which necessitates chemical and capital costs for
    post-treatment.~\cite{shannon_science_2009,fritzmann_state---art_2007}
%    \item The well-known permeability-selectivity tradeoff, common to dense polymer 
%    membranes, directly correlates an increase in water permeability with an 
%    increase in solute permeability.~\cite{geise_water_2011}
    \item Addressing these problems with highly selective separations may lead to 
    more gains.~\cite{deshmukh_desalination_2015}
 \end{itemize}

  \subsection{Applications of Selective Membranes}

  Selective separations are useful in a wide range of applications. They can be used
  to separate components in both gaseous and aqueous phases.~\cite{baker_gas_2014}. We 
  will focus this section on aqueous separations since they are the target process
  for the following studies.

  \textit{Desalination}: Creating potable water from seawater or brackish water is of 
  paramount importance in water-scarce areas. 
  \begin{itemize}
    \item It is estimated that by 2025, 1.8 billion people may live in water-stressed
    regions.~\cite{navarro-ortega_managing_2015} %There's a water stress index in fritzmann_2007
    \item Membrane deslination via RO and thermal muli-stage flash distillation are 
    both widely used processes for creating potable water in these regions.~\cite{fritzmann_state---art_2007}
  	\item Compared with thermal distillation techniques, RO has been shown to 
  	be more environmentally friendly and economical due to its lower energy
  	requirements.\cite{morton_environmental_1997}
% 	\item In general, RO has a smaller footprint % BJC: need citation
  	\item However, thermal techniques are still preferred where excess waste heat or cheap 
  	thermal energy is available, such as cogeneration plants.\cite{bhojwani_technology_2019}
  \end{itemize}

  \textit{Organic Micropollutants}: Municipal and industrial wastewaters are contaminated by harmful
  micropollutants, which can have adverse effects on human health even at low 
  concentrations.\cite{schwarzenbach_challenge_2006}
  \begin{itemize}
    \item Micropollutants include pharmaceuticals, personal care products, 
    hormones, pesticides and industrial chemicals which find their way into
    our drinking water supply.~\cite{barbosa_occurrence_2016}
    \item Sources of micropollutants incude industrial wastewater, 
    agricultural runoff, landfill leachates and domestic effluents.~\cite{mompelat_occurrence_2009}
    %BJC2: There is a review I need to wade through to bulk this up.
    \item Unfortunately, most waste water treatment plants are not 
    designed for complete removal of these contaminants.~\cite{tijani_review_2013}
    \item There is a need for a new generation of membranes that can better
    target specific molecules.
  \end{itemize}

  \textit{Recovery of Valuable Dissolved Species}: In addition to removing contaminants 
  to clean waste water, we can also use highly selective membranes in order to recover
  potentially valuable dissolved species from complex municipal and industrial waste
  streams.~\cite{guest_new_2009,daigger_evolving_2009}

  Municipal waste-waters are rich in carbon, nitrogen and phosphorus-containing 
  compounds.~\cite{romero_raw_2013}
  \begin{itemize}  
    \item The recovery of such products, which can be achieved using a 
    selective membrane, has numerous potential uses.\cite{sales_resource_2015}
    \item For example, nitrogen and phosphorus recovery can help sustain fertilizer 
    production which will help meet global food demand as population continues
    to increase.\cite{xie_membrane-based_2016}
  \end{itemize}

  Industrial waste-waters are often quite complex with up to six times more 
  total dissolved solids than seawater.\cite{werber_materials_2016} 
  \begin{itemize}
    \item For example, flowback water, produced during hydraulic fracturing of
    shale formations consists of relatively high concentrations of salts, metals,
    and soluble organic compounds.~\cite{orem_organic_2014}
    \item The majority of this water is disposed through deep well injection, 
    however there is a growing public concern about its management which has 
    prompted the use of separation technologies such as RO and NF in order to 
    reduce the volume of contaminated water.~\cite{gregory_water_2011}
    \item Some of the dissolved organic compounds in flowback water, such as acetate, 
    are potentially valuable and can be recovered with highly selective membranes
    \cite{dischinger_application_2017}.
  \end{itemize}

%BJC: this is more of a gas separation
%  \textit{Breathable barriers}:
%  
%  Finally, there is a great deal of military interest in creating breathable 
%  barriers which selectively allow the passage of water in both directions but
%  blocks out harmful contaminants.
%  \begin{itemize}
%    \item Need to allow sweating
%    \item Can even put catalytic groups in the pores to break down contaminants
%  \end{itemize}

  \section{Historical Development of Membrane Technology}
  %BJC2: maybe better to title this section "Historical Development of Membrane Technology"

  Membranes for separations are a relatively young technology. Over the past 60 years, 
  the field has grown into a commercially viable technology and the focus has shifted
  primarily towards incremental improvements in performance and most importantly,
  targeted separations. This section will briefly cover a subset of the most relevant
  membrane technologies which have ultimately inspired this thesis.

  \subsection{Reverse Osmosis Membranes}

  The first aqueous separations membranes were developed for RO applications 
  in the 1960s and were made of the amorphous polymer cellulose acetate (CA).~\cite{kesting_semipermeable_1965}
  \begin{itemize}
    \item Reid and Spencer were among the first to show the ability of
    these membranes to reject NaCl to high degrees ($\sim$ 99 \%) while subject 
    to high applied pressures (270 atm).~\cite{reid_ultrafiltration_1960}
    \item The most limiting drawback to CA membranes is their chemical instability, 
    undergoing severe hydrolysis at pH values outside the range 4--6.~\cite{vos_kinetic_1966}
  \end{itemize}

  Amorphous thin film composite (TFC) polyamide membranes are the current industry 
  standard for high purity RO separations.~\cite{baker_membrane_2012}
  \begin{itemize}  
    \item Developed in the 1980s, polyamide membranes are capable of rejecting salt
    to a higher degree and can operate at wider pH ranges and higher feed pressures than 
    CA membranes.~\cite{cadotte_interfacial_1981}
    \item TFC polyamide membranes are typically formed by the interfacial polymerization
    of a polyamide selective layer on the surface of a microporous polysulfone support 
    layer.    
    \item The support layer gives the membrane mechanical strength without hindering the
    flow of water thereby permitting a very thin selective layer where molecular 
    separations occur.~\cite{kucera_biofouling_2019}
  \end{itemize}

  The dense polyamide selective layers of TFC membranes require high feed pressures,
  in the range of 5 -- 120 bar, in order to achieve a useful permeate flux.~\cite{van_der_bruggen_review_2003}
  \begin{itemize}
    \item While the energy requirements to attain these pressures are still less than 
    the energy requirements of thermal distillation\cite{morton_environmental_1997}, it 
    is often desirable to create membranes that can operate at lower feed pressures.
    %BJC2: probably could use some citations here
    \item It is not always necessary to achieve the high purity permeate streams
    resulting from RO.
    \item It is also possible that a target separation may be possible with 
    membranes containing larger pores.
    \item This has driven research into nanofiltration.
  \end{itemize}

  \subsection{Nanofiltration Membranes}

  Since their introduction in the late 1980s, NF membranes have been applied towards
  an increasing number of separations since they bridge the gap in capabilities
  between reverse osmosis and ultrafiltration.~\cite{eriksson_nanofiltration_1988}
  \begin{itemize}
    \item Nanofiltration membranes have pores on the order of 1 nm in size. 
    \item Larger pore sizes allow lower operating pressures (3 -- 20 bar) which lowers 
    energy consumption.~\cite{van_der_bruggen_review_2003}
    \item An ideal NF membrane should have densely packed, uniform-sized and non-tortuous pores.~\cite{jackson_nanoporous_2010}
    This combination has been practically difficult to realize.
  \end{itemize}

  The most widely used methods of NF membrane fabrication are interfacial polymerization
  and phase inversion.
  \begin{itemize}
    \item NF membranes created by interfacial polymerization are made similarly to
    RO membranes. By modifying the monomer materials, one can control the structural morphology
    in order to create pores larger than those in RO membranes.~\cite{seman_nanofiltration_2010}
    \item During phase inversion, one submerges a polymer, dissolved in a solvent, in
    a non-solvent. A solid, porous polymer membrane is all that remains once all
    solvent has been removed by non-solvent exchange.~\cite{smolders_microstructures_1992}
    \item Although commercially dominant, both of these techniques suffer from
    tortuous pores and polydisperse pore sizes which results in far from ideal 
    selectivities.
  \end{itemize}

  Track-etching is another NF fabrication technique which attempts to eliminate 
  polydispersity and tortuosity. During track-etching, a polymer film is bombarded with 
  charged particles, then chemically etched to create pores.~\cite{apel_track_2001}
  \begin{itemize}
	\item Although the pores are uniform and non-tortuous, benefiting selectivity, the 
	membranes have an extremely low pore density and subsequently low permeability. 
  \end{itemize}

  \subsection{Nanostructured Membranes}

  Nanostructured membrane materials have the potential to achieve the high selectivity
  of RO with the low feed pressure requirements of NF. Nanostructured materials of 
  interest typically have explicit nm-size pores that are uniform in size, eliminating
  issues with polydispersity. There have been a number of explorations into different
  kinds of technologies for this application.

  %BJC2: would a figure showing the nanostructures go well here? (a graphene film, a carbon nanotube, a MOF, a zeolite)

  %BJC2: Each tecnhology below will likely have its own heading and be split into 2 or more 
  % paragraphs.
  \textit{Graphene and graphene oxide} Ultrathin-film graphene and graphene oxide membranes
  are an active area of research because they are atomically thick materials and therefore 
  offer potential for extremely high permeability membranes.~\cite{humplik_nanostructured_2011} 
  \begin{itemize}
    \item The tightly packed aromatic network of $sp^2$ hybridized carbon atoms that 
    make up graphene sheets precludes the passage of even the smallest monoatomic species, helium.~\cite{bunch_impermeable_2008}
    \item Controllably-sized pores have been created in pristine graphene by plasma etching
    and by charged particle bombardment.~\cite{surwade_water_2015,russo_atom-by-atom_2012}
    \item Graphene oxide (GO), a highly oxidized modification of graphene, consist of epoxy,
    hydroxyl and carboxylate functional groups.~\cite{hegab_graphene_2015}
    \item Many groups study GO membranes because they have improved antibacterial properties 
    over graphene, leading to less membrane fouling, and because their functionality can be
    modified in order to achieve solute-specific separations.~\cite{liu_antibacterial_2011,he_bioinspired_2013}
    \item Despite their promise, scalable synthesis of monolayer graphene and GO membranes without
    introducing microscopic, performance degrading defects has not yet been achieved.~\cite{cohen-tanugi_multilayer_2016,wei_multilayered_2018}
    \item Multilayered GO membranes have shown considerably more progress since their 
    synthesis procedures are low cost and scalable.~\cite{zhang_size-controlled_2009}
    \item GO sheets are particularly easy to make into multilayered membranes since they 
    stack on top of each other, held together by hydrogen bonds.~\cite{homaeigohar_graphene_2017}
    \item The potential of graphene and GO membrane certainly merits their continued exploration. Many
    groups are still working in order to work on better control and standardization of their synthesis,
    improving their designability for specific separations, and further improving their scalability.~\cite{wei_multilayered_2018}
  \end{itemize}

  \textit{Carbon nanotubes}: Carbon nanotubes (CNTs) have shown promise as aqueous separations membranes
  due to unprecedentedly fast water transport.\cite{humplik_nanostructured_2011}
  \begin{itemize}
    \item CNTs are an allotrope of carbon consisting of tubular graphitic layers.~\cite{iijima_helical_1991}
    \item One can control CNT diameters down to 0.5 nm in size which has inspired the
    creation of CNT membranes for reverse osmosis applications.~\cite{sun_creating_2000}
    \item Molecular dynamics simulations have predicted and experiments have confirmed that
    CNTs are extremely fast transporters of water.~\cite{hummer_water_2001,joseph_why_2008}
    \item MD simulations by Kalra \textit{et al.} predicted water transport at rates comparable
    to the transmembrane protein aquaporin-1.~\cite{kalra_osmotic_2003}
    \item Experiments have exhibited water transport rates orders of magnitude higher than 
    predictions by continuum modeling.~\cite{holt_fast_2006}
    \item This enhancement has been explained as a result of nearly friction-less single-file water
    flow through the hydrophobic CNT interiors.~\cite{kalra_osmotic_2003}
    \item Functionalization of the CNT pore entrances can potentially lead to highly selective separations.~\cite{fornasiero_ion_2008}
	\item Practically, dispersing and aligning CNTs into a polymer matrix is extremely
    difficult because they tend to agglomerate due to van der Waals forces.~\cite{sahoo_polymer_2010}
    \item Some work has been to functionalize the carbon nanotubes in order to better incorporate them,
    however facile incorporation at scale remains a challenge yet to be overcome.~\cite{kim_polysulfone_2007}
  \end{itemize}

  \textit{Zeolites}: Zeolite membranes offer the potential for permeabilities comparable to ultrafiltration, 
  with selectivities as good as NF and RO.~\cite{hoek_nanotechnology-based_2014}
  \begin{itemize}  
    \item Zeolites have highly uniform nm-sized crystalline structures with cage-like
    cavities that allow movement and trapping of small solutes. 
    \item The crystalline frameworks are typically formed by networks of silicon 
    and aluminum each attached to 4 oxygen atoms in a tetrahedral arrangement.~\cite{auerbach_handbook_2003}
    \item One can replace the silicon and aluminum atoms via ion exchange in order
    to control the size of the cavities and hence its molecular-seiving properties.~\cite{li_novel_2007}
    \item Zeolite-coated ceramic membranes have shown the most promise for RO applications
    because of their mechanical stability under high applied pressure and chemical
    stability for withstanding disinfectants.~\cite{pendergast_review_2011}
    \item A number of studies have tested the permeability and sodium salt rejection
    of various zeolite membranes, however their high permeabilities typcially come
    at the cost of low selectivities and vice versa.~\cite{li_desalination_2004,duke_seawater_2009}
    \item Zeolites are also prone to defects in the crystalline structure which 
    would require substantial modification to the traditional synthesis procedures
    in order to fix.~\cite{kumakiri_application_2000}
  \end{itemize}

  \textit{Metal organic frameworks}: Finally, metal organic frameworks (MOFs) have been
  recently introduced to the field of aqueous separations.
  \begin{itemize}
    \item MOFs are highly porous and flexible crystalline structures made of metal
    clusters coordinated to organic ligands.~\cite{furukawa_chemistry_2013}
    \item There is a much wider breadth of research into MOFs as gas separation membranes
    because they tend to be unstable in water.~\cite{feng_construction_2013}
    \item But a new generation of water-stable MOFs is making it possible to use them
    in aqueous separations.~\cite{wang_ultrastable_2015}
    \item Purification of both aqueous and gaseous streams using MOFs are 
    typically achieved by selective adsorption or catalytic degradation.~\cite{li_metalorganic_2012}    
    \item MOFs have been applied towards the removal of a number of common water 
    contaminants including dyes, pesticides, endocrine disruptors, and pharmaceuticals.
    \cite{zhang_fabrication_2015,seo_adsorptive_2015,li_metalorganic_2015,hasan_adsorptive_2012}
    \item Although aqueous MOF separations are theorized to surpass the performance 
    of conventional membrane separations, it is still a budding field. The stability and
    reusability of the material must be guaranteed before they can be implemented
    at scale.~\cite{dias_towards_2015}
  \end{itemize}

  \section{Cross-linked Self Assembled Liquid Crystal Membranes for Selective Aqueous Separations}

  Amphiphilic molecules that self-assemble to form ordered nanostructures have shown
  potential as nanoporous membranes.
  \begin{itemize}
    \item People are particularly interested in self-assembled materials because they
    offer the potential of uniform-sized pores and tunable chemical and physical 
    properties.~\cite{werber_materials_2016}
    \item There are two major classes of self-assembled amphiphilic materials: block 
    copolymers and liquid crystals.
  \end{itemize}

  Block copolymers contain more than one distinct repeating unit which facilitates 
  the formation of self-assembled structures based on differences in the subunit 
  amphiphilicity.~\cite{jackson_nanoporous_2010} 
  \begin{itemize}
%    \item The varying amphiphilic character of block copolymers, polymers with more than
%    one distinct repeating unit, allows them to form self-assembled structures.~\cite{jackson_nanoporous_2010} 
    \item Block copolymer membranes typically have pores on the order of 5-50 nm which
    makes them useful for UF applications, however there is progress being made towards
    smaller pores in the NF regime.~\cite{gu_tailoring_2015}
  \end{itemize}

  Under the right conditions, the shape and amphiphilic character of liquid 
  crystal monomers drives their self-assembling into ordered nanostructures with
  uniform-sized pores on the order of 1 nm in size.~\cite{gin_polymerized_2008}
  \begin{itemize}
    \item They have been used as membranes before.~\cite{gin_polymerized_2001}
    \item As shown in Figure TBD, monomer 1 will form the inverted hexagonal phase 
    (H\textsubscript{II})~\cite{smith_ordered_1997} while monomer 2 will form the 
    bicontinuous cubic phase (Q\textsubscript{I})~\cite{carter_glycerol-based_2012}
    \item The tails have vinyl groups which can be cross-linked for mechanical strength.
    \item In this work we attempt to understand how one can design LLC membranes as 
    alternative nanostructured materials for highly selective aqueous separations.
    \item We chose to study liquid crystals because they are already suitable as NF 
    membranes and there has been sufficient experimental characterization.~\cite{gin_polymerized_2001,feng_scalable_2014,feng_thin_2016}
  \end{itemize}

  %BJC: figure TBD here showing each monomer and the phase they form

  \subsection{The H\textsubscript{II} Phase}

  The H\textsubscript{II} phase is characterized by hexagonally packed, uniform-sized
  and straight pores.~\cite{smith_ordered_1997}
  \begin{itemize}
    \item The hydrophilic head groups aggregate in the pore centers to create
    cylindrical aqueous channels. 
    \item These pore channels are lined with the chemical functionality of the LLC monomers
    and have the potential to interact with solutes in a chemically-dependent manner.
    \item The inhomogeneity of the pores suggests that the effective pore size may depend
    on the solute. Both the size, shape and chemical functionality will influence
    how a solute partitions within the membrane. %BJC2: need to rework this sentence.
    \item In theory, this pore topology and geometry is ideal for high flux and 
    highly selective separations.
  \end{itemize}

  Unfortunately, it is a difficult task to align the self-assembled hexagonal
  mesophases into continuous pores that traverse the thickness of the membrane.
  \begin{itemize}
    \item Although the membranes have shown high experimental selectivity, their 
    flux has been very low due to misalignment of the hexagonal mesophases.~\cite{zhou_supported_2005}
  \end{itemize}

%  Recently, Feng et al. has come up with two methods which can be used to 
%  macroscopically align the hexagonal mesophases. 
  Considerable recent efforts have made progress towards the macroscopic
  alignment of the hexagonal mesophases.
  \begin{itemize}
    \item Feng \textit{et al.} leveraged the magnetic anisotropy of the columnar
    mesophases in order to control their alignment with a magnetic field.~\cite{feng_scalable_2014}
    \item This study showed an 85-fold increase in ionic conductivity in the aligned
    versus unaligned state, demonstrating a huge increase in feasibility as an industrial membrane.
    \item Feng \textit{et al.} also used an approach called soft confinement which
    takes advantage of the hexagonal columns preference towards anchoring
    perpendicular to either a PDMS or glass substrate.~\cite{feng_thin_2016}
    \item Finally, Feng \textit{et al.} designed a third technique which uses a structure
    directing molecule in order to template the assembly of a fatty acid 
    molecule into ordered columnar phases.~\cite{feng_highly_2017}
%    \item Aligned hexagonal phase membranes exhibited an 85 fold increase in 
%    ionic conductivity.~\cite{feng_scalable_2014}
  \end{itemize}

  \subsection{The Q\textsubscript{I} Phase}

  The bicontinuous cubic, or Q\textsubscript{I} phase, is a class 
  of nanostructured phases characterized by a network of 3 dimensionally
  interconnected pores.~\cite{hyde_bicontinuous_1996}
  \begin{itemize}
    \item Aside from its tortuous pore architecture, it shares the 
    uniform size and complex topological features of the H\textsubscript{II} phase
    which lends itself to highly selective separations.
    \item Although water and solutes must follow a longer path in order to pass
    through Q\textsubscript{I} membranes, they are more practical to synthesize
    because the mesophases do not require alignment.~\cite{zhou_new_2007}
  \end{itemize}

  The space group of the Q\textsubscript{I} phase configuration formed by
  monomer 1 is uncertain and thought to be either Pn3m or Ia3d.  %BJC: will have accompanying figure
  \begin{itemize}
    \item 6 Q phase architectures have been identified in small molecule
    amphiphile systems.~\cite{mariani_cubic_1988}
    \item Of these, only 4 are consistent with diffraction data generated
    by the gemini surfactant used here.~\cite{pindzola_cross-linked_2003}
    \item Q$^{230}$ (Ia3d), Q$^{224}$ (Pn3m), Q$^{229}$ (Im3m) or Q$^{212}$ (P4\textsubscript{3}32) phase of type I configuration
    \item The presence of $1 / \sqrt{6}$ and $1 / \sqrt{8}$ peaks rules out the
    Q$^{227}$ (Fd3m) and Q$^{223}$ (Pm3n) configurations.
    \item The Pn3m and Ia3d architectures are the most common~\cite{mariani_cubic_1988,wiesenauer_nanoporous_2012}
    \item The most likely phase is the type I Q$^{230}$ (Ia3d) because it is the
    most common phase observed between lamellar and hexagonal phases and the
    monomeric alkyltrimethylammonium salts used to synthesize the gemini LCs
    exhibit clear Ia3d symmetry.
  \end{itemize}

  Due to its more facile synthesis, there has been significantly more development
  of Q\textsubscript{I} phase membranes.
  \begin{itemize}
    \item The first generation of Q\textsubscript{I} membranes was made by 
    the self-assembly of a gemini phosphonium monomer in water.~\cite{pindzola_cross-linked_2003}
    \item Hatakeyama \textit{et al.} improved the industrial viability of Q\textsubscript{I} 
    phase membranes by using a gemini ammonium monomer which is both easier and cheaper
    to synthesize.~\cite{hatakeyama_nanoporous_2010}
    \item Free standing films of both first and second generation Q\textsubscript{I} phase 
    membranes cannot withstand high pressure. Because solution casting was ineffective, Zhou
    et. al used a hot-pressing method to make mechanically strong membranes which involves 
    heating the initial monomer mixture to 70\degree C and pressing it with 12 tons of force
    into a thick microporous, hydrophilic polymer support.~\cite{zhou_new_2007}
    \item The most recent generation of Q\textsubscript{I} membranes uses an imidazolium-based
    gemini LLC monomer which is capable of being solution cast into defect-free thin films on
    porous supports. This improvement resulted in a ten-fold increase in flux while retaining
    selectivities similar to earlier generations.~\cite{carter_glycerol-based_2012}
  \end{itemize}

  %BJC: figure with different generation Q1 monomers

  State of the art Q\textsubscript{I} membranes offer selectivities that
  are competitive with existing commercial technologies.
  \begin{itemize}
    \item When separating organic solutes from NaCl, Q\textsubscript{I}-phase
    membrane filtration experiments have shown selectivity 2--3 times higher than
    commercial RO and 6--12 times higher than commercial NF membranes
    \cite{dischinger_application_2017}.
    \item Water permeability is higher than commercial RO membranes but
    less than commercial NF. There is work being done to reduce the
    thickness of the selective layer in order to increase permeability. 
  \end{itemize}

  %\section{Atomistic Molecular Simulation of LLC Membranes}
  \section{Modeling Membrane Separations}

  %BJC2: taken from stochastic transport paper.
  Mathematical descriptions of transport in complex separations membranes are a
  powerful way to understand mechanisms and formulate design principles.
  \cite{vinh-thang_predictive_2013,geens_transport_2006,darvishmanesh_mass_2016}
  The complexity of a well-fit model generally parallels the complexity of the
  transport mechanism being studied, as well as the transport information the
  model conveys. In dense homogeneous membranes, the solution-diffusion model
  can extract diffusion and partition coefficients and has successfully
  predicted solute transport rates.~\cite{wijmans_solution-diffusion_1995}
  Analogously, pore-flow models yield predictions of diffusion coefficients and
  solute transport rates in nanoporous membranes.~\cite{paul_diffusive_1974}
  Modern single particle tracking approaches have taken researchers beyond
  continuum modeling allowing them to characterize complex diffusive
  behavior.~\cite{manzo_review_2015} At the molecular level, one can use
  molecular dynamics (MD) simulations to study both single particle dynamics
  and bulk transport properties with atomic-level
  insight.~\cite{coscia_chemically_2019,maginn_best_2018} All of these
  approaches facilitate generation of hypotheses about the molecular origins of
  separations by attempting to give a more intuitive understanding of how
  solutes move as a function of their environment, in turn suggesting
  experiments that could be performed.

  \subsection{Macroscopic Treatments}

  The most widely used approaches to modeling transport in membranes of all types
  are based on macroscopic theory. In this section we will discuss the two prevailing
  theories: the pore flow model and the solution-diffusion model.

  \noindent Membrane permeation is the result of a chemical potential gradient, $\frac{d\mu_i}{dx}$.
  \begin{itemize}
	  \item The flux of component $i$ is proportional to this gradient:
	  \begin{equation}
	    J_i = -L_i \frac{d\mu_i}{dx}
	  \end{equation}
	  where $L_i$ is a coefficient of proportionality.~\cite{wijmans_solution-diffusion_1995}
	  \item $L_i$ is an intrinsic materials property dependent on ... % Wijman's doesn't say. I don't want to speculate. Get book he referenced from library
      %MRS2: mention what L depends on molecularly (even if uncalculatable)
	  \item Chemical potential gradients can be induced by concentration, pressure, 
	  temperature and electromotive forces.
	  \item We will limit our discussion to concentration and pressure gradients because 
	  these are primarily forces used for separations with the types of membranes we 
	  will be studying.
  \end{itemize}

  %MRS: is there are reason you are talkling about just water flux here?
  % This follows the discussion of werber_material_2016
  %BJC2: add solute flux
  For pressure driven membrane processes, water flux through both porous and 
  amorphous membranes can be described by
  \begin{equation}
  J_w = A(\Delta P - \Delta \pi_m)
  \end{equation}
  where $J_w$ is volumetric water flux, $A$ is the water permeability coefficient,
  $\Delta P$ is the applied hydraulic pressure and $\Delta \pi_m$ is the 
  trans-membrane osmotic pressure difference. Estimates of the water permeability 
  coefficient and the solute flux are dependent on how one models flow through the 
  membrane.

  %MRS: be clearer on presenting the context for this theory.  Will your findings be in contradiction?
  %MRS: is this theory inadequate?  Will it be tested?
  In porous membrane architectures, water and solute flux are modeled as laminar flow through 
  cylindrical pores.
  \begin{itemize}
    \item The derivation follows Hagen-Poiseuille flow with inclusion of morphological
    characteristics:~\cite{baker_membrane_2012}
    \begin{equation}
      A = \frac{\varepsilon r_p^2}{8\mu\delta_m}
    \end{equation}
    where $\varepsilon$ is the surface porosity, $r_p$ the pore radius, $\delta_m$ the
    membrane thickness, and $\mu$ the solution viscosity.
    \item Solute flux, $J_s$ is related to $J_w$ but scaled by the empirically measured
    permeate concentration:
    \begin{equation}
    J_s = c_p J_w
    \end{equation}
    \item Analytical expressions for solute flux, such as the Donnan Steric Pore Model (DSPM), 
    are typically quite complicated because they attempt to incorporate more system features
    leading to less assumptions.~\cite{bowen_modelling_2002}
  \end{itemize}

  Solute rejection is the quantity typically used to characterize the performance of a 
  filtration membrane.
  \begin{itemize}
    \item Under the pore flow model, solutes with radii smaller than the pore will not 
    pass through the membrane. 
    \item A membrane with uniform-sized pores smaller than the solute being targeted 
    for separation should have a solute flux of 0. However, in practice, most membranes
    have a distribution of pore sizes, some large enough to allow solutes to pass.
    \item Therefore, separation performance is typically characterized in terms of solute
    rejection, $R$. 
    \begin{equation}
      R = 1 - \frac{c_p}{c_f}
    \end{equation}
    where $c_p$ and $c_f$ are the concentrations of the rejected species in
    the permeate and feed respectively.
    \item The DSPM has been used to directly predict rejection.~\cite{hatakeyama_water_2011,bowen_modelling_2002}
  \end{itemize}

  % BJC: mention does not enter Brownian regime
  In amorphous membranes, water and solute flux are modeled using the solution-diffusion 
  model.
  \begin{itemize}
    \item It is hypothesized that water and solute molecules partition into the
    membrane and diffuse across the polymer matrix due to a chemical potential
    gradient and then desorb into the permeate.
    \item The product of a component's solubility and diffusivity is commonly
    referred to as its permeability, $P$.~\cite{wijmans_solution-diffusion_1995}
    \item $P$ is an intrinsic material property and is thus independent of membrane
    thickness.  % BJC: is intensive a more appropriate word than intrinsic?
    \item The water permeability coefficient, $A$, \textit{is} dependent on thickness, $\delta_m$,
    and can be formulated in terms of the pure water permeability, $P_w$.
    \item \begin{equation}
    A = \frac{P_wV_w}{\delta_mR_gT}
    \end{equation}
    where $V_w$ is molar volume of water, $R_g$ is the universal gas constant and $T$ is
    absolute temperature.
    \item Solute flux is modeled as:
    \begin{equation}
    J_s = \frac{P_s}{\delta_m}\Delta c_m = B\Delta c_m
    \end{equation}
    where $B$ is a solute permeability coefficient and $\Delta c_m$ is the concentration
    difference across the selective membrane layer.~\cite{geise_fundamental_2014}
    \item Under these assumptions, rejection can be reformulated as
    \begin{equation}
    R = \frac{A}{A + B}
    \end{equation}
    which approaches unity as the solute flux tends towards zero.~\cite{van_der_bruggen_review_2003}
  \end{itemize}

  These equations provide a somewhat limited perspective on the molecular details
  of transport.
  \begin{itemize}
    \item The pore flow model neglects molecular level interactions with the membrane
    and assumes all transport occurs within the cylindrical pore region. In this work, we
    will show that solutes are not confined to the ``pore region" and that motion is 
    heavily influenced by interactions with membrane.
    \item The solution-diffusion model assumes that solutes undergo Brownian motion. As will
    be discussed, solutes appear to undergo subdiffusive behavior on short
    timescales. 
    \item Therefore, we may need a more descriptive theory in order
    to appropriately address questions of membrane design.
%    \item The parameters are experimental observables which allow one to 
%    speculate on the atomic-level which cause them.
  \end{itemize}

  \subsection{Atomistic Molecular Simulation of LLC Membranes}

  %BJC: following two paragraphs taken from first paper.
  %BJC2: Could probably reorganize the sentences of these next two paragraphs 
  % to flow better with rest of intro.
  %BJC2: These next two paragraphs might function better in previous section.
%  Our current understanding of the molecular details of LLC polymer membranes'
%  nanostructure is not sufficient to be able to precisely design them for
%  specific separations. Dischinger \textit{et al.}~attempted to use an empirical model
%  that correlates the physiochemical properties of the counterion used in a
%  Q\textsubscript{I}-phase LLC membrane to solute rejection\cite{dischinger_effect_2017}.
%  Although their model showed some qualitative agreement with experiment, the
%  quality of fit of their model was limited due to complex solute-membrane
%  interactions that could not easily be modeled. Additionally, they observed
%  an unexpected discrepancy in the relationship between uncharged solute
%  rejection and water permeability, which will require a more in-depth knowledge of
%  the difference between solute and solvent transport.
%
%  Over the past 20 years, H\textsubscript{II}-phase LLC polymer membrane
%  studies have been limited primarily to the Na-GA3C11 monomer with some
%  characterization done after minor structural modifications. For example, Resel
%  \textit{et al.}~varied the length of the monomer tails and the counterion used and
%  observed its effect on pore spacing \cite{resel_structural_2000}. In a later
%  study of rejection performance, it was shown that membranes formed by
%  cross-linked Na-GA3C11 in the H\textsubscript{II} phase cannot separate solutes
%  less than 1.2 nm in diameter because the pores are too large.
%  %MRS: this above is interesting, since that seems somewhat inconsistent with the 
%  \cite{zhou_supported_2005} We do not yet understand how to controllably reduce
%  the effective pore size or how to tune the chemical environment in the
%  nanopores of this or related materials for small molecule separations. The only
%  source of predictive modeling for LLC systems have been macroscopic models that
%  likely do not adequately describe transport at these length scales
%  \cite{hatakeyama_water_2011}. Modeling with molecular detail could provide
%  sufficient information about the mechanisms and chemical features to better
%  inform experimental design of similar nanostructured membranes.

  A molecular-level understanding of the relationship between monomer structure
  and solute transport can help provide guidelines to reduce the
  large chemical space available to design monomers for creation of
  separation-specific membranes.
  \begin{itemize}
    \item Atomistic MD simulations can provide the required level of detail 
    (Figure \ref{fig:detailed_pore}), assuming the force fields are sufficiently accurate. 
    \item With such an atomistic model, we can directly observe
    molecular-level solute transport and suggest governing mechanisms. 
    \item We can also observe how the choice of head group and counterions interacts
    with solutes of interest. 
  \end{itemize}

  \begin{figure}[!htb]
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{cartoon_pore.pdf}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{detailed_pore.pdf}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure} 
    \caption{(a) Previous understanding of the LLC pores are essentially speculations 
    based on limited chemical and experimental data. (b) We use detailed molecular 
    modeling in this paper in order to appropriately model the pore's complex architecture,
    which is crucial to understanding the mechanism of solute transport. In both 
    pictures, the head group region is colored blue and the tail region is colored cyan.}~\label{fig:detail}
  \end{figure}

  Although the Q\textsubscript{I} phase shows the most promise for
  practical applications, the focus of this work will be on the H\textsubscript{II}
  phase.
  \begin{itemize}
    \item The H\textsubscript{II} phase is an easier to geometry to model
    and analyze.
    \item We also have detailed structural data which is necessary for validating
    our model.
    \item All of the analysis techniques we use can be equally adapted to
    the more complicated Q\textsubscript{I} phase geometry.
  \end{itemize}

  There are few molecular simulation studies which study structure and transport
  in LLC systems.
  \begin{itemize}
    \item Mondal \textit{et al.} used MD simulations to study the self assembly of gemini 
    surfactants and observed the formation of H\textsubscript{II} and Q\textsubscript{I}
    phases depending on water content.\cite{mondal_self-assembly_2013}
    \item Mantha and Yethiraj as well as Roy \textit{et al.} studied the dynamics of confined 
    water in these systems and showed orders of magnitude differences in their motion
    dependent on system geometry.~\cite{mantha_dynamics_2016,roy_water_2016}
    \item Jackson \textit{et al.} as well as Mantha \textit{et al.} have combined experiment with simulation
    in order to show how the choice of monomer head group counterion regulate water 
    dynamics.~\cite{jackson_ion-specific_2018,mantha_counterion-regulated_2018}
    \item Sakamoto \textit{et al}. and Nada \textit{et al}. observed the dynamics of water molecules
    and ions and hypothesized transport mechanisms in a simplified model of an LLC
    nanopore.~\cite{sakamoto_development_2018,nada_transport_2020}
  \end{itemize}

  Because there is relatively sparse coverage of these types of simulation systems in 
  the literature, we have built a detailed molecular model from the ground up. There 
  are four primary research questions that we will address in this work.
  \begin{enumerate}
    \item What is the nanoscopic structure?
    	\begin{itemize}
    	  \item Before we can try to understand the molecular mechanisms of solute
    	  transport, we need to ensure that we model the chemical environment
    	  within the nanopores in a way that is consistent with experiment.
    	  \item We use experimental structural data in order to validate our model.
    	\end{itemize}
    \item Which solute-membrane interactions have the greatest influence on transport rates?
    	\begin{itemize}
    	  \item After gaining a detailed picture of the nanopore structure, we can 
    	  feel confident that solutes in this system will experience the same 
    	  interactions which are present in a real system.
    	  \item We create independent systems for each of 20 small polar solutes 
    	  and observe transport mechanisms whose dominance is dependent on
    	  solute chemical functionality.
    	  \item We characterize three different trapping mechanisms which lead
    	  to subdiffusive transport behavior.
    	\end{itemize}
    %MRS: maybe talk more about the stochastic modeling itself; for example, could phrase it in terms of
    %how you can test mechanistic models to see if they are consistent with the timeseries, and then how those models
    %can be further extrapolated to physical observables even if the MD simulations cannot. 
    %MRS: this MIGHT include using the effective diffusion in a solution/diffusion model.
    %\item Can we estimate experimentally-relevant macroscopic properties?
    %BJC2: modified thesis
    \item Can we formulate mechanistic models whose dynamical properties are consistent with our
    molecular simulations and can be used to extrapolate macroscopic behavior which MD simulations cannot.
    	\begin{itemize}
    	  \item Using our qualitative understanding of the dominant trapping mechanisms,
    	  we develop stochastic time series models which we can use to mimic solute dynamic
    	  behavior on time scales orders of magnitude longer than our simulations. 
    	  \item We attempt to reproduce both quantitative and qualitative solute trajectory
    	  behavior on MD simulation-length timescales.
    	  \item We then show how we can use our most promsiing model in order to connect 
    	  microscopic transport to macroscopic flux and selectivity.
    	\end{itemize}

    \item How can time series data directly inform us about transport mechanisms?
    	\begin{itemize}
    	  \item Although our stochastic models show great promise, their development 
    	  require some qualitative and quantitative understanding of dominant transport
    	  mechanisms.
    	  \item In the final chapter, we use the infinite hidden Markov model in order
    	  to automatically detect and parameterize a unknown number of hidden dynamical modes
    	  exhibited by solute time series. 
    	  \item This more flexible approach allows us to both infer mechanisms based 
    	  on differences in dynamical behavior and generate stochastic trajectory realizations
    	  which we can use to predict flux and selectivity. 
    	\end{itemize}
  \end{enumerate}

  The next four chapters offer answers to each of the above questions and are adapted from 
  my first author manuscripts at various stages in the publication process as described
  below. The introduction to each adaptation is condensed in order to avoid redundancy and 
  all other sections are left intact. The supporting information documents for each manuscript are
  given in the appendices. \\

  %BJC2: Should I just put title with reference? Or full citation?
  \noindent\textit{Chapter 2}: adapted from manuscript entitled ``Understanding the 
  Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes." 
  published in the Journal of Physical Chemistry B.~\cite{coscia_understanding_2019} \\

  \noindent\textit{Chapter 3}: adapted from manuscript entitled ``Chemically Selective 
  Transport in a Cross-Linked H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane."
  published in the Journal of Physical Chemistry B.~\cite{coscia_chemically_2019} \\

  \noindent\textit{Chapter 4}: adapted from a manuscript in review with Physical Review E, entitled 
  ``Capturing Subdiffusive Solute Dynamics and Predicting Selectivity in Nanoscale Pores
  with Time Series Modeling". The final manuscript may change based on reviewer feedback. \\

  \noindent\textit{Chapter 5}: adapted from the draft of a manuscript in preparation with the working
  title ``Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series 
  of Solute Trajectories in Nanostructured Membranes." The content will change as the
  draft is finalized and after reviewer feedback.

%  Chapter 2 is adapted 
%  \begin{itemize}
%    %BJC2: Should I just put title with reference? Or full citation?
%    \item Chapter 2 is adapted from Coscia, B. J.; Yelk, J.; Glaser, M. A.; Gin, D. L.; Feng, X.; Shirts,
%    M. R. Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer
%    Membranes. J. Phys. Chem. B 2019, 123, 289-309.
%    \item Chapter 3 is adapted Coscia and Shirts. Chemically Selective Transport in a Cross-Linked 
%    H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane. J. Phys. Chem. B 2019, 123, 6314-6330. 
%    \item Chapter 4 is adapted from a manuscript in review with Physical Review E, entitled 
%    "Capturing Subdiffusive Solute Dynamics and Predicting Selectivity in Nanoscale Pores
%    with Time Series Modeling". The final manuscript may change based on reviewer feedback.
%    \item Chapter 5 is adapted from the draft of a manuscript in preparation with the working
%    title "Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series 
%    of Solute Trajectories in Nanostructured Membranes."
%  \end{itemize}