### 2020-04-09 Ben_Manuscripts/thesis/thesis.tex
Beginning of supporting info and figures. Draft updates, abstract for thesis

\documentclass[defaultstyle,11pt]{thesis}

\usepackage{amssymb}		% to get all AMS symbols
\usepackage{graphicx}		% to insert figures
%\usepackage{hyperref}		% PDF hyperreferences??
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{textcomp} % gets rid of \perthousand and \micro warnings that come from gensymb
\usepackage{gensymb}
\usepackage{subcaption}
\usepackage{tcolorbox}
\usepackage[export]{adjustbox}
\usepackage{listings}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{diagbox}
\usepackage{array}

\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\newcolumntype{C}[1]{>{\PreserveBackslash\centering}p{#1}}
\newcolumntype{R}[1]{>{\PreserveBackslash\raggedleft}p{#1}}
\newcolumntype{L}[1]{>{\PreserveBackslash\raggedright}p{#1}}

\DeclareUnicodeCharacter{03C9}{$\omega$}
\DeclareUnicodeCharacter{03C0}{$\pi$}

\definecolor{amber}{rgb}{0.99, 0.76, 0.0} % golden poppy

\title{Understanding Structure and Selectivity in Nanostructured Lyotropic Liquid Crystal Membranes using Molecular Simulations}

\author{Benjamin Joseph}{Coscia}

\otherdegrees{B.A., University of Connecticut, 2014 \\
	          M.S., University of Colorado Boulder, 2017}

\Degree{Doctor of Philosophy}		                %  #1 {long descr.}
	   {Ph.D., Chemical Engineering}		%  #2 {short descr.}

\dept{Department of}			                %  #1 {designation}
	 {Chemical and Biological Engineering}		%  #2 {name}

\advisor{Prof.}				        %  #1 {title}
	    {Michael Shirts}			%  #2 {name}

\reader{Prof. Richard Noble}		%  2nd person to sign thesis
\readerThree{Prof. Douglas Gin}		%  3rd person to sign thesis
\readerFour{Prof. Matthew Glaser}
\readerFive{Prof. Robert Davis}

\abstract{

  % \OnePageChapter	% use if abstract is less than one page

  The ability to perform highly selective aqueous separations at low cost
  is beneficial to a variety of chemical industries. Periodic, nanostructured 
  porous polymer membranes made from the cross-linked inverted hexagonal phase
  of self-assembled lyotropic liquid crystals (LLCs) are a promising class of 
  materials for selective separations. In this thesis, we use atomistic molecular 
  modeling in order investigate an experimentally characterized LLC polymer 
  membrane.

  First, we gain a deep understanding of nanoscopic pore structure in order 
  to make sure we can appropriately model the chemical environment in the pores.
  In particular, we compare simulated X-ray diffraction (XRD) patterns with 
  experimental XRD data to quantify and understand the major structural
  features that we must incorporate into our model. We explore the composition 
  and structure of the nanopores and reveal that there exists a composition gradient
  rather than an abrupt partition between the hydrophilic and hydrophobic regions.
  A caveat is that the time scales of the dynamics are extremely long for this system,
  resulting in simulated structures that appear too ordered, thus requiring
  careful examination of the metastable states observed in order to draw any
  conclusions.

  We use our carefully constructed molecular model in order to observe transport of
  water, sodium ions and 20 small polar solutes within the pores of a lyotropic 
  liquid crystal (LLC) membrane using atomistic molecular simulations. We find 
  that the transport of a species is dependent not only on molecular size, but on
  chemical functionality as well. We observe that all solutes perform intermittent 
  hops between lengthy periods of entrapment. Three different trapping mechanisms 
  are responsible for this behavior. First, solutes that drift out of the pore can
  become entangled among the dense monomer tails. Second, solutes can donate 
  hydrogen bonds to the monomer head groups. Third, solutes can coordinate with 
  sodium counter ions. The degree to which a solute is affected by each mechanism 
  is dependent on the chemical functionality of the solute.

  The remainder of this thesis focuses on developing stochastic models meant to 
  generate realizations of solute trajectories with similar qualitative and
  quantitative dynamics to our MD trajectories solute trajectories. Mathematically 
  modeling complex transport phenomena at the molecular level can be a powerful tool
  for connecting microscopic chemically-dependent solute motions that do not
  follow simple diffusive behavior with macroscopic membrane performances. 
  We develop three different stochastic time series models, parameterized by long 
  molecular dynamics (MD) simulation trajectories.

  First, using anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time
  random walk. From our MD simulations, we parameterize the distribution of
  dwell times, hop lengths between dwells and correlation between hops.

  Next, we generalize Markov state models, treating the configurational states of 
  the system as a Markov process where each state has distinct transport properties. 
  For each state and transition between states, we parameterize the distribution 
  and temporal correlation structure of positional fluctuations as a means of 
  characterization and to allow us to predict solute MSDs.

  Building on what we learned from the first two models, we develop a third approach,
  called the infinite hidden Markov model, which is capable of automatically 
  distinguishing an unknown number of hidden dynamical modes in multidimensional
  solute time series. This model offers a flexibile approach to modeling solute
  motion and helps us to uncover mechanisms with minimal human effort.
  }

\dedication[Dedication]{	% NEVER use \OnePageChapter here.
	To all of the fluffy kitties.
	}

%  Here's where you acknowledge folks who helped. But keep it short, i.e., no more than one page,
%  as required by the Grad School Specifications.
\acknowledgements{	
  \OnePageChapter	% *MUST* BE ONLY ONE PAGE!

  I would like to acknowledge and thank a number of people whose support and
  guidance has not gone unnoticed. First, I thank my advisor, Michael Shirts, 
  who has been instrumental to my growth as a competent and confident researcher,
  pushing me to do high quality science. I thank the entire Shirts group for 
  their feedback and encouragement as well as for making our lab an enjoyable to work.
  I thank Richard Noble and Douglas Gin for their guidance on past and present 
  experimental research on LLC membranes. I also thank Michael McGrath, Gregory 
  Dwulet and Sarah Dischinger, members of the Gin and Noble groups, for providing
  invaluable experimental perspective. I thank Matthew Glaser for his 
  contributions to my understanding of liquid crystal systems. Our first paper 
  would not have been complete without his insight. I thank Joseph Yelk for his 
  effort creating the X-ray diffraction code on which our first paper relied.
  I thank Christopher Calderon for his guidance on applying Bayesian non-parametric
  techniques to solute time series.
%  I would like to acknowledge my financial support: The ACS Petroleum Research Fund
%  grant \#59814-ND7 and the Graduate Assistance in Areas of National Need (GAANN) 
%  fellowship which is funded by the U.S. Department of Education. 
%
%  % BJC: This acknowledgement is how they want to be cited in papers. Could I
%  % shorten it for this?
%  Finally, I would like to acknowledge my high performance computing resources:
%  Molecular simulations were performed using the Extreme Science and
%  Engineering Discovery Environment (XSEDE), which is supported by National
%  Science Foundation grant number ACI-1548562. Specifically, it used the Bridges
%  system, which is supported by NSF award number ACI-1445606, at the Pittsburgh
%  Supercomputing Center (PSC). This work also utilized the RMACC Summit supercomputer,
%  which is supported by the National Science Foundation (awards ACI-1532235 and
%  ACI-1532236), the University of Colorado Boulder, and Colorado State
%  University. The Summit supercomputer is a joint effort of the University of
%  Colorado Boulder and Colorado State University.

  }

% BJC: The list of figures and tables is obnoxious because the captions are so long.
% BJC: Can I leave them out? (would just uncomment \emptyLoF and \emptyLoT)

% \IRBprotocol{E927F29.001X}	% optional!

% \ToCisShort	% use this only for 1-page Table of Contents

% \LoFisShort	% use this only for 1-page Table of Figures
% \emptyLoF	% use this if there is no List of Figures

% \LoTisShort	% use this only for 1-page Table of Tables
% \emptyLoT	% use this if there is no List of Tables

\begin{document}

\graphicspath{{../transport/figures/}{../structure_paper/figures/}{../transport/supporting_figures/}{../stochastic_transport/figures/}{../stochastic_transport/supporting_figures/}}

\input macros.tex
\input intro.tex

%\input structure.tex
%\input structure_si.tex

%\input transport.tex
%\input transport_si.tex

%\input stochastic_transport.tex

%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{ieeetr}	% or "siam", or "alpha", etc.
%\nocite{*}		% list all refs in database, cited or not
\bibliography{refs}		% Bib database in "refs.bib"

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
\appendix
%\input structure_si.tex
%\input transport_si.tex
%\input stochastic_transport_si.tex

\end{document}