### 2020-04-10 Ben_Manuscripts/thesis/thesis.tex
Shortened abstract

\documentclass[defaultstyle,11pt]{thesis}

\usepackage{amssymb}		% to get all AMS symbols
\usepackage{graphicx}		% to insert figures
%\usepackage{hyperref}		% PDF hyperreferences??
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{textcomp} % gets rid of \perthousand and \micro warnings that come from gensymb
\usepackage{gensymb}
\usepackage{subcaption}
\usepackage{tcolorbox}
\usepackage[export]{adjustbox}
\usepackage{listings}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{diagbox}
\usepackage{array}

\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\newcolumntype{C}[1]{>{\PreserveBackslash\centering}p{#1}}
\newcolumntype{R}[1]{>{\PreserveBackslash\raggedleft}p{#1}}
\newcolumntype{L}[1]{>{\PreserveBackslash\raggedright}p{#1}}

\DeclareUnicodeCharacter{03C9}{$\omega$}
\DeclareUnicodeCharacter{03C0}{$\pi$}

\definecolor{amber}{rgb}{0.99, 0.76, 0.0} % golden poppy

\title{Understanding Structure and Selectivity in Nanostructured Lyotropic Liquid Crystal Membranes using Molecular Simulations}

\author{Benjamin Joseph}{Coscia}

\otherdegrees{B.A., University of Connecticut, 2014 \\
	          M.S., University of Colorado Boulder, 2017}

\Degree{Doctor of Philosophy}		                %  #1 {long descr.}
	   {Ph.D., Chemical Engineering}		%  #2 {short descr.}

\dept{Department of}			                %  #1 {designation}
	 {Chemical and Biological Engineering}		%  #2 {name}

\advisor{Prof.}				        %  #1 {title}
	    {Michael Shirts}			%  #2 {name}

\reader{Prof. Richard Noble}		%  2nd person to sign thesis
\readerThree{Prof. Douglas Gin}		%  3rd person to sign thesis
\readerFour{Prof. Matthew Glaser}
\readerFive{Prof. Robert Davis}

\abstract{

  % \OnePageChapter	% use if abstract is less than one page

  Periodic, nanostructured porous polymer membranes made from the cross-linked
  inverted hexagonal phase of self-assembled lyotropic liquid crystals (LLCs) are
  a promising class of materials for selective separations. Our current understanding
  of these systems is not sufficient to design them for specific applications. In 
  this thesis, we use detailed atomistic molecular modeling in order to investigate 
  structure and transport of solutes within LLC polymer membrane.

  We begin by enhancing our understanding of the nanoscopic pore structure in order 
  to ensure we appropriately model the chemical environment in the pores. We use radial
  distribution functions and simulated X-ray diffraction patterns to quantify and 
  understand the major structural features of the model.

  We use our carefully constructed molecular model in order to observe transport of
  water, sodium ions and 20 small polar solutes within the pores of a lyotropic 
  liquid crystal (LLC) membrane using atomistic molecular simulations. We 
  observe that solutes perform intermittent hops between lengthy periods of 
  entrapment and describe how transport of a species is dependent not only on 
  molecular size, but on chemical functionality as well.

  Based on our qualitative understanding of solute transport, we focus the remainder
  of our efforts on the development of three different stochastic models which can be
  used to generate realizations of solute trajectories with qualitative and quantitative 
  dynamics similar to our MD solute trajectories. Mathematically modeling complex 
  transport phenomena at the molecular level can be a powerful tool for connecting 
  microscopic chemically-dependent solute motions that do not follow simple diffusive
  behavior with macroscopic membrane performance.

  First, using anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time
  random walk. We parameterize the distribution of dwell times, hop lengths between
  dwells and correlation between hops. Next, we generalize Markov state models, treating
  the configurational states of the system as a Markov process where each state has 
  distinct transport properties. Finally we implement a third approach,
  called the infinite hidden Markov model, which is capable of automatically 
  distinguishing an unknown number of hidden dynamical modes in multidimensional
  solute time series. This model offers a flexibile approach to modeling solute
  motion and helps us to uncover mechanisms with minimal human effort.
  }

\dedication[Dedication]{	% NEVER use \OnePageChapter here.
	To all of the fluffy kitties.
	}

%  Here's where you acknowledge folks who helped. But keep it short, i.e., no more than one page,
%  as required by the Grad School Specifications.
\acknowledgements{	
  \OnePageChapter	% *MUST* BE ONLY ONE PAGE!

  I would like to acknowledge and thank a number of people whose support and
  guidance has not gone unnoticed. First, I thank my advisor, Michael Shirts, 
  who has been instrumental to my growth as a competent and confident researcher,
  pushing me to do high quality science. I thank the entire Shirts group for 
  their feedback and encouragement as well as for making our lab an enjoyable to work.
  I thank Richard Noble and Douglas Gin for their guidance on past and present 
  experimental research on LLC membranes. I also thank Michael McGrath, Gregory 
  Dwulet and Sarah Dischinger, members of the Gin and Noble groups, for providing
  invaluable experimental perspective. I thank Matthew Glaser for his 
  contributions to my understanding of liquid crystal systems. Our first paper 
  would not have been complete without his insight. I thank Joseph Yelk for his 
  effort creating the X-ray diffraction code on which our first paper relied.
  I thank Christopher Calderon for his guidance on applying Bayesian non-parametric
  techniques to solute time series.
%  I would like to acknowledge my financial support: The ACS Petroleum Research Fund
%  grant \#59814-ND7 and the Graduate Assistance in Areas of National Need (GAANN) 
%  fellowship which is funded by the U.S. Department of Education. 
%
%  % BJC: This acknowledgement is how they want to be cited in papers. Could I
%  % shorten it for this?
%  Finally, I would like to acknowledge my high performance computing resources:
%  Molecular simulations were performed using the Extreme Science and
%  Engineering Discovery Environment (XSEDE), which is supported by National
%  Science Foundation grant number ACI-1548562. Specifically, it used the Bridges
%  system, which is supported by NSF award number ACI-1445606, at the Pittsburgh
%  Supercomputing Center (PSC). This work also utilized the RMACC Summit supercomputer,
%  which is supported by the National Science Foundation (awards ACI-1532235 and
%  ACI-1532236), the University of Colorado Boulder, and Colorado State
%  University. The Summit supercomputer is a joint effort of the University of
%  Colorado Boulder and Colorado State University.

  }

% BJC: The list of figures and tables is obnoxious because the captions are so long.
% BJC: Can I leave them out? (would just uncomment \emptyLoF and \emptyLoT)

% \IRBprotocol{E927F29.001X}	% optional!

% \ToCisShort	% use this only for 1-page Table of Contents

% \LoFisShort	% use this only for 1-page Table of Figures
% \emptyLoF	% use this if there is no List of Figures

% \LoTisShort	% use this only for 1-page Table of Tables
% \emptyLoT	% use this if there is no List of Tables

\begin{document}

\graphicspath{{../transport/figures/}{../structure_paper/figures/}{../transport/supporting_figures/}{../stochastic_transport/figures/}{../stochastic_transport/supporting_figures/}}

\input macros.tex
\input intro.tex

%\input structure.tex
%\input structure_si.tex

%\input transport.tex
%\input transport_si.tex

%\input stochastic_transport.tex

%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{ieeetr}	% or "siam", or "alpha", etc.
%\nocite{*}		% list all refs in database, cited or not
\bibliography{refs}		% Bib database in "refs.bib"

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
\appendix
%\input structure_si.tex
%\input transport_si.tex
%\input stochastic_transport_si.tex

\end{document}