### 2020-04-10 Ben_Manuscripts/hdphmm/Supporting_Information.tex
added missing file

\documentclass{article}

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath} % or simply amstext
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\newcommand{\angstrom}{\textup{\AA}}
\newcommand{\colormap}{jet}  % colorbar to use
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}
\usepackage{xr}

\externaldocument[M-]{Draft}

\title{Supporting Information: Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series 
       of Solute Trajectories in Nanostructured Membranes.}

\author{Benjamin J. Coscia, Christopher P. Calderon \and Michael R. Shirts}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thepage}{S\arabic{page}}
\renewcommand{\thetable}{S\arabic{table}}

\begin{document}

  \graphicspath{{./supporting_figures/}}
  \maketitle

  \section{Demonstration of IHMM Parameterization Procedure}\label{section:ihmm_procedure}

  Parameterization of trajectories by the IHMM is a multi-stepped procedure. In the figures
  that follow, we graphically illustrate the procedure which is described in detail in
  Section~\ref{M-method:IHMM} of the main text.

  \begin{enumerate}
  	\item Parameterize in $x$, $y$, $z$ coordinates with $x$, $y$ coordinates relative to the nearest
  	pore center (see Figure~\ref{fig:xyz_hmm}). 	
  	\item Convert x and y coordinates to radial coordinates, $r$, then fix the state sequence
  	and use the inference component of the IHMM in order to estimate the parameters in terms
  	of $r$ and $z$ (see Figure~\ref{fig:rz_hmm}). 	 	
  	\item Cluster the VAR parameters from the states found in all 24 trajectories. Reassign the state
  	sequence so that segments which belong to the same cluster are labeled the same across all solute
  	trajectories.
  	\item Subtract the mean of each segment, obtained before clustering, in order to zero out the
  	trajectory.
  	\item Fix clustered state sequence and infer parameters of each state, assuming a mean of zero
  	for all states (see Figure~\ref{fig:zeroed_hmm}). 
  \end{enumerate}

  \begin{figure}
    \centering
	\begin{subfigure}{0.8\textwidth}
		\includegraphics[width=\textwidth]{xyz_hmm.pdf}
		\caption{}\label{fig:xyz_hmm}
	\end{subfigure}  	
	\begin{subfigure}{0.8\textwidth}
		\includegraphics[width=\textwidth]{rz_hmm.pdf}
		\caption{}\label{fig:rz_hmm}
	\end{subfigure}  
	\begin{subfigure}{0.8\textwidth}
		\includegraphics[width=\textwidth]{zeroed_hmm.pdf}
		\caption{}\label{fig:zeroed_hmm}
	\end{subfigure}  
	\caption{IHMM parameterization is a multi-step procedure. In the above plots, distinct colors
	correspond to distinct state behavior. (a) First we parameterize the trajectories in terms of
	the $x$, $y$, $z$ solute center-of-mass coordinates with the $x$ and $y$ coordinates relative
	to the nearest pore center (b) Next, we convert the $x$ and $y$ 	
	coordinates to radial coordinates, $r$, then fix the state sequence and use the inference 
	component of the IHMM in order to estimate the parameters in terms of $r$ and $z$. Note that
	the segment colors in (b) are the same as (a) because we fixed the state sequence.
	(c) After clustering the VAR parameters and reassigning the state sequence so that segments 
	which belong to the same cluster are labeled the same across all solute trajectories, we subtract
	the mean of each segment, obtained before clustering, in order to zero out the
  	trajectory, and then fix the clustered state sequence to infer the parameters of each clustered 
  	state, assuming a mean of zero for all states. Note that the colors are no longer aligned with 
  	figures (a) and (b) because the states are relabeled in terms of their clusters. There are
  	also significantly less states in this single trajectory.
	}\label{fig:hmm_demo}
  \end{figure}

  \newpage

  \section{Parameter Convergence}\label{section:convergence}

  Given enough data, parameter estimates converge within only a few steps. In 
  Figure~\ref{fig:convergence}, we plot the entries of $A$ and $\Sigma$ as 
  a function of IHMM iteration. Some states are sampled far more frequently
  than others. High sampling frequently results in higher certainty converged
  parameter estimates.

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.7\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_4.pdf}
  \caption{15 total emissions}\label{fig:convergence_MET_low}
  \end{subfigure}
  \begin{subfigure}{0.7\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_21.pdf}
  \caption{1289 total emissions}\label{fig:convergence_MET_medium}
  \end{subfigure}
  \begin{subfigure}{0.7\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_15.pdf}
  \caption{24468 total emissions}\label{fig:convergence_MET_high}
  \end{subfigure}
  \caption{As the total number of emissions observed from each state
  increases, the uncertainty in the converged parameters estimates
  decreases.}\label{fig:fixed_state_convergence}
  \end{figure}

  \newpage

  \section{Choosing Prior Parameters}\label{section:prior_guesses}

  The states identified by the IHMM are heavily influenced by the Gaussian prior
  placed on $\mathbf{c}$ in Equation~\ref{M-eqn:var} of the main text. The 
  entries of $A$ and $\Sigma$ do not vary over a wide range, so the final
  parameters were relatively insensitive to the priors. In order to 
  maximally automate the IHMM procedure, we attempted to parameterize the
  prior on $\mathbf{c}$ in an intelligent way. The prior parameters should
  be chosen such that the mean level of each state lies within a region of 
  reasonable probability of the prior (see Figure~\ref{fig:prior_guesses}).
  In each dimension, we defined the prior mean to lie halfway between the 
  maximum and minimum of each trajectory dimension. To parameterize the 
  prior's variance in each dimension, we defined the maximum and minimum to 
  be 2 standard deviations from the prior mean. Although this approach has 
  worked quite well for the data in this work, it is important to check the 
  results to determine whether further adjustments to the prior might be needed.
  In Figure TBD, we show the result of a parameterization where the prior
  parameters of $c$ were poorly chosen. %TODO

  \begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{prior_guesses.pdf}
  \caption{The parameters of the prior on the mean vector, $c$, (black line), which
  represents the coordinates at which solute are trapped, should be chosen such
  that the mean levels of each state identified in the trajectory (blue line) lie within
  regions of the prior with reasonable probability. We chose the mean of the prior 
  as halfway between the maximum and minimum of each trajectory dimension. We chose 
  $\sigma$ of the prior by defining the maximum and minimum to be 2 standard deviations
  from the mean (dashed lines).}\label{fig:prior_guesses}
  \end{figure}

  \newpage

  \section{Agglomerative Clustering Versus Gaussian Mixture Models}\label{section:agglomerative}

  % BJC: will show Gaussian mixture model which has a cluster engulfing another cluster, so
  % it's tails are on both sides of the center cluster.

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.95\textwidth}
  \includegraphics[width=\textwidth]{bayesian_A.pdf}
  \caption{Bayesian Gaussian mixture model}\label{fig:bayesian_A}
  \end{subfigure}
  \begin{subfigure}{0.95\textwidth}
  \includegraphics[width=\textwidth]{agglomerative_A.pdf}
  \caption{Agglomerative Clustering}\label{fig:agglomerative_A}
  \end{subfigure}
  \caption{To reduce the state space, we prefer agglomerative clustering over Bayesian
  Gaussian mixture modeling. In the plots above we show the results of clustering the
  diagonal entries of the autoregressive coefficient matrices, $A$, of methanol. (a) 
  Bayesian Gaussian mixture models tend to delocalize the clusters in parameter space
  since the Gaussians can overlap. Agglomerative clustering prevents overlap of 
  clusters and ensures that all states within each cluster have similar parameters.
  }\label{fig:clustering_choice}
  \end{figure}

  \newpage

  \section{Choosing the Number of Clusters}\label{section:nclusters}

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{silhouette_MET.pdf}
  \caption{methanol}\label{fig:silhouette_MET}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{silhouette_URE.pdf}
  \caption{urea}\label{fig:silhouette_URE}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{silhouette_GCL.pdf}
  \caption{ethylene glycol}\label{fig:silhouette_GCL}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{silhouette_ACH.pdf}
  \caption{acetic acid}\label{fig:silhouette_ACH}
  \end{subfigure}
  \caption{We used the silhouette score test in order to quantitatively
  inform our decision on the numbrer of clusters. Silhouette scores 
  closer to one indicate good clustering. The crosses on the plot indicate
  our choice in number of clusters after careful consideration of the
  silhouette scores in conjunction with qualitative examination of the clusters
  formed by the choice. Although the test often
  favors fewer states, we elected use intermediate number of states in 
  the cases of $A$ and $\Sigma$ because choosing 2--3 clusters 
  insufficiently separates the dynamical modes. }\label{fig:silhouette}
  \end{figure}

\end{document}