### 2020-04-13 Ben_Manuscripts/thesis/thesis.tex

\documentclass[defaultstyle,11pt]{thesis}

\usepackage{amssymb}		% to get all AMS symbols
\usepackage{graphicx}		% to insert figures
\usepackage{hyperref}		% PDF hyperreferences??
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{textcomp} % gets rid of \perthousand and \micro warnings that come from gensymb
\usepackage{gensymb}
\usepackage{subcaption}
\usepackage{tcolorbox}
\usepackage[export]{adjustbox}
\usepackage{listings}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{diagbox}
\usepackage{array}

\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\newcolumntype{C}[1]{>{\PreserveBackslash\centering}p{#1}}
\newcolumntype{R}[1]{>{\PreserveBackslash\raggedleft}p{#1}}
\newcolumntype{L}[1]{>{\PreserveBackslash\raggedright}p{#1}}

\DeclareUnicodeCharacter{03C9}{$\omega$}
\DeclareUnicodeCharacter{03C0}{$\pi$}

\definecolor{amber}{rgb}{0.99, 0.76, 0.0} % golden poppy

\title{Understanding Structure and Selectivity in Nanostructured Lyotropic Liquid Crystal Membranes using Molecular Simulations}

\author{Benjamin Joseph}{Coscia}

\otherdegrees{B.A., University of Connecticut, 2014 \\
	          M.S., University of Colorado Boulder, 2017}

\Degree{Doctor of Philosophy}		                %  #1 {long descr.}
	   {Ph.D., Chemical Engineering}		%  #2 {short descr.}

\dept{Department of}			                %  #1 {designation}
	 {Chemical and Biological Engineering}		%  #2 {name}

\advisor{Prof.}				        %  #1 {title}
	    {Michael Shirts}			%  #2 {name}

\reader{Prof. Richard Noble}		%  2nd person to sign thesis
\readerThree{Prof. Douglas Gin}		%  3rd person to sign thesis
\readerFour{Prof. Matthew Glaser}
\readerFive{Prof. Robert Davis}

\abstract{

  % \OnePageChapter	% use if abstract is less than one page

  Periodic, nanostructured porous polymer membranes made from the cross-linked
  inverted hexagonal phase of self-assembled lyotropic liquid crystals (LLCs) are
  a promising class of materials for selective separations. There is a wide 
  design space available for liquid crystal monomer design which may allow one to
  tailor solute interactions within the membrane pores. Our currrent
  understanding of these membrane is based primarily on experimental measurements
  and macroscopic modeling which limits our insight and ability to design them 
  for specific applications.

  In this thesis, I develop a detailed atomistic molecular model, maximally consistent
  with experimental structural data, in order to investigate structure and transport
  of solutes within LLC polymer membranes. I characterize the dynamics of water, sodium
  ions and 20 small polar solutes within its pores. I present evidence which suggests that 
  LLC membranes are characterized by a gradual, rather than abrupt, transition in 
  composition from the primarily aqueous pore region to the hydrophobic monomer
  tail region. Within this continuous and heterogeneous environment, solutes perform
  intermittent hops between lengthy periods of entrapment. I show how transport rates
  of these species are dependent not only on molecular size, but on chemical functionality
  as well.

  In order to connect these microscopic chemically-dependent solute motions to
  macroscopic membrane performance, I develop three different stochastic models 
  of varying complexity.  
  First, using anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time
  random walk. We parameterize the distribution of dwell times, hop lengths between
  dwells and correlation between hops. Next, we generalize Markov state models, treating
  the configurational states of the system as a Markov process where each state has 
  distinct transport properties. Finally apply the infinite hidden Markov model, 
  an approach capable of automatically distinguishing and parameterizing an unknown 
  number of hidden dynamical modes in multidimensional solute time series. This model
  offers a flexibile approach to modeling solute motion and helps us to uncover 
  mechanisms with minimal human effort.

  }

\dedication[Dedication]{	% NEVER use \OnePageChapter here.
	To my girls Jenna, Maclynn and Aria.
	}

%  Here's where you acknowledge folks who helped. But keep it short, i.e., no more than one page,
%  as required by the Grad School Specifications.
\acknowledgements{	
  \OnePageChapter	% *MUST* BE ONLY ONE PAGE!

  I would like to acknowledge and thank a number of people whose support and
  guidance has not gone unnoticed. First, I thank my advisor, Michael Shirts, 
  who has been instrumental to my growth as a competent and confident researcher,
  pushing me to do high quality science. I thank the entire Shirts group for 
  their feedback and encouragement as well as for making our lab an enjoyable
  place to work. I thank Richard Noble and Douglas Gin for their guidance on past
  and present experimental research on LLC membranes as well as their
  students Michael McGrath, Gregory Dwulet and Sarah Dischinger, for providing 
  invaluable experimental perspective. I thank
  Matthew Glaser for his contributions to my understanding of liquid crystal systems.
  Our first paper would not have been complete without his insight. I thank Joseph
  Yelk for his effort creating the X-ray diffraction code on which our first paper
  relied. Finally, I thank Christopher Calderon for his guidance on applying 
  Bayesian non-parametric techniques to solute time series, allowing a leap in
  the flexibility of our dynamical modeling approaches.

  Outside of the University, I was fortunate enough to have full support of my
  friends and family. Moving to Colorado was a big leap and I would not
  have been successful without their encouragement.

  }

% BJC: The list of figures and tables is obnoxious because the captions are so long.
% BJC: Can I leave them out? (would just uncomment \emptyLoF and \emptyLoT)

% \IRBprotocol{E927F29.001X}	% optional!

% \ToCisShort	% use this only for 1-page Table of Contents

% \LoFisShort	% use this only for 1-page Table of Figures
\emptyLoF	% use this if there is no List of Figures

% \LoTisShort	% use this only for 1-page Table of Tables
\emptyLoT	% use this if there is no List of Tables

\begin{document}

\graphicspath{{../transport/figures/}{../structure_paper/figures/}{../transport/supporting_figures/}{../stochastic_transport/figures/}{../stochastic_transport/supporting_figures/}{../hdphmm/figures/}{../hdphmm/supporting_figures/}}

\input macros.tex
\input intro.tex

\input structure.tex
%\input structure_si.tex

\input transport.tex
%\input transport_si.tex

\input stochastic_transport.tex

\input hdphmm.tex

%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{ieeetr}	% or "siam", or "alpha", etc.
%\nocite{*}		% list all refs in database, cited or not
\bibliography{refs}		% Bib database in "refs.bib"

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
\appendix
\input structure_si.tex
\input transport_si.tex
\input stochastic_transport_si.tex
\input hdphmm_si.tex

\end{document}