### 2020-07-12 Ben_Manuscripts/stochastic_transport/response_to_reviewers.tex
Incorporated reviewer responses into main document

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{xr}
\usepackage{xcolor}
\usepackage{siunitx}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{scrextend}
\usepackage[normalem]{ulem}

\externaldocument[S-]{Supporting_Information}
\externaldocument[M-]{Final_Draft}

\begin{document}

\graphicspath{{./figures/}}

\begin{center}
\textbf{Response to reviewers: Capturing Subdiffusive Solute Dynamics and Predicting
Selectivity in Nanoscale Pores with Time Series Modeling} \\
Authors: Benjamin J. Coscia and Michael R. Shirts
\end{center}

We thank the reviewers for carefully reading over our manuscript and providing
thoughtful and constructive feedback. We have taken the suggestions into consideration 
and made appropriate revisions to the manuscript document. The comments have been reproduced
in italics and all changes to the text have been documented below them. Any newly added
references are given in the bibliography of this document and are placed in the main 
text with appropriate numbering.

\section*{Response to Reviewer 1}

\begin{enumerate}[label={Comment \theenumi :}, leftmargin=3.9\parindent]

  % comment 1
  \item \textit{\textbf{Page 5 line 19:} In this work, we use the output of our MD simulations
        to construct \textbf{two classes} of mathematical models which aim to predict membrane 
        performance while providing quantitative mechanistic insights. \\
        \textbf{Page 5 line 47:} We treat our system in terms of \textbf{two well-known} classes of 
        anomalous subdiffusion: \textbf{fractional Brownian motion} (FBM) subordinate to a 
        \textbf{continuous time random walk (CTRW)}, or subordinated FBM (sFBM) for short. \\ \\
        I am a little confused by these sentences, does the two well-known classes correspond to FBM
        and CTRW, or refer to different classes of modeling approaches used later in the manuscript?}

  Author reply: We thank the reviewer for identifying our potentially confusing choice of words.
  The first sentence (Page 5, line 19) refers more broadly to the two types of stochastic models 
  formulated in this work: the anomalous diffusion model and the Markov state-dependent dynamical
  model. The second quoted sentence (Page 5, line 47) refers to the different types of
  motion (FBM and CTRW) that give rise to anomalous diffusion specifically. The paragraph expands
  upon the introduction given in the previous paragraph. We have modified the second quoted 
  sentence as follows to remove ambiguity.

  Original text:
  \begin{quote}
  We treat our system in terms of two well-known classes of anomalous subdiffusion: 
  \end{quote}

  New text:
  \begin{quote}
  In this first approach, we treat our system in terms of two well-known types of molecular
  motion which lead to anomalous subdiffusion: 
  \end{quote}

  % comment 2  
  \item \textit{\textbf{MD-section} \\
        \textbf{Page 8 and 9} \\
        The authors refer to ref 13 (where these simulations have been performed before), 
        nevertheless I find that the MD section does not provide sufficient amount of details 
        to be self-contained. Could the authors specify the force field, as well as summarize 
        its reliability for this particular system (often simulations involving changed solutes
        through pores require polarizable force fields)?
   }

  Author reply: We thank the reviewer for pushing us to include more details about our 
  molecular simulations in order to boost reproducibility. The length of the manuscript made
  us wary to include information that could be accessed elsewhere, however there are 
  benefits to keep this very important section self-contained.

  We added the following text to address comments related to our choice of force field:

  New text:
  \begin{quote}
  We modeled water molecules using the TIP3P water model,~\cite{jorgensen_comparison_1983}
  We used the antechamber module shipped with AmberTools18 to parameterize monomers and 
  solutes using the Generalized Amber Force Field (GAFF).~\cite{wang_development_2004} We 
  assigned charges using the AM1-BCC model.~\cite{jakalian_fast_2000,jakalian_fast_2002}
  GAFF is very commonly used in simulations with diverse organic chemical functionality, 
  and was developed for consistency within the AMBER ecosystem, which used TIP3P as the
  water model. We chose these standard force fields and generalizable parameterization 
  approaches so that a range of solutes and monomers could be self-consistently and easily 
  studied. The main goal of this work is to develop stochastic models which can reproduce
  solute behavior on MD time scales, independent of the choice of force field. In future 
  work, it may be beneficial to employ specialized force fields, including polarizable 
  force fields or other higher-order effects in order to improve the accuracy of structure 
  parameters and selectivity predictions.
  \end{quote}

  % comment 3
  \item \textit{Next, reading this section I would also like to know how well finite 
			  size effects are controlled for with this specific membrane and these 
			  specific solutes.}

		Author reply:

        %MRS: so I'm looking at
        %https://pubs.acs.org/doi/10.1021/jp0477147 and .  I think the
        %key point here is that most of the finite size corrections
        %have to do with the viscosity of the bath that the solute is
        %in.  I THINK that we can consider the viscosity essentially infinite, or at least
        %really really high, and therefore the corrections

        We have not made any explicit finite size corrections, but they are less 
        important for this work than for diffusion in other soft matter systems 
        where diffusion is studied. In our previous work, we have shown that 
        monomer diffusion is negligible on simulation time scales. Finite-size 
        corrections to diffusion are primarily due to the hydrodynamics of the 
        surroundings~\cite{venable_lipid_2017,yeh_system-size_2004}, and the LLC
        membrane (outside of intramolecular motion and quasiharmonic vibration 
        around its average position) is therefore mostly motionless, with a very
        high effective viscosity. This high surrounding viscosity should nearly 
        eliminate these typical finite size effects. Additionally, we are usually 
        comparing ratios of diffusion when we look at selectivity, and thus finite 
        size effects should also mostly cancel out there.

        Verifying that there are indeed negligible finite size effects may be 
        interesting in future work, but we do not feel it is justified for this 
        study given the computational cost of studying larger systems. We also 
        believe it would have little influence on the way in which we fit our models
        to the time series.

  %MRS: I think it's always good practice to have a text change
  %corresponding to every reviewer comment, even if it's just to
  %clarify.  This proves that MRS: we are taking the reviewer
  %seriously.

  %BJC: MRS, can you check my placement of this paragraph in main text (search "r1_comment3" in Final_Draft_JCTC.tex)
  To clarify this logic, we have added the following text and citations:
  \begin{quote}
    Although we studied diffusion, we did not explicitly account for finite size effects
    in this system. Finite size corrections to diffusion are primarily due to the
    hydrodynamics of the surroundings~\cite{venable_lipid_2017,yeh_system-size_2004}, and
    as discussed in our previous work~\cite{coscia_understanding_2019} the LLC membrane
    (outside of intramolecular motion and quasiharmonic vibration around it's average 
    position) is mostly motionless, and therefore has a very high effective viscosity.
    Nonetheless, studies of finite size effects in this system may be worthwhile in
    future work. However, it likely would not affect the modeling approach described in
    this work.
  \end{quote}

  % comment 4      
  \item \textit{Lastly, it would be natural to report the size of the system in this
              section (I found it eventually in the introduction after searching for it 
              throughout the manuscript).}

              Author reply:

              We thank the reviewer for pointing out this missing piece of information. We
              have added the following text in order to inform the reader of the system's size:
			  \begin{quote} 
			  Each unit cell contains $\sim$ 62,000 total atoms.
			  \end{quote}

  %comment 5
  \item \textit{\textbf{Figure 5 page 25:} \\
		First, the symbols within the figure are very small and barely readable, and probably not
		according to authors guidelines for figures. Second, the caption, as with the other captions of
		this manuscript, contain a lot of interpretation that normally are placed in the main text. This
		might be a particular style of the authors that has its benefits, however in this specific case the
		caption fails at explaining what exactly this figure is showing. The way I interpret this figure is:
		the symbol on the y-axis marks a general parameter and the boldface symbol marks the
		parameter shown in the histogram. Am I correct? If so, I would strongly recommend modifying
		y-axis and the caption to make this clear.}

		Author reply:

		We thank the reviewer for their constructive feedback. The fontsize in the figure is indeed too small.
		We have modified the plots in the figure to have larger, more readable font. We also took this 
		opportunity to increase the font size on Figures 6--8.

		The reviewer is correct in their interpretation of the figures. For clarity, we modified the y-axes as
		well as the caption in order to further aid the reader in reading them. We also made efforts to cut
        down on the interpretation that is already presented in the main text. We made similar modifications
        to Figure 6 since it is an analogous figure. Our caption changes are presented below.

		This comment caused us to reconsider the detail in the captions of other figures as well. We feel
		the level of the detail is justified in most cases. However, we decided to remove some redundant
		information from Figure 13, as shown below.

		Figure 5, original caption:

		\begin{quote}

		  The parameters of the one mode model reveal differences in dynamics
	      between solutes. (a) We parameterized Gaussian,
	      $\mathcal{N}(\sigma)$, and L\'evy stable, $L(\sigma, \alpha_h)$,
	      distributions to describe solute hop lengths. We assume the mean
	      ($\mu$) to be zero for these distributions and no
	      skewness ($\beta = 0$) in the L\'evy stable distributions. High
	      values of $\sigma$ and lower values of $\alpha_h$ result in larger
	      hops. (b) We parameterized a pure power law, $P(\alpha)$, and a
	      truncated power law, $P_T(\alpha, \lambda)$, distribution to describe
	      solute dwell times. Lower values of $\alpha$ lead to heavier power
	      law tails and higher values of $\lambda$ truncate the distribution at
	      lower dwell times. (c) Finally, we parameterized the hop
	      autocorrelation function, $\gamma(H)$, to describe the degree of
	      correlation between hops. Simulations with higher values of $H$ display
	      behavior closer to the Brownian limit.

		\end{quote}

		Figure 5, modified caption:

		\begin{quote}

		  The parameters of the one mode model reveal differences in dynamics
	      between solutes. (a) We parameterized Gaussian, $\mathcal{N}(\sigma)$, 
	      and L\'evy stable, $L(\sigma, \alpha_h)$, distributions to describe
	      solute hop lengths. 
		  % BJC: added	      
	      Each bar represents the value of a single parameter,
	      highlighted in bold, of the associated hop distribution.
		  % BJC: removed	      
	      % We assume the mean ($\mu$) to be zero for these distributions and no
%	      skewness ($\beta = 0$) in the L\'evy stable distributions. 
	      In general, higher values of $\sigma$ and lower values of $\alpha_h$ 
	      result in larger hops. 
	      (b) We parameterized 
	      % BJC: removed. I think this was left over from when the pure power law was still in the paper
	      % a pure power law, $P(\alpha)$, and 
	      a truncated power law distribution, $P_T(\alpha, \lambda)$,  to describe
	      solute dwell times. 
		  % BJC: not sure if should reiterate the bold highlights here.	      
	      Lower values of $\alpha$ lead to heavier power
	      law tails and higher values of $\lambda$ truncate the distribution at
	      lower dwell times. (c) Finally, we parameterized the hop
	      autocorrelation function, $\gamma(H)$, to describe the degree of
	      correlation between hops. Simulations with higher values of $H$ display
	      behavior closer to the Brownian limit.

		\end{quote}

		Figure 6, original caption:

	    \begin{quote}

		  The two mode model parameterizes solute behavior in the pore and tails
	      separately. We consider solutes to be within the pore region if they
	      are 0.75 nm from a given pore center, otherwise they are in the
	      tails. (a) Generally, movement is much more restricted in the tail
	      region, parameterized by lower $\sigma$ values (smaller hops) for the
	      Gaussian and L\'evy stable distributions. Values of $\alpha_h$ are
	      significantly lower for urea and acetic acid meaning there is a
	      larger probability that they will take large hops. (b) Dwell times
	      are longer in the tails. Lower values of $\alpha$ correspond to power
	      laws with heavier tails and thus higher probabilities of long dwell
	      times. There is no easily discernible trend in $\lambda$ of the
	      truncated power law distribution. Note that we used the same Hurst
	      parameter for both modes (shown in Figure~\ref{M-fig:hurst_barchart})
	      due to a low number of sufficiently long sequences of hops in each
	      mode. (c) Solutes spend various amounts of time in the tail and pore
	      region dependent on their size, shape and chemical functionality.
	      Methanol's small size favors occupation of the much larger accessible
	      volume in the tails. Urea and acetic acid are fairly stable in both
	      regions since they are small and hydrophilic. Ethylene glycol has a
	      slight preference for the pores likely because it is a larger
	      molecule with two hydrophilic hydroxyl groups.

        \end{quote}

	    Figure 6, modified caption:

	    \begin{quote}
		  % Figure 6 caption
		  The two mode model parameterizes solute behavior in the pore and tails
	      separately. We consider solutes to be within the pore region if they
	      are 0.75 nm from a given pore center, otherwise they are in the
	      tails. 
	      %BJC: added next two sentences
	      For (a) and (b), each bar represents the value of a single parameter, 
	      highlighted in bold, of the associated hop or dwell time distribution.
	      Parameterizations of solute motion in the pores are visualized as bars
	      in the upper half of the plots, while the parameters describing motion 
	      in the tails are reflected into the lower half of the plots.
	      %BJC: modified this sentence to cut down on detail
	      (a) Generally, movement is more restricted in the tail region, but 
	      with occasionally large hops, as implied by lower $\sigma$ (smaller
	      hops) and $\alpha_h$ (heavier tails of the hop distribution) values.
	      %for the Gaussian and L\'evy stable distributions. 
%	      Values of $\alpha_h$ are significantly lower for urea and acetic acid
%	      meaning there is a larger probability that they will take large hops.
	      (b) Dwell times are longer in the tails as implied by lower values 
	      of $\alpha$.
	      %correspond to power laws with heavier tails and thus higher probabilities of long dwell
	      %times. 
	      There is no easily discernible trend in $\lambda$ of the
	      truncated power law distribution. Note that we used the same Hurst
	      parameter for both modes (shown in Figure~\ref{M-fig:hurst_barchart})
	      due to a low number of sufficiently long sequences of hops in each
	      mode. (c) Solutes spend various amounts of time in the tail and pore
	      region dependent on their size, shape and chemical functionality.
	      %BJC: removing details that is already in the text
%	      Methanol's small size favors occupation of the much larger accessible
%	      volume in the tails. Urea and acetic acid are fairly stable in both
%	      regions since they are small and hydrophilic. Ethylene glycol has a
%	      slight preference for the pores likely because it is a larger
%	      molecule with two hydrophilic hydroxyl groups.

        \end{quote}

        Figure 13, modified caption:

        \begin{quote}

          The selectivity between pairs of species changes monotonically with
	      pore length. The strength of dependence on pore length depends on
	      the difference between $\beta$ values. 
		  %BJC: All of this is in the text.	      
	      \sout{The largest differences in solute
	      flux result in high selectivities at any pore length. This membrane
	      may be a good candidate for the separation of ethylene glycol from
	      acetic acid. Ethylene glycol has the lowest $\beta$ value while
	      acetic acid has the second highest, leading to strong length
	      dependence. Ethylene glycol also has the highest flux and acetic acid
	      has the lowest resulting in relatively high selectivities independent
	      of pore length.}

        \end{quote}

  % comment 6
  \item \textit{\textbf{Page 6, line 55}: The second measure of success is based on the qualitative
        comparison between individual realizations of solute trajectories generated ...\\ \\
		Here I am a little curious. Are there no measures which can be used to quantify the similarity of
		the trajectories, such as autocorrelation functions etc, or are such measures inputs for the
		parametrization of the model which cannot be used?}

		Author reply: In short, we do not believe there are any other suitable methods for comparing
		similarity beyond the checks which we show in this work. Autocorrelation is indeed an input 
		to the anomalous diffusion and Markov state-dependent dynamical models, specifically through
		the Hurst parameter which is generated by fitting to autocorrelation of solute jumps. The
		anomalous diffusion model is only parameterized by the distribution of hop length, the 
		distribution of dwell times and the degree of autocorrelation between hops, all of which
		are measured directly from MD. It may be possible to think of other metrics of success
		for the Markov state-dependent dynamical model, but it fails the qualitative test and
		thus does not merit any further checks.

		%BJC: I don't think we need to modify the text for this comment?

  % comment 7
  \item \textit{One question that came to mind reading the paper was: why is the model only in 
  		one dimension and not two (for example using cylindrical coordinates)?}

    %MRS: we probably want to phrase it a bit differently. I'm thinking you should you lead with the fact that it is quasi-2D, since radial information is built into having different states in the pores and the tails for the MSDDM and the 2-state anomalous diffusion models, then go into the other details.  Thoughts?
    %BJC: modified to lead with quasi-2D

  		Author reply: We thank the reviewer for their sensible question. First, we would like
  		to point out that the two mode anomalous diffusion model is a quasi-2D 
  		modeling approach. In the paper, we do acknowledge that the radial location of solutes has
  		an influence on solute hopping behavior. However, we are primarily interested in axial 
  		transport, therefore the radial behavior is only important to the extent that it
  		affects solute motion in that direction.

		From a more mathematical standpoint, there are two factors which pose challenges with
		increasing dimensionality. First, we have a relatively low amount of data, and therefore
		it is challenging to adequately parameterize behavior as a continuous function of both 
		$r$ and $z$. Second, it is not trivial to simulate fractional motion in multiple dimensions.
		We were fortunate to find an efficient python implementation of FBM and we needed to
		develop our own approximation of 1D fractional L\'{e}vy motion based on literature. 
		We felt there was little to gain by models more complicated than those showcased in our
		work.

		However, we would like to acknowledge that we have a strong desire to incorporate data
		from multiple dimensions. We are actively investigating a new modeling approach which 
		allows us to overcome these issues and to use cylindrical representations with ease. 
		%MRS: could cite the arXiv paper here.
		%BJC: currently not on arXiv

  % comment 8
  \item \textit{\textbf{Page 35 line 12-14}: \\
   		We do not consider the two mode AD model because it has a broken correlation structure \\ \\
        I was a little disappointed to see only the single mode being used. Could the authors 
        elaborate on this point? Exactly how does this broken correlation structure hinder such
        an investigation?}

        Author reply: We are pleased with the reviewer's interest in the two-mode model. While it
        would be technically simple to apply the same mathematical framework to study the long time
        scale performance of the two-mode model, we feel it may be misleading to the reader.
        As described in the text, the predicted MSDs of the two-mode model may indeed show better 
        qualitative MSD behavior than the one-mode model. Specifically, the two-mode model becomes
        linear on long timescales which may be more consistent with our MD results. However, this 
        long time scale linear behavior of the two-mode model has no physical basis. As discussed
        in the text, the one-mode model has persistent curvature since hop correlation persists 
        indefinitely for fractional motion. However, for the two-mode model, each time a solute 
        transitions between the pore and tail regions, we must generate a new FBM process because
        we lack a proven technique for simulating fractional motion that incorporates time-dependent
        variation in the width of the hop length distributions.

        Perhaps, most importantly, Section 3.3 is primarily used to demonstrate our mathematical
        framework. Applying it to the two mode model would be trivial. We chose to use the one-mode 
        model for demonstration because it passes our qualitative and quantitative checks and is
        mathematically sound. As the reviewer points out, it is likely that our approach to 
        modeling the solute time series is not yet suitable to make quantitative predictions of 
        flux and selectivity. But our approach opens the door for applying the same analysis to 
        improved time series modeling techniques.

        Original text:
        \begin{quote}
        We do not consider the two mode AD model because it has a broken correlation structure 
        and we do not consider the MSDDM because its realizations do not display the expected 
        hopping and trapping behavior. 
        \end{quote}

        New text:
        \begin{quote}
        Although it would be trivial to test the two mode model as well, we chose not to because
        we feel the results would be misleading to the reader since the stochastic trajectories
        generated by this method lack a firm physical basis as discussed in Section~\ref{M-section:AD_all_data}. 
        We also do not consider the MSDDM because its realizations do not display the expected 
        hopping and trapping behavior. 
        \end{quote}

  % comment 9
  \item \textit{\textbf{Page 35 line 20-42:}\\
        A pore-length of 50 nm is a little underwhelming compared to 7 micrometers. Reading the paper
		I do not understand what lies behind this limitation - what hinders the sampling one really would
		like to have when doing such modeling? Also, you write about RAM-requirements scaling more
		than linearly (assuming with L). What exactly lies behind this scaling? Finally, just out of
		curiosity, are there ways of coarse-graining/renormalizing the random walk such that larger time
		and length scales can be reached? That being said, I am impressed by fit against the scaling
	  	law.}

	  	Author reply: We can appreciate why the reviewer has questioned this limitation.
		When simulating fractional motion, each step must be correlated to all of its previous 
		steps. Therefore the computational cost increases with each added step. We show in the paper
		that the mean first passage time across the pores scales with L$^{\alpha}$, and $\alpha$ is
		always $\geq$ 2. As the degree of anti-correlation between hops increases, the scaling
		parameter also increases. Therefore, we need to generate very long trajectories, with 
		correlation between all data points, in order to sufficiently sample the MFPT distribution 
		before fitting them to Equation 9. This is especially challenging with slow-moving solutes
		like acetic acid. We needed to simulate over 2 million timesteps in order to sufficiently
		sample the MFPT over a 50 nm pore. This procedure becomes computationally expensive very
		quickly, and requires ample RAM to store the trajectories as they are constructed.

		Given the quality of our fits in Figure 11b, we felt it was not necessary to carry out the
		calculation for longer pores which would likely offer little extra benefit to the insight we
		already gained.

	  	This may be an issue that could be overcome by developing specialized algorithms and perhaps 
	  	by using a compiled computer language like C++.

        We have modified the text to express the current limitations in a more straightforward manner:

	  	Original text:
	  	\begin{quote}
	  	The RAM requirements and performance scales greater than linearly and thus would take an 
	  	infeasible amount of memory and time to simulate transport through a pore over 100 times
        longer. 
	  	\end{quote}

	  	Modified text:
	  	\begin{quote}
	  	At this time, it is infeasible for us to simulate transport through a pore over 100 times
	  	longer without investing significant effort into writing and optimizing new code that can
	  	handle simulation of large trajectories with correlation between all steps.
	  	\end{quote}

  % comment 10        
  \item \textit{In this paper the models are only evaluated by their ability of achieving 
		  self-consistency with the trajectories used as input. This is conceptually not ideal. 
		  Are there no experimental data or mathematical models which act as a reference for 
		  the models? Please elaborate.}

		  Author reply: The reviewer makes a valid point. This work started in parallel with experimental
		  work which promised solute flux and selectivities that we could compare directly to our
		  simulations. However, synthesizing an aligned hexagonal phase membrane has been a practical
		  challenge for our experimental collaborators and much of their focus has shifted to 
		  membranes formed by liquid crystals that self-assemble into bicontinuous cubic architectures
		  which are much simpler to synthesize. We have also started to shift our focus towards models 
		  of the bicontinuous cubic phase, which do have much more experimental data. However, our 
		  research into generating physically accurate molecular models of that phase lags our work on
		  time series modeling. In future work, we will be able to apply this technique to these 
		  bicontinuous systems for which we have experimental data.

		  %BJC: I'm not sure this is worth bringing up

                  %MRS: probably should bring them up, since it allows you
                  %to respond to the question, and point out they are
                  %really insufficient because of he level of chemical
                  %detail.  Can just put them in the introduction
                  %describing the sort of model we want to move beyond.

		  As far as mathematical models, there are none that we know of to which we can reasonably
		  compare. Some work has been done to apply Donnan steric pore models to experimental data, 
		  but these models are continuum-based and have not been used in a predictive capacity.

		  We have added a mention of these mathematical models to the introduction in order to give
		  a more complete picture of the modeling approaches which preceded molecular dynamics:
		  \begin{quote}
		    Some work has been done to apply Donnan steric pore models to experimental LLC membrane 
		    data, but these models are continuum-based and are not effective as tools for explaining
		    complex solute-membrane interactions.~\cite{hatakeyama_water_2011}
		  \end{quote}

  \end{enumerate}

\noindent Optional comment to consider:\\
\begin{addmargin}[5.8em]{0em}
	\textit{Having read the paper thoroughly twice, I still struggle with the acronyms of the different random
	walks. Given that this is not a journal of stochastic processes, it is my opinion that the readability
	and accessibility of the manuscript would be greatly improved by writing the full names of the
	random walks in the text while keeping acronyms in figures. \\}

	We thank the reviewer for their constructive feedback. In response we have reduced our usage of
	acronyms for the various stochastic models in favor of their full unabbreviated form for clarity.

\end{addmargin}

\noindent Small typos to consider: \\
\begin{addmargin}[3.3em]{0em}
\begin{itemize}
  \item \textit{Micro in micrometer should be written with upright Greek letter.} \\ \\

  \item \textit{d as “dt” in integrals should be non italic.} \\ \\

  \item \textit{Page 10 line 28: T should be in symbol form.} \\ \\

  \item \textit{Symbol T is used for both time and transition matrix, perhaps use some matrix notation.} \\ \\
  %BJC: Maybe change matrix to bold italic? https://tex.stackexchange.com/questions/199789/which-bold-style-is-recommended-for-matrix-notation

  \item \textit{The references should be checked thoroughly by the authors. Ref 13, which is arguably
  the most important reference for this paper is missing the volume number.} \\ \\
\end{itemize}

\end{addmargin}

Author reply: We thank the reviewer for catching these errors and have corrected all of the instances
we could identify.

\section*{Response to Reviewer 2}

\begin{enumerate}[label={Comment \theenumi :}, leftmargin=3.9\parindent]

    % comment 1
    \item \textit{ Abstract: here it is difficult to understand what is meant with "dwell times, hop 
    lengths between dwells and correlation between hops", without reading the whole Introduction (and 
    possibly reference 13, which by the way is not completely listed in the References section). The 
    authors should briefly describe the complex transport behavior observed in their previous study of
    these systems (for example, by including information from the Introduction, i.e., the last paragraph
    of page 3 and first paragraph of page 4).}

    Author reply: We thank the reviewer for pushing us to increase the clarity of the abstract. We have
    made the following changes in order to give more immediate context to our work.

    Original text:

    \begin{quote}
    Mathematically modeling complex transport phenomena can be a powerful tool
    for extracting important physical information from molecular simulations. 
    In this study, we present two new approaches that use stochastic time series
    modeling to predict long time-scale behavior and macroscopic properties 
    from molecular simulation which can be generalized to other molecular systems
    where complex diffusion occurs. Specifically, we parameterize our models using
    long molecular dynamics (MD) simulation trajectories of a cross-linked 
    H\textsubscript{II} phase lyotropic liquid crystal (LLC) membrane in order to
    predict solute mean squared displacements (MSDs), solute flux, and
    solute selectivity in macroscopic length pores.    
    \end{quote}

	Modified text (highlighted in red):

	\begin{quote}
    %MRS: some suggested tweaks to describe the process more directly.
    \textcolor{red}{Using molecular simulations to fit mathematical models of complex transport phenomena
    that have more direct connection to experimental observables
    can be a powerful tool
    for extracting important physical information from these molecular simulations}. 
    In this study, we present two new approaches that use stochastic time series
    modeling to predict long time-scale behavior and macroscopic properties 
    from molecular simulation which can be generalized to other molecular systems
    where complex diffusion occurs. \textcolor{red}{In our previous work, we studied long 
    molecular dynamics (MD) simulation trajectories of a cross-linked 
    H\textsubscript{II} phase lyotropic liquid crystal (LLC) membrane where we
    observed subdiffusive solute transport behavior characterized by intermittent
    hops separated by periods of entrapment. In this work, we use our models to 
    parameterize the behavior of the same systems, so we can generate characteristic
    trajectory realizations that can be used} to predict solute mean squared 
    displacements (MSDs), solute flux, and solute selectivity in macroscopic length pores.
    \end{quote}

    % comment 2
    \item \textit{Some of the acronyms were not defined (or perhaps I missed them), for example 'AD' 
    (page 7) and 'MLE' (page 11)}

    Author reply:

    We thank the reviewer for pointing out these oversights. We have also defined MLE as maximum
    likelihood estimate before its first usage. In consideration of this comment and a similar
    comment by reviewer 1, we have decided to reduce the number of abbreviations in the main text 
    in order to eliminate confusion over acronyms.

    % comment 3
    \item \textit{In the Introduction, specifically the last paragraph of page 3 and first paragraph
    of page 4, the authors should refer to Figure 1 to complement their description of the complex solute
    dynamics.}

    Author reply:

    We thank the reviewer for this suggested improvement and we have added the appropriate
    figure reference with some modifications to maintain the overall sentence flow:

    Original text:
    \begin{quote}  
    We identified three mechanisms responsible for the solute trapping behavior: 
    \end{quote}

    Modified text:
    \begin{quote}
    The complex membrane topology (see~\ref{M-fig:membrane_structure}) gives rise to
    three mechanisms responsible for the solute trapping behavior: 
    \end{quote}

    \item \textit{Caption of Figure 7 should refer the reader to Table 2 for completeness}

    Author reply:

    We thank the reviewer for helping increase the clarity of the manuscript. We have added
    the following text to the caption of Figure 7:
    \begin{quote}
    The state numbers correspond to those described in Table 2.
	\end{quote}

\end{enumerate}

\bibliographystyle{ieeetr}
\bibliography{stochastic_transport}

\end{document}